% equilibria.tex
% Predrag extracted from newton.tex 		jul  3 2006
% Predrag					jun 20 2006
% Vaggelis					may 20 2006
% $Author$ $Date$

\section{Equilibria} % of the \KSe}
\label{sec:stks}

% Predrag                                       05dec2004
% Lan                                           25nov2004
% from Lan thesis                                8jun2004

\noindent
Equilibria 
% (or the steady solutions)
are the simplest invariant sets in
the phase space. They,  and 
the connections between them form the
coarsest geometrical framework for organizing
phase space orbits. %\rf{ksgreene88}.
% In certain parameter ranges 
% equilibria they play prominent role in observed dynamics.
% For example, the $n$-cell state
% is stable in discrete parameter windows
% windows for arbitrary large system sizes. %\rf{FSTks86}.

The equilibrium condition $u_t=0$ for the {\KSe} PDE \refeq{ks-L} 
is the ODE
\[
(u^2)_x-u_{xx}- u_{xxxx}=0 
% \,.
\]
which can be analyzed as a dynamical system in its own right.
Integrating once we get
\beq
u^2-u_x- u_{xxx}=c 
\,,
\label{eq:stdks}
\eeq
where $c$ is an integration constant 
whose value strongly influences the nature of
the solutions. %of \refeq{eq:stdks}. 
Written as a 3-dimensional dynamical system
with spatial coordinate $x$ playing the role of ``time'',
%\refeq{eq:stdks} 
this is a volume preserving flow
\beq
u_x = v \,,\qquad
v_x = w \,,\qquad
w_x = u^2-v-c \,,
  \label{eq:3dks}
\eeq
with the ``time'' reversal symmetry, 
\[
x \to -x,\quad u \to -u, \quad v \to v, \quad w \to -w \,.
\]
 From \refeq{eq:3dks} we see that
\[
(u+w)_x=u^2-c \,.
\]
If $c<0$, $u+w$ increases without bound with $x \to \infty$,
and every solution escapes to infinity.
If $c=0$, the origin $(0,0,0)$ is the
only bounded solution. 

For $c>0$ there is much
$c$-dependent interesting dynamics, with
complicated fractal sets of bounded solutions.
The sets of the solutions of the equilibrium condition 
\refeq{eq:3dks} are themselves in turn organized by the  
equilibria of the equilibrium condition, and 
the connections between them.
    For $c>0$ the equilibrium points of \refeq{eq:3dks} are
$c_{+}=(\sqrt{c},0,0)$ and $c_{-}=(-\sqrt{c},0,0)$.
Linearization of the flow around
$c_{+}$ yields 
stability eigenvalues 
$[ 2\lambda \,, -\lambda \pm i \theta ]$
with 
\[
\lambda=\frac{1}{\sqrt{3}}\sinh \phi
\,,\qquad
\theta=\cosh \phi \, ,
\]
and $\phi$ fixed by $\sinh 3\phi=3\sqrt{3c}$. 
Hence $c_{+}$ has a {1-dimensional}
unstable manifold and a 2-dimensional
stable manifold along which solutions spiral 
in. 
By the $x \to -x$ ``time reversal'' symmetry, the 
invariant manifolds of $c_{-}$ 
have reversed stability properties.

However, we do not need to explore the fractal set of the 
\KS\ equilibria for infinite size system here;
for a fixed system size
$L$ with periodic boundary condition, the only surviving equilibria  are
those with periodicity $L$.
They satisfy 
the equilibrium condition for \refeq{expan}
\PC{\rf{ksgreene88} to remarks}
\beq
(k/\tilde{L})^2\left( 1 - (k/\tilde{L})^2  \right)b_k 
	 + i (k/\tilde{L}) \sum_{m=-\infty}^{+\infty} b_m b_{k-m} = 0
\,.
\label{eq:stfks}
\eeq 
% We have proved in \refsect{sec:kspr} that 
Periods of spatially periodic equilibria are multiples of $L$.
Every time $\tilde{L}$ crosses an integer value  $\tilde{L}=n$,
$n$-cell states
are generated through pitchfork bifurcations. 
In the full phase space they
form an invariant circle due to the translational invariance of \refeq{ks-L}. 
In the antisymmetric subspace considered here, they corresponds to two points,
half-period translates of each other of the form
\[
u(x,t)=-2\sum_k b_{kn}\sin (knx) \,,
\]
where $b_{kn} \in \mathbb{R}$.
% By rescaling $u,x$ and $\nu$, the $n$-cell states transform to each other.

%      With the increase of $L$ these periodic solutions 
% may bifurcate into more complicated ones.
For any fixed period $L$
%, however, 
the number 
of spatially periodic solutions is finite up to a spatial translation.
This observation can be heuristically motivated as follows. 
% \PC{this argument keeps worrying me: there are lots of solutions, like
% $u=0$, that are equilibria, but isolated -
% they are noplace near asymptotic dynamics.
% Do they belong to the invariant manifold?
%    }
% Equilibria are solutions valid for all times, and are thus points
% on the finite-dimensional compact inertial manifold\rf{infdymnon}.
Finite dimensionality of the inertial manifold
bounds the size of Fourier components of all solutions.
% This
% compact inertial manifold and the dynamics on it can be 
% described by analytic functions of a finite number of Fourier modes.
\PC{explain the theory; say that in practice it is useless}
On a finite-dimensional compact manifold,
an analytic function can only have a finite number
of zeros. So, the equilibria, {\em i.e.}
the zeros of a smooth velocity field on
the inertial manifold, are finitely many.
% The number of equilibria increases exponentially with $L$,
% \PC{give reference for ``exponential growth''}
% for infinite system size $L \to \infty$,
% there are infinitely many equilibria. 
% \PC{is there a reference where this is explained?}

For a sufficiently small $L$ 
the number of equilibria is small,
mostly
concentrated on the low wave number end of the Fourier spectrum.
These solutions may
be obtained by solving the truncated versions of \refeq{eq:stfks}. 
% Understanding the structure of these solutions, requires a study
% of the full phase space of the 3-dimensional dynamical system \refeq{eq:3dks},
% not attempted here.

% Locally coherent structures are observed for arbitrary system size,
% see \reffig{f:flameFlut}~(b). %\reffig{f:ksev}.


\underline{Important \KS\ equilibria:}{ 
% \label{exam:KurSivEquil}
% \index{Kuramoto-Sivashinsky equilibria}
% \index{equilibria!Kuramoto-Sivashinsky}
%

	} %end \example{Important \KS\ equilibria



\PC{say somewhere: ``
The task of the theory is to describe this spatio-temporal
turbulence and yield quantitative predictions for its measurable
consequences.
   ''}



\section{Why does a flame front flutter?}
%\label{s:StabEqui}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Predrag           5jun2005
% extracted from \Chapter{stability}{ 2apr2005}
%       Predrag                  14/3-95
% taken from ks.tex
% Spatiotemporal chaos in terms of recurrent patterns
% PC            30/6/96 

% \index{local!stability}
% \index{linear!stability}
% \index{stability!linear}
% \index{equilibrium!point}

We start by considering the case where
$a_\stagn$ is an equilibrium point \refeq{EquilPoint}. 
Expanding around the equilibrium point $a_\stagn$,
and using the fact that the matrix
${\bf \Mvar}={\bf \Mvar}(a_\stagn)$
% , its matrix of its stability exponents
in \refeq{die}
% local expansion rate
is constant, 
we can apply the simple formula \refeq{eqPointStab} also to the 
{\jacobianM}
of an equilibrium point of a PDE, 
\[
 \jMps^t(a_\stagn) = e^{{\bf \Mvar} t}
    \,\qquad
 {\bf \Mvar}={\bf \Mvar}(a_\stagn)
\,.
\]

The \KS\ flat flame front $u(x,t)=0$ is an 
equilibrium point of \refeq{ks}. The matrix of variations
\refeq{DerMatrix}
follows from \refeq{expan}
% \index{Kuramoto-Sivashinsky system}
\beq
{\Mvar}_{kj}(a) ={\pde v_k(a)\over \pde a_j  }
=((k/\tilde{L})^2- (k/\tilde{L})^4)\delta_{kj} - 2(k/\tilde{L}) a_{k-j}
\,.
\ee{expanMvar}
For the $u(x,t)=0$ equilibrium  solution the matrix of variations
is diagonal, and -- as in \refeq{EqDyn164} -- so is the {\jacobianM}
$
\jMps^t_{kj}(0) = \delta_{kj} e^{((k/\tilde{L})^2- (k/\tilde{L})^4)t}
\,.
$

For $\tilde{L} < 1$,  $u(x,t)=0$ is the globally attractive stable 
equilibrium.
As the system size $\tilde{L}$  is increased,
the ``flame front'' becomes increasingly unstable and turbulent,
the dynamics goes through a rich sequence of
bifurcations on which we shall not dwell here.
% studied e.g. in \refref{KNS90}. 
% , one quickly finds a
% myriad of unstable periodic solutions whose number
% grows exponentially with period.

The $|k|<??$ 
long wavelength perturbations of the flat-front equilibrium
are linearly unstable, while all 
$|k|> ??$ short wavelength perturbations are strongly contractive.  
The high $k$ eigenvalues, corresponding to rapid variations of
the flame front, decay so fast that the corresponding eigendirections
are physically irrelevant.
% \index{Lyapunov exponent!equilibrium}
To illustrate the rapid contraction in the non-leading eigendirections
we plot  in [MAYBE INCLUDE] % \reffig{f:eigenvalues}
the eigenvalues of the equilibrium in the unstable regime,
for relatively small system size, % low viscosity $\nu$,
and compare them with the
stability eigenvalues of the least unstable cycle for the same 
system size.
% value of $\nu$. 
The equilibrium solution is very unstable,
in 5 eigendirections,
the least unstable cycle only in one. 
Note that for $k>7$ the rate of contraction
is so strong that higher eigendirections are numerically meaningless for 
either solution; even though the flow is infinite-dimensional, the attracting
set must be rather thin.

While in general
for $\tilde{L}$ sufficiently large
one expects many 
coexisting attractors in the phase space%
%Hyman and Nicolaenko
\rf{HNZks86} ,
in numerical studies most random initial
conditions settle converge to the same chaotic attractor. 

From \refeq{expan} we see that the origin $u(x,t) = 0$
has Fourier modes as the  linear
stability eigenvectors. 
When $|k| \in (0,\tilde{L})$, the corresponding Fourier modes are
unstable.
The most unstable modes has $|k|=\tilde{L}/\sqrt{2}$ and defines the scale of basic building
blocks of the spatiotemporal dynamics of the {\KSe} in large system size limit,
as shown in \refsect{sec:KSnumer}. 


\noindent
Consider now the case of initial $a_k$ sufficiently small
that the bilinear $ a_m a_{k-m}$ terms in \refeq{expan} can
be neglected. Then we have a set of decoupled linear
equations for $a_k$ whose solutions are exponentials, at most
a finite number for  which
$k^2 > \nu k^4$
is growing with time, and infinitely many with
$
\nu k^4 > k^2
$
decaying in time.
The growth of the unstable long wavelengths (low $|k|$) excites
the short wavelengths
through the nonlinear term in \refeq{expan}.  The excitations thus
transferred are dissipated by the strongly damped short wavelengths,
and a ``chaotic equilibrium'' can emerge. The very short
wavelengths $|k| \gg 1 / \sqrt{\nu}$ remain small for all times,
but the intermediate wavelengths of order $|k| \sim 1 / \sqrt{\nu}$
play an important role in maintaining the dynamical equilibrium.
As the damping parameter decreases, the solutions increasingly take on
% Burgers type
shock front
character poorly represented by the Fourier basis, and many
higher harmonics may need to be kept
% \rf{KNS90,GEP}
in truncations of
\refeq{expan}.


Hence, while one may truncate the high modes in the expansion
\refeq{expan}, care has to be exercised to ensure that no modes
essential to the dynamics are chopped away. 

In other words, even though our starting point
\refeq{ks}
is an infinite-dimensional dynamical system, the asymptotic dynamics
unfolds on a finite-dimensional attracting manifold, and so we are back on
the familiar territory of \refsect{SecDynFlows}:
the theory of a finite number of ODEs applies to this
infinite-dimensional PDE as well.

    {\bf When is an equilibrium important?} There are two kinds of roles
equlibria play:

{\em ``Hole'' in the natural measure}.
The more unstable eigendirections it has (for example, the
$u=0$ solution), the more unlikely it is  that
an orbit will recur in its neighborhood.

{\em unstable manifold of a ``least unstable''equilibrium}.
 Asymptotic dynamics
spends  a large fraction of time in
neighborhoods of a few  equilibria with
only a few unstable eigendirections.


\underline{Stability of \KS\ equilibria:}{ 
% \label{exam:KSEquilStab}
% \index{Kuramoto-Sivashinsky equilibria}
\begin{table}
\caption[]{
Important \KS\ equilibria:
% in the antisymmetric solution 
% space of the Kuramoto-Sivashinsky equation with periodic boundary % % % % condition,
% $ \nu =1$, $L=38.5$;
% their labels,
the first few stability exponents
%, with complex pairs written together.
}

\vskip 1.5cm

{\small
%\lineup
\begin{tabular}{@{}ccccc}
\hline %\br
$~S~~~$ & $~~~~\lambda_1 \pm \,i\,\theta_1$ 
                                & $~~~~\lambda_2 \pm \,i\,\theta_2$ 
                                        & $~~~~\lambda_3 \pm \,i\,\theta_3$ 
\\ 
\hline %\mr
${C_1}$    &{0.04422 $\pm \,i\,$0.26160}   &-0.255 $\pm \,i\,$0.431 
&-0.347 $\pm \,i\,$0.463         \\
% ${C_2}$    &0.33053  & 0.097 $\pm \,i\,$0.243 
% &-0.101 $\pm \,i\,$0.233        \\
\hline %\mr
${R_1}$   &{0.01135 $\pm \,i\,$0.79651} & -0.215 $\pm \,i\,$0.549 
&-0.358 $\pm \,i\,$0.262        \\
%  ${R_2}$   &  0.33223  & -0.001 $\pm \,i\,$0.703  
%  & -0.281 $\pm \,i\,$0.399      \\
\hline %\mr
${T}$     & 0.25480  & -0.07 $\pm \,i\,$0.645 &-0.264  
\\
\hline %\br
\end{tabular}
}
\label{t:stationary}
\end{table}

{\em 
spiraling out in a plane}, all other directions contracting


{\bf
Stability of ``center'' equilibrium
	}

linearized stability exponents: 
\[ % \beq
(\lambda_{1}\pm\,i\,\theta_{1},\lambda_{2} \pm\,i\,\theta_{2}, \cdots)
	= (0.044 \pm \,i\,0.262\,,\,
		-0.255 \pm \,i\,0.431\,,\,\cdots)
\] %\eeq

The plane spanned by $\lambda_{1} \pm\,i\,\theta_{1}$ eigenvectors rotates with angular period
$\period{} ~\approx~2\pi/\theta_{1}=24.02$.
% 2*4*a(1)/0.26160 = 24.0182924586375

a trajectory 
that starts near  the $C_1$~equilibrium point spirals 
away per one rotation
with multiplier
$\ExpaEig_{\mbox{radial}}~\approx~\exp(\lambda_{1}\period{})=2.9$.
% 2*4*a(1)/0.26160*0.04422 = 1.062
% e(1.0620888) = 2.8924063421

each Poincar\' e section return, 
contracted into the stable manifold by 
factor of
{
$\ExpaEig_{2}\approx\exp(\lambda_{2}\period{})=0.002$
}
%2*4*a(1)/0.26160*(-0.255) = -6.12466
%e(-6.12) = .0022


The local Poincar\' e return map is 
{\em
in practice $1-dimensional$
}
	} %end \example{{Stability \KS\ equilibria


% Staring at the solution
% as it evolves in time we should start getting a glimpse of the
% repertoire of the spatiotemporal patterns charcterizing
% the turbulent dynamics.


\underline{Model PDE systems.}{
The theorem on finite dimensionality of inertial manifolds
of phase-space contracting PDE flows is proven in \refref{Foias88}.
% \index{inertial manifold}
% \index{Kuramoto, Y.}
% \index{Sivashinsky, G.I.}
% \index{Kuramoto-Sivashinsky system}
The Kuramoto-Sivashinsky equation was introduced in \refrefs{ku,siv}.
Holmes, Lumley
and Berkooz\rf{Holmes96} offer a delightful discussion of why this system
deserves study as a staging ground for studying turbulence in 
full-fledged Navier-Stokes equation. 
How good 
a description of a flame front this equation
is need not concern us here; suffice it to say that such model
amplitude equations for interfacial instabilities arise in a variety
of contexts - see e.g.~\refref{saddks} - and this one is perhaps the
simplest physically interesting spatially extended nonlinear system.
\PC{Comment om MAWs, BECS and CGLe}
\PC{refer to Trefethen's program for fast integration}
This chapter
% \refsect{s_extend} 
is based on V. Putkaradze's term project
(see \wwwcb{/extras}).
%a gratifying example of a successful course project which led to a
%full-fledged research paper by
and on the Christiansen {\em et al.} article \rf{Christiansen:97}. 
\PC{refer the reader to \refref{Lan:Thesis} for a
 review of the $L \to \infty$ equilibria.
 Shouldn't Michelson\rf{Mks86} be given credit here?}

% \label{r:KuramShiva}
}

\underline{\KS\ system, truncations.}{
We describe here our criterion for reliable
truncations of the infinite ladder of 
ordinary differential equations (\ref{expan}).

Adding an extra dimension to a truncation of the system (\ref{expan})
introduces a small
perturbation, and this can (and often will) 
throw the system into a totally different asymptotic state. 
A  chaotic attractor for $N=15$ can become a period three 
window for $N=16$, and so on. 
If we compute, for example, the Lyapunov exponent
$\Lyap(\nu,N)$ for the strange attractor of the 
system (\ref{expan}), there is no reason to 
expect $\Lyap(\nu,N)$ to smoothly converge to the limit  
value $\Lyap(\nu,\infty)$ as $N \rightarrow \infty$. 
The situation is different in the periodic windows, 
where the system is structurally stable, and it makes sense to compute 
 Lyapunov exponents, escape rates, etc. for the 
{\em repeller}, \ie, the closure of the set of all 
{\em unstable} periodic orbits. 
Here the power of cycle expansions comes in: 
to compute quantities on the repeller by direct averaging methods is 
generally more difficult, because the motion quickly collapses to the 
stable cycle. 
	} %end \remark{Kuramoto-Sivashinsky system, numerical results.}


The problem one faces with high-dimensional flows is 
that their topology is hard to
visualize, and that even with a decent starting guess for a point on
a periodic orbit, methods like the Newton-Raphson method are likely to fail.
Methods that start with initial guesses for a number of points along the
cycle, such as the multipoint shooting method of \refsect{s-MultShoot},
are more robust.
The relaxation  (or variational) methods take this strategy to its
logical extreme,
and start by a guess of not a few points along a periodic orbit,
but a guess of the entire orbit.

At present the theory is in practice applicable only to systems
with a low intrinsic {\em dimension}
-- the minimum number of coordinates necessary to
capture its essential dynamics.
% \index{dimension!intrisic}
% \index{degree of freedom}
If the system is very turbulent
(a description of its long time dynamics requires a space of high
intrinsic dimension) we are out of luck. 
% \index{turbulence}

% \input{chapter/refsPDEs}

%%% end

