% history.tex
%
% Predrag created file				jul 3 2006
% $Author$ $Date$


\section{{\Rpo s} - a brief history}

Chenciner\rf{ChencinerLink}
says Poincar\'e introduced them in the 3-body problem.
Consider motion of a test particle of mass
$\mu \ll 1$ in the
restricted three-body problem\rf{rtb},
under the
influence of the gravitational force of two heavy bodies with masses $1$ and
$\mu \ll 1$ fixed at $(-\mu,0)$ and $(1-\mu,0)$. \Reqv\ of this problem
(known in this context as the Lagrange points, stationary in
the co-rotating frame) are circular motions in the inertial frame,
and {\rpo s} correspond to quasiperiodic motions. 

Lan has some relative equilibria (travelling waves) for KS in his
thesis\rf{Lan:Thesis}, %http://chaosbook.org/projects/theses.html
 and for complex LG in a paper on "MAWs".
Viswanath\rf{ViswanathPC06} % arXiv.org/physics/0604062
found them in the plane Couette problem.
Siminos has for Davidcheck box size $L=22$ some equlibria, and all
orbits his \rpo\ search finds are in the set you sent us.


The strategy
is to minimize an action functional defined on a space of paths $\gamma$ in the
configuration space $\pS$ which satisfy the property
\beq
                               \gamma (t + T ) = g \cdot \gamma (t)                       
\label{McC1}
\eeq
for a fixed {\em relative period} $T$ and some fixed diffeomorphism $g$ of $\pS$, 
sometimes referred to as a {\em phase}, that leaves the 
action functional invariant. If $g^k=1$ is of finite order
$k$, then the corresponding orbit is periodic with period $k T$. 

Striking applications of this idea has been the discovery
of ``choreographies" of $N$-body problems\rf{McC7,McC8,McC}.
\PC{some McCord references}

If $g$ does not have finite order $k$, the
paths can be periodic after the action of $g$. 
In either case, order we refer to the loops
in $\pS$ satisfying \refeq{McC1} as 
{\em relative loops} and
to the corresponding solutions of the equations of motion
as {\em \rpo}. 

    Let $A[\gamma]$ be an action functional defined loops $\Loop$
emebedded in the state space manifold $\pS$, invariant
under a diffeomorphism $g : \pS \to\pS$. Our aim is to determine
the critical points of the action functional
$ %\beq
	A[\gamma]
  % =\int_0^T  dt L(\gamma,\dot\gamma, t)
$ % \label{McC2}
  % \eeq
defined on $\Lambda_g^1(\pS)$, a Sobolev space of paths in $\pS$ satisfying 
\refeq{McC1}.
\PC{define Sobolev space as adding a vector space?}

such as the relative equilibrium solutions and
the relative periodic solutions. They respectively appear from
equilibria and periodic solutions of the Hamiltonian reduced by the symmetries.
They are in many physical situations, which include rigid bodies, gravitational
N-body problems, molecules and nonlinear waves.

