% energy.tex
% $Author$ $Date$


% \section{Representation in terms of symmetry invariant moments} % of the \KSe}
% \label{sec:energy}
% Predrag                   Apr 12 2007

In physical settings where the observation times are much longer
than the dynamical ``turnover'' and Lyapunov times (statistical mechanics,
quantum physics, turbulence) 
periodic orbit theory\rf{DasBuch}
provides highly accurate predictions of measurable
long-time averages such as the turbulent drag\rf{GHCW07}.
Physical predictions have to be independent of a
particular choice of ODE representation of 
the PDE under consideration
and, most importantly, 
invariant under all symmetries of the dynamics.
In this section we discuss
a set of such physical observables for
the  1$D$ KS invariant under reflections and translations.
They offer a representation of
dynamics in which the symmetries are explicitly quotiented out.
We illustrate  this in  \refsect{sec:energyL22} by projecting 
a set of explicit solutions on these coordinates. 

The {space average} of a function $\obser = \obser(\pSpace,t)$  on
the interval $L$,
\beq
    \expct{\obser} = \Lint{\pSpace}\, \obser(\pSpace,t)
    % \expct{\obser} = \frac{1}{L}\int_0^{L} d\pSpace\, \obser(\pSpace,t)
    \,,
    \label{rpo:spac_ave}
\eeq
is in general time dependent. 
Its mean value is given by the {time average}
\beq
\timeAver{\obser}
    =
\lim_{t\rightarrow \infty} {1\over t} \int_0^t \! d\tau \, \expct{\obser}
    =
\lim_{t\rightarrow \infty} {1\over t} \int_0^t \! 
    \Lint{\tau}  d\pSpace\, \obser(\pSpace,\tau)
    \,.
\label{rpo:tim_ave}
\eeq
The mean value
% $\timeAver{\obser}$,
of $\obser = \obser(u_q) \RLDedit{\equiv \obser_q}$ evaluated on
$q$ \eqv\ or {\reqv} $u(\pSpace,t) = u_q(\pSpace-ct)$ is
\RLD{I'm somewhat confused by the notations here.  Notions
with subscripts, such as $\expct{\cdots}_q$, are not defined, 
so I'm not sure if my changes make sense.}
\beq
%         \obser_q = \timeAver{\obser}_q = \expct{\obser}_q
\RLDedit{\timeAver{\obser}_q = \expct{\obser}_q = \obser_q }
\,.
\label{rpo:u-eqv}
\eeq
Evaluation of the infinite time average \refeq{rpo:tim_ave}
on a function of a 
\po\ or \rpo\ $u_p(\pSpace,t)=u_p(\pSpace,t+\period{p})$
 requires only a single $\period{p}$ 
traversal, 
\beq
       \RLDedit{\timeAver{\obser}_p} = \frac{1}{\period{p}}
    \int_0^{\period{p}} \! d\tau \, \expct{\obser}
\,.
\label{rpo:u-cyc}
\eeq

Equation \refeq{ks} can be written as % in ``potential'' form
\beq
    u_t=- V_x
        \,,\qquad
    V(x,t)={\textstyle\frac{1}{2}}u^2+u_{x} + u_{xxx}
    \,.
\ee{ksPotent}
$u$ is related to the ``flame-front height'' $h(x,t)$ by
$u=h_x$, so \expctE \RLDedit{, defined in \refeq{eq:stdks},} 
can be interpreted as the mean energy density.
% \refeq{ksEnergy}.
%
So, even though KS is a phenomenological
small-amplitude equation, the time-dependent quantity
\beq
    \expctE=
%\frac{1}{L}\int_0^{L} \!dx \, 
\RLDedit{\Lint{\pSpace}}
V(x,t)=
%\frac{1}{L}\int_0^{L}\! dx \,
\RLDedit{\Lint{\pSpace}}
\frac{u^2}{2}
\label{ksEnergy}
\eeq
% \PC{this is missing a prefactor 1/2. Check if someone in literature
% - other than Greene and Kim - defines it right. }
has a physical interpretation\rf{ksgreene88}
as the average ``energy'' density of the flame front.
This analogy to the corresponding definition of the
mean kinetic energy density for
the Navier-Stokes motivates what follows.
\PC{bit weird: can use Galilean invariance to
    set $\expctE=0$ for any given $u(x,t)$?}
\RLD{But it's the same as with kinetic energy, which is zero 
in the reference frame that is moving with the object. Right?}

The energy \refeq{ksEnergy} is intrinsic to
the flow, independent of the particular ODE basis set 
chosen to represent the PDE. However, as the Fourier
amplitudes are eigenvectors of the translation operator,
in the Fourier space the energy is a diagonalized
quadratic norm,
\beq
\expctE   % = (u,u) 
          = \sum_{k=1}^{\infty} E_k
\,,\qquad
E_k = % a_{-k} a_k =
    {\textstyle\frac{1}{2}}|a_k|^2
\,,
\ee{EFourier}
and explicitly invariant term by term 
under translations \refeq{eq:RPO}.

Take time derivative of the energy density \refeq{ksEnergy},
substitute \refeq{ks} and integrate by parts. Total derivatives vanish
by the spatial periodicity on the $L$ domain:
\bea
   \dot{\expctE} &=&
     \expct{u_t \, u}
    % \frac{1}{L}\int_0^{L}u \, u_t\, dx
     = - \expct{\left({u^2}/{2} + u \, u_{x} + u \, u_{xxx}\right)_x u }
    \continue
    &=&
\expct{+ u_x \, {u^2}/{2} + u_{x}{}^2 + u_x \, u_{xxx}}
    \,.
\label{rpo:ksErate}
\eea
Substitution by \refeq{KSeqvCond}
verifies that for an \eqv\ $\expctE$ is constant:
\[
   \dot{\expctE} =
\expct{ \left({u^2}/{2} + u_{x} + u_{xxx} \right) u_x}
    = \expctE \expct{ u_x }=0
    \,.
\]
The first term in \refeq{rpo:ksErate} vanishes by
integration by parts,
\(
 \expct{(u^3)_x} = 3 \expct{ u_x \, u^2}=0
\,,
\) % {EnNonl0}
and integrating the third term by parts yet again 
one gets\rf{ksgreene88} that the energy variation
\beq
   \dot{\expctE} = P - D
                \,,\qquad
      P =  \expct{u_{x}{}^2}
                \,,\quad
      D =  \expct{u_{xx}{}^2}
\ee{EnRate}
balances the power $P$ pumped in by anti-diffusion $u_{xx}$
against the energy dissipation rate $D$
by hyper-viscosity $u_{xxxx}$
in the KS equation \refeq{ks}.

% PC worked out with Bridges 
An infinite number of identities for moments of
solutions of KS follow\rf{Bridges_priv}. For example,
integrating by parts $\expct{u_{x} u_t}$,
$\expct{u_{xx} u_t}$,
and
$\expct{u^2 u_t}$,
respectively, one obtains \reqva\ relations
\PC{please re-derive these three. Are there more
    important ones that I have missed?
    Also, I - unnecessarily - specialized to \reqva,
    but these identities will be especially useful for \rpo s}
\RLD{I don't quite follow this process of deriving moments.  What is 
the reason we look for some combinations of $u$ and its derivatives?
Why not just take $\expct{u_{x\cdots x}}$?
Is there any physical motivation for such combinations, 
beyond $E$, $P$ and $D$?  Maybe use things like $\dot{P}$, 
$\dot{D}$, etc.?}
\bea
% \expct{u_{x} u_t} 
%     \qquad \to \qquad
\expct{u \, u_{x}{}^2} &=& c P 
\label{Bridges1}\\
% \expct{u_{xx} u_t} 
%     \qquad \to \qquad
  \expct{u_{xxx}{}^2} + {\textstyle\frac{1}{2}}\expct{u_{x}{}^3} &=&   P 
\label{Bridges2}\\
\expct{u \, u_{xx}{}^2} &=& c P 
\label{Bridges3}
\,.
\eea
When moments such as \refeq{Bridges1} are added as
coordinate axes to \reffig{f:drivedragConn}, \reqva\ and
\rpo s are separated from \eqva\ and \po s. Furthermore,
as higher moments have more and more powers of $u$ and derivatives
$u_{xx\cdots x}$, their magnitudes should be strongly suppressed, 
providing a symmetry-invariant basis set that can be safely truncated to 
a finite-dimensional \statesp.
See  \refsect{sec:energyL22} for how this works out in practice. 

The time averaged energy density  $\timeAver{E}$
computed on a typical orbit goes to a constant, so
the expectation values \refeq{rpo:EtimAve} of drive and dissipation
exactly balance each out:
\beq
    \timeAver{\dot{E}}  =
    \lim_{t\rightarrow \infty}
        {1\over t} \int_0^t d\tau \, \dot{\expctE}
=
      \timeAver{P} - \timeAver{D}
%       \timeAver{(u_{x})^2} - \timeAver{(u_{xx})^2}
= 0
    \,.
\ee{rpo:EtimAve}
In particular, the \eqva\
and \reqva\ fall onto the diagonal in \reffig{f:drivedrag},
and so do time averages computed on \po s and \rpo s:
\beq
\timeAver{E}_p =
{1\over \period{p}} \int_0^\period{p}d\tau \, E(\tau)
    \,,\qquad
\timeAver{P}_p =
{1\over \period{p}} \int_0^\period{p} d\tau \, P(\tau)
    =
      \timeAver{D}_p
% \timeAver{(u_{x})^2}_p =
% {1\over \period{p}} \int_0^\period{p} d\tau \, \expct{(u_{x})^2}
%     =
%       \timeAver{(u_{xx})^2}_p
    \,.
\label{poE}
\eeq
In the Fourier basis \refeq{EFourier} the conservation of energy on average
takes form
\beq
0 = \sum_{k=1}^{+\infty} ( (k/\tildeL)^2 - ( k/\tildeL)^4 )\,
    \timeAver{E_k}
\,,\qquad
E_k(t) =  |a_k(t)|^2
\,.
\ee{EFourier1}
The large $k$ convergence of this series is insensitive to the
system size $L$; $\timeAver{E_k}$ have to decrease much faster than
$(k/\tildeL)^{-4}$.
\PC{determine the decay rate, presumably exponential in $k$}
Deviation of $E_k$ from this bound for small $k$ determines the active modes.
This may be useful to bound the number of equilibria, with
the upper bound given by zeros of a small number
of long wavelength modes.
For \eqva\ the $L$-independent bound
    on $E$ is given by Michaelson\rf{Mks86}. 
The best current bound\rf{GiacoOtto05,bronski-2005} on the long-time limit
of $E$
as a function of the system size $L$ scales as 
$E \propto L^{3/2}$.
%
\PC{
 % Past: Michael Loss has not taught us how to bound $E$ by
 %  Sobolev bounds. Neither has Spiegel. Next: But 
  Constantin says that the answer is in
    \refrefs{temam85ks}. Eckmann says: the  best bound is by Otto\rf{GiacoOtto05}; the
    only bound close to k=0, better in essential way. See also \refref{bronski-2005}.
    Eckmann had $L^{8/5}$, but conceptually Otto is the best. Recheck whether it is
    $|u|$ or $E \propto L^{3/2}$.
    When the solution is big, how long can it stay big? They found it cannot stay big for
    long. 
       }
\PC{
from bounds on energy, we might be able
to bound the number of equilibria as function of systems size $L$, and thus
be sure we have them all.
   }
%
\PC{Next for fluid guys: read Lieb and Ruelle to learn
    how to bound $E$ for plane Couette by Sobolev bounds}
