% KSe.tex
% $Author$ $Date$

% Predrag extracted from newton.tex         jul  3 2006
% Predrag                   jun 20 2006
% Vaggelis                  may 20 2006

\section{\KSe}
\label{s-KS}
% Predrag                    4jul2006
% extracted from ~dasbuch/book/chapter/PDEs.tex  5jun2005 version
% Predrag                   17sep1999

% \PC{incorporate missing refs from chapter/refsPDEs.tex}

The \KS\ [KS] system\rf{ku,siv}, which
arises in the description of stability of
flame fronts, reaction-diffusion systems and many other physical settings,
is one of the simplest nonlinear PDEs that
exhibit spatiotemporally chaotic behaviour.
In the formulation adopted here, the time evolution of the 
``flame front velocity'' $u=u(x,t)$ defined on a periodic domain
$u(x,t) = u(x+L,t)$
is given by
\beq
    u_t=(u^2)_x-u_{xx}- u_{xxxx}
    \,,\qquad   x \in [0,2\pi \tildeL]
    \,.
\ee{ks}
Here $t \geq 0$ is the time and
$x$ is the spatial coordinate.
The subscripts $x$ and $t$ denote partial derivatives with respect to
$x$ and $t$. We shall use interchangeably the ``dimensionless''
system size $\tildeL$, or the periodic domain size $L= 2\pi \tildeL$,
as the single system parameter.


\subsection{Symmetries of \KSe}
% former \subsection{Scaling and symmetries}
% former symm.tex
% Predrag extracted from newton.tex         jul  3 2006

The  \KSe\ is
Galilean invariant: if $u(x,t)$ is a solution, then 
$v+u(x+2vt,t)$, with $v$ an arbitrary constant velocity, is also a solution. 
Without loss of generality, in our calculations we shall set 
the mean velocity of the  front to zero,
% \index{Galilean invariance}
% \index{invariance!Galilean}
\beq
\int dx \, u = 0
\,.
\ee{GalInv}

% In terms of the system size $ 2\pi L$, the only length scale available,
% the dimensions of terms in \refeq{ks} are
% $ %\[
% [x]=L
% $, $%\,,\quad
% [t]=L^2
% $, $%\,,\quad
% [u]=L^{-1}
% $, $%\,,\quad
% [\nu]=L^2
% \,.
% $ %\]
% What is the non-dimensional ``Rayleigh'' number for the
% \KS\ system? 
%  Scaling out the ``viscosity'' $\nu$ 
% \[ 
% x \to x \nu^{\frac{1}{2}}
% \,,\quad
% t \to t \nu
% \,,\quad
% u \to u \nu^{-\frac{1}{2}}
% \,,
% \]
% brings the \KSe\ \refeq{ks}
% to a non-dimensional form
% \beq
% u_t=(u^2)_x-u_{xx}- u_{xxxx}
% \,,\qquad 
%   x \in  [0,L\nu^{-\frac{1}{2}}] = [0,2\pi\tilde{L}]
% \,.
% \ee{ks-L}
% In this way the ``viscosity'' $\nu$
% and the system size $L$ are trade in for a single
% dimensionless system size parameter
% \beq
%   \tilde{L}={L}/{(2 \pi \sqrt{\nu})}
% % \tilde{L}=\frac{L}{2 \pi \sqrt{\nu}}
% % \,,
% \ee{tildeL}
% which plays the role of a ``Reynolds number''
% for the \KS\ system.

% In the literature sometimes 
% $L$ is used as the system parameter, with $\nu$ fixed to $1$, and
% at other times $\nu$ is varied with $L$ fixed to either $1$ or $2\pi$.
% To minimize confusion,
% in what follows we shall state results of all 
% calculations in units of dimensionless system size $\tilde{L}$.
% \PC{motivate $2\pi$ factor by the mean wavelength,
%     refer to the equation number}
% Note that the time units also have to be
% rescaled; if $T^*_p$ is a period
% of a periodic solution of \refeq{ks} with a given
% $\nu$ and $L=2\pi$, then
% the corresponding solution of the non-dimensionalized \refeq{ks-L}
% has period 
% \beq
%  T_p= T^*_p/\nu
% \ee{Trescl}
%
The KS equation  \refeq{ks} is time translationally invariant,
and 
space translationally invariant
under the 1-$d$ Lie group of $O(1)$ rotations: if
$u(x,t)$ is a solution, then $u(x+d,t)$ is an equivalent
solution for any $-L/2 < d \leq L/2$.
As a result,
KS can have \rpo\ solutions with nonzero shift
\beq
u(x+d,\period{}) = u(x,0)
\,.
\ee{KSrpos}
where $\period{}$ is the period.
\ES{Is this obvious to everybody? I cannot really see why it is so.}
iiiiiii
The KS equation is also invariant under
reflection (``parity'' or ``inversion'') operation
\beq
\Refl u(x) = -u(-x)
\ee{KSparity}
and the shift symmetry operation 
\beq
\Shift u(x)=u(x+L/2)
\,. 
\ee{KSshift}
In the Fourier modes decomposition  this
symmetry corresponds to invariance under
\refeq{FModInvSymm}.
This shift symmetry is a particular case of the
above translational $O(1)$ invariance; more generally,
KS allows for solutions invariant under any rational shift by
$L/m$. Any \rpo\ with such rational shift is eventually periodic.
\Rpo s without a discrete symmetry have irrational shifts
$d$, and are not segments of a periodic solution.


Relations $\Refl^2=\Shift^2=1$
induce a linear decomposition of the space of solutions into 4 invariant
subspaces\rf{KNSks90} via projection operators
$\Refl_1=(\matId+\Refl)/2$,
$\Refl_2=(\matId-\Refl)/2$,
$\Shift_1=(\matId+\Shift)/2$, and
$\Shift_2=(\matId-\Shift)/2$. The nonlinear term $(u^2)_x$ in the KS equation
mixes these subspaces, leading,
according to \refref{KNSks90}, to four invariant subspaces
(labeled by the corresponding projection operators):
\begin{itemize}
 \item[$R_1$:] The space of antisymmetric functions,
 \item[$S_1$:] The space of L/2 periodic functions,
 \item[$R_1 S_1$:] The intersection of the above two,
 \item[$L$:] The space of functions invariant under $x\mapsto L/2-x,\ u\mapsto -u$.
 
\end{itemize}
%\ES{Using notation of \rf{KNSks90} here, not sure if it exists elsewere.}
The dynamics within the $R_1$ space of antisymmetric functions
was studied in \refref{Christiansen:97,Lan:Thesis,lanCvit06}.

