% reducesymm/QFT/correspnd.tex
% $Author$ $Date$
% Predrag  switched to github.com               jul  8 2013

\section{Is QED finite? Correspondence}
\label{sect:correspnd}


\begin{bartlett}{
I am not sure I'll will tell you anything useful any time soon, as I'm a
lapsed field theorist, but here is for your amusement my
\HREF{https://vimeo.com/79605418} {recent rant}. To
cite my wife: ``You may well be right".
        }
\bauthor{
\HREF{http://www.goodreads.com/work/quotes/3226250-keep-the-aspidistra-flying}
{Keep the Aspidistra Flying!}
    }
\end{bartlett}
\bigskip



\begin{description}

\item[2007-03-18 Dirk Kreimer] % kreimer@ihes.fr via nbi.dk
I was pretty much looking forward to grab you for a chat on Dyson
Schwinger etc whilst in Bonn this starting week. Now I fail to be there
myself due to a stupid mechanical problem with a vertebrae.
Pondering about other options, let me just point out that the IHES
(www.ihes.fr) has an excellent visitor program. If you foresee some time
you fancy to pay us a visit, please go ahead.

I am running a little seminar at Boston whilst being there, so besides
what works out with IHES,  whenever you are available in the above
period, you are herewith invited for a talk at BU.
Would you be able to come for a seminar to Boston, Thu, Nov 29 this year?

\item[2011-08-23 Predrag] to Dirk:

I feel like some stuck-up god knows who for not having answered sooner -
but it is not that. I have been focusing on turbulence (with the goal of
using what I learn on Yang-Mills, eventually - if it works, Feynman
diagrammar will be of secondary importance, the first step is fully and
totally non-perturbative) so I feel like fraud talking about QFT when I
am suspended between the perturbative past and the elusive
non-perturbative future.

If someone wants to hear about turbulence as in
\HREF{http://ChaosBook.org/tutorials} {ChaosBook.org/tutorials}, I can do
that honestly - otherwise we wait until I actually do any QFT worth
hearing about...

I got \HREF{http://birdtracks.eu} {birdtracks.eu} published, finally -
you get 100 birdtracks per 1\$, a real deal if you are into that kind of
thing.

Guilty as Charged -
Predrag

\item[2011-08-26 Dirk]
June 01-05 2009 I am organizing a workshop at IHES. It has a
non-perturbative component via Dyson--Schwinger eqs. Would you be
interested to participate and come?

\item[2017-06-16 Dirk]
I will show your notes to
Henry Kissler\rf{Kissler16}   <kissler@physik.hu-berlin.de>
and
\HREF{http://people.physik.hu-berlin.de/~borinsky/}
{Michael Borinsky}\rf{Borinsky14,Borinsky17}    <borinsky@physik.hu-berlin.de> .

There should be a few people at Les Houches June 2018 who have something
to say about asymptotics of solutions of Dyson-Schwinger equation for
example.


\end{description}

\input{WDSmith} % 2013-10-23

\input{Laporta17figuragau}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{rrrrr}
\hline
   1  & (1,3,0) & - 1.971075616835818943645699655337264406980    & - 1/2  \\%01
   2  & (2,2,0) &  - 0.142487379799872157235945291684857370994  & \phantom{+} 1/2 &~(!)\\%02
   3  & (1,2,1) &  - 0.621921063535072522104091223479317643540  & - 1/2  \\%03
   4  & (2,1,1) &  \phantom{+} 1.086698394475818687601961404690600972373  & \phantom{+} 1/2  \\%04
   5  & (3,1,0) &  - 1.040542410012582012539438620994249955094   & - 1/2  \\%05
   6  & (4,0,0) &  \phantom{+} 0.512462047967986870479954030909194465565  & \phantom{+} 1/2  \\%06
\hline
   7  &  & \phantom{+} 0.690448347591261501528101600354802517732  \\%07;
   8  &  & - 0.056336090170533315910959439910250595939  \\%08;
   9  &  & \phantom{+} 0.409217028479188586590553833614638435425  \\%09;
   10 &  & \phantom{+} 0.374357934811899949081953855414943578759  \\%11;
   11 &  & - 0.091305840068696773426479566945788826481  \\%131;
   12 &  & \phantom{+} 0.017853686549808578110691748056565649168  \\%141;
   13 &  & - 0.034179376078562729210191880996726218580  \\%132
   14 &  & \phantom{+} 0.006504148381814640990365761897425802288  \\%142
   15 &  & - 0.572471862194781916152750849945181037311  \\%15;
   16 &  & \phantom{+} 0.151989599685819639625280516106513042070  \\%16;
   17 &  & \phantom{+} 0.000876865858889990697913748939713726165  \\%21;
   18 &  & \phantom{+} 0.015325282902013380844497471345160318673  \\%22;
   19 &  & \phantom{+} 0.011130913987517388830956500920570148123  \\%23
   20 &  & \phantom{+} 0.049513202559526235110472234651204851710  \\%24
   21 &  & - 1.138822876459974505563154431181111707424  \\%12;
   22 &  & \phantom{+} 0.598842072031421820464649513201747727836  \\%19;
   23 &  & \phantom{+} 0.822284485811034346719894048799598422606  \\%20;
   24 &  & - 0.872657392077131517978401982381415610384  \\%17;
   25 &  & - 0.117949868787420797062780493486346339829  \\%18;
%  1+2+{\ldots}+6 & - 2.176866027739540077443259355895893938670  \\
% -2*1.9710-2*0.14249-2*0.6219+1.0867-2*1.0405+0.5125 =
% -1.9710-0.14249-0.6219+1.0867-1.0405+0.5125 =  -2.17669
\hline
\end{tabular}
\end{center}
\caption{
Contribution to $A^{(8)}_1$ of the 25 gauge-invariant sets of
\reffig{Laporta17figuragau}, as reported by Laporta\rf{Laporta17}.
}
\label{Laporta17:tableset}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{table}
%\small
%\begin{center}
%\begin{tabular}{ccrrr}
%gauge set &$(k,m,m')$& value~~ & prediction \\
%\hline
%   (1)  & (1,3,0) & - 1.9710    & - 1/2  \\%01
%   (2)  & (2,2,0) &  - 0.1424   & \phantom{+} 1/2 &~(!)\\%02
%   (3)  & (1,2,1) &  - 0.6219   & - 1/2  \\%03
%   (4)  & (2,1,1) &  \phantom{+} 1.0867  & \phantom{+} 1/2  \\%04
%   (5)  & (3,1,0) &  - 1.0405   & - 1/2  \\%05
%   (6)  & (4,0,0) &  \phantom{+} 0.5125  & \phantom{+} 1/2  \\%06
%\hline
%%  1+2+{\ldots}+6 & - 2.176866027739540077443259355895893938670  \\
%% -2*1.9710-2*0.14249-2*0.6219+1.0867-2*1.0405+0.5125 =
%% -1.9710-0.14249-0.6219+1.0867-1.0405+0.5125 =  -2.17669
%\end{tabular}
%\end{center}
%\caption{
%Contribution to $A^{(8)}_1$ of the 6 gauge sets of
%\reffig{Laporta17figuragauShort}, as reported by Laporta\rf{Laporta17}
%(for the full 25 gauge-invariant sets, see \reftab{Laporta17:tableset}).
%The last column: 1977 Cvitanovi\'c predictions\rf{Cvit77b}
%for the six no-electron loops gauge sets.
%Signs are right, except for the set (2) = $(2,2,0)$, which is anomalously small,
%and the remaining sets are surprisingly close to multiples of 1/2.
%There might be factors of 2 having to do with symmetries, missing from the
%guesses of \refref{Cvit77b}, but I cannot see how that would work.
%Only (4) = $(2,1,1)$ and  (6) = $(4,0,0)$ are symmetric,
%but (1) = $(1,3,0)$, (4) and (5) = $(3,1,0)$ seem to
%have an extra factor of 2 or 4.
%}
%\label{Laporta17:tablesetShort}
%\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\newpage

\begin{description}

\item[2017-04-30 Predrag] to
Stefano Laporta <stefano.laporta@bo.infn.it>:

Thanks for the listings of gauge sets, and of course, for your whole amazing
project. While in detail my 1977 guesses are wrong, the overall finiteness conjecture
still looks good - it's crazy how small all these individual contributions are.

The reason I got into a collaboration with Kinoshita (nominally, Tung-Mow
Yan was my adviser) is that he gave me the 3-loop ladder diagram (at
4-loops, it would belong to your gauge set (6) = $(4,0,0)$) to evaluate,
and I derived a compact formula for the parametric integrand of $n$-loops
ladder diagram. So maybe gauge sets of type (6) = $(4,0,0)$ are the
easiest to evaluate?

Is evaluating gauge set $(5,0,0)$ with 5 loops feasible? If that turns out to be
+1/2, that would be sweet.

\item[2017-05-05 Laporta] to Predrag

To my knowledge, there are no ideas how to bound the sizes of the gauge sets.

In the set (6) = $(4,0,0)$ the ladder is easy to evaluate, but the others are not.
The evaluation of the set $(5,0,0)$ with five loops would be a real challenge.

Among the sets (1)-(6), the easiest to evaluate is the set (3) = $(1,2,1)$.

\item[2017-06-17 Laporta] on \emph{transcedentals}:
There are no cancellations of transcendentals at 2 loops.

At 3 loops there is only one:\\
$\zeta(2)\log(2)$ cancels in the sum of  the gauge sets
(1,2,0)+(2,1,0)+(1,1,1)

At 4 loop there are several:\\
$\zeta(2)^2\log(2)$ cancels in the sum of the gauge sets 8,9,10
(using my numbering of gauge sets) \\
$\zeta(2)^2$        cancels in the sum
of the gauge sets 5,6,20 (or 10,13,19) \\
a weight-7 elliptic constant
cancels in the sum of the gauge sets 2,3,4;
(I have not checked all possible combinations of gauge sets)


\item[2017-05-25 Predrag] to Stefano

I've now searched through the literature. There seem to be 2 approaches
that might be relevant to establishing bounds on, and actually
computation of gauge-invariant sets (I am ignoring various $N=2$ and
$N=4$ supersymmetric models), Hopf algebraic approach of Kreimer and
collaborators, and the worldline formalism pursued by Schubert and
collaborators. It operates with $N$-photon propagators, the ingredient
that defines the quenched gauge-invariant sets for $(g-2)$. The summary
of this literature is in \refsect{sect:future}.

\item[2017-05-21 Predrag] to
Sergey  A. Volkov <volkoff\_sergey@mail.ru>
%\\
%Skobeltsyn Institute of Nuclear Physics,
%\\
%Moscow State University, Moscow, 119234 Russia
%
%Dear Sergey

I have read your
{\em New method of computing the contributions of graphs without lepton
loops to the electron anomalous magnetic moment in {QED}}\rf{Volkov17}
(and the earlier \refref{Volkov16}) with great interest, and I can see
that you are set to compute the 5-loop correction to the electron
$(g-2)$. This is a very hard calculation, and approaching it
strategically is a necessity.

May I suggests that you order your calculation by gauge sets of
\reffig{tabGaugeSets}, as illustrated (for the 4-loop case) in Stefano
Laporta's \reffig{Laporta17figuragauShort}? My hunch is that the gauge
set (6) = $(4,0,0)$ would be the most interesting, though Stefano thinks
it too hard, and suggests starting with a 5-loop relative $(1,3,1)$ (or
$(1,2,2)$?) of the set (3) = $(1,2,1)$ instead. While the contributions
of individual vertex graphs (and self-energy sets\rf{AoHaKiNi15}) are all
over the place, all gauge-invariant sets are insanely small up to order
8, and it would be very sweet to see that this continues through order 10
(at least for the 5-loop graphs with no electron loops).

By the way, to check my conjecture one needs the gauge sets only to two
significant digits or so, no high accuracy is needed.

\item[2017-06-18 Sergey]
\textbf{About gauge sets}:
It is possible to calculate the contributions of $(k,m,m')$ by my
method. It will require more computer resources than the calculation of
the total contribution without splitting into sets, because the sets
$(k,m,m')$ cut the families that I use for calculation into many pieces.
I will try to calculate 5-loop contributions of $(k,m,m')$ after
some refinement of my code (this requires some months). Even a not
precise calculation needs a lot of computer time.

\item[2017-06-18 Sergey] About
\textbf{``That QED on-mass shell amplitudes are IR-free must be an old result..."}:
Is there a rigorous mathematical proof that they are IR-free? I didn't
find such a proof. I saw only some approximate considerations about this.
The proof of the Bogoliubov-Parasiuk theorem is quite large and
complicated. I think that the total proof of IR and UV cancellation
should be larger. However, all considerations that I saw are much
simpler. I think that an existence of such a proof should affect
computational methods. However, at the present moment, the best
calculation of the QED contributions to g-2 (Aoyama, Kinoshita et al.)
uses the subtraction procedure that needs a handwork at 5-loop level (see
the 2008 Aoyama \etal\ article\rf{AoHaKiNiWa08}, and the link to it in
the end of their 2012 article\rf{AoHaKiNi12a}). On the other hand, all
mathematical proofs of IR cancellations that I saw are connected with IR
divergences of other nature (not on-shell). Is there a mathematical proof
that the QED contributions to $g-2$ are finite at any order of the
perturbation series (only for each term, not for the sum)?

\item[2017-06-18 Predrag to Sergey]
I thought that my thesis argument on IR divergences subtraction in the
Feynman-parametric space in \rf{CviKin74b} establishes IR finitness for
(g-2).
We show that the UV and IR finite part of amplitude $M_G$ is given by
\beq
\Delta M_G =\prod_{ij}(1-I_{G/S_i})(1-K_{G/S_j}) M_G
\ee{UV-IRfiniteGcorr}
where the products are over all self-energy and vertex subdiagrams $S_i$ and
$S_j$. We then use Ward identities to show that UV and IR divergent parts
of counterterms $I_{G/S_i} M_G$, $K_{G/S_j} M_G$, \etc\ cancel on mass
shell (and the finite parts contribute to the anomaly). In the paper I
check various cases, and I seem to have missed some that do not arise in
the 3-loop order. But how could anything be simpler than
\refeq{UV-IRfiniteGcorr}? And how could such elegant formula be wrong?

I think the UV/IR finitness is very clean and clear to see in the
dimensional regularization.
I have now reread my 1976 {\em {Yang-Mills} theory of the
mass-shell}\rf{MassShell}, and it is a rather amusing, but probably not a
persuasive answer to your query. The paper is good and original, so it is
never cited. At least it got published. The original draft was to be
exactly \emph{one} page of Physical Review Letters, {had} \emph{one}
reference, a homage to Schwinger\rf{Schwinger48}. Those were happier
times - I wrote it in a couple of days while visiting Helsinki Research
Institute for Theoretical Physics, mailed it, impressing my hosts
infinitely - they have never seen a PR Lett written at that speed, and
\HREF{http://www.cns.gatech.edu/~predrag/friends/Predrag/MidnightRider75/mrider.html}
{\em bicycled off into Lapland}. But referees made me quadruple the
length of the Letter, making it as boring as any other PR Lett. For the
proof of absence of on-mass shell $(g-2)$ IR divergences I cite no less
than three of my own unpublished papers (!!!), and nobody else.
What
happened? I had understood that (the finiteness of QED
conjecture had much to do with it) that doing more Feynman graphs would
be stupid (expanding around the wrong vacuum), and thus started my
nonlinear detour. I started thinking about functional equations on May 1,
1976, and wrote down
\HREF{http://chaosbook.org/~predrag/papers/universalFunct.html}
{\em the period doubling fixed point function equation} on May 3, 1976.
In my office I do have a very thick folder marked UNPUBLISHED, and it
must contain the drafts of these IR papers. But how would I know that
anyone would care 40 years later?

The one that has IAS number was actually published\rf{IRstruct}, see
\HREF{http://chaosbook.org/~predrag/papers/IRstruct.pdf}
{here}, and it says:
\begin{quote}
``The anomaly, expressed in terms of the unrenornormalized coupling
constant, is free of infra-red divergences for QED (\ie, all $a^{(2n)}$
are IR finite). This is particularly easy to demonstrate using the method
for separation of IR divergences given in \refref{CviKin74b}. The
analysis is rather technical, but the result is very intuitive: a Feynman
integral $M_G$ contributing to the magnetic moment is IR divergent
whenever the corresponding diagram G can be split into a vertex
subdtagram S and a cloud of soft gluons attached to the external quark
lines (diagram G/S). In general some gluons in G/S can be hard, provided
they are within UV divergent subdiagrams, In all cases, the IR divergent
part of $M_G$ factors into $M_S$ times the IR divergent part of
$L_{G/S}$, where $L_{G/S}$ is the charge form factor computed from the
diagram G/S.''
\end{quote}
and so on - the paper is focused on QCD, but I believe it establishes IR
finiteness of QED on-mass shell amplitudes, though it most likely not as
rigorously as you would like it.

Yennie and S. C. Frautschi and H. Suura\rf{YeFrSu61}
{\em The infrared divergence phenomena and high-energy processes}
(cited over 800 times!) and
Kinoshita\rf{Kinoshita62}
{\em Mass singularities of {Feynman} amplitudes}
say nothing about cancellation of IR divergences by renormalization.
Herzog and Ruijl\rf{HerRui17}
{\em The {R*}-operation for {Feynman} graphs with generic numerators}
and Chetyrkin, Tkachov and
Smirnov\rf{CheTka82,CheSmi84,SmiChe85}
are not of interest, as {R*}-operation is not about on-shell IR-divergences.
 (Herzog and Ruijl\rf{HerRui17} cite Chetyrkin, Tkachov and
Smirnov\rf{CheTka82,CheSmi84,SmiChe85} but no Volkov\rf{Volkov16,Volkov17} or
Cvitanvi\'c\rf{CviKin74b,MassShell,IRstruct,QCDmshell,NPB81} papers.)

We also cited in \refref{CviKin74b}
Johnson and Zumino\rf{JohZum59} {\em Gauge dependence of the
wave-function renormalization constant in {Quantum Electrodynamics}}

An unrelated IR problem: Kinoshita\rf{Kinoshita50} {\em Note on the infrared
catastrophe}, about which he writes\rf{Kinoshita16}: ``
I developed a method to handle the infrared problem within the
perturbative framework of Feynman–Dyson theory\rf{Kinoshita50}. My
insight was not deep enough, however, and I overlooked a subtlety in the
infrared problem. This became the source of an error in the calculation
of radiative correction to the $\mu-e$ decay.
''


\item[2017-06-11 Predrag] to
\\
%Professor
Christian Schubert <schubert@ifm.umich.mx>
%\\
%Instituto de Física y Matem{\'a}ticas,\\
%Universidad Michoacana de San Nicol{\'a}s de Hidalgo,\\
%Ciudad Universitaria, C-3, Apdo. 2-82,\\
%C.P. 58040 Morelia, Michoacan, MEXICO
%
%\bigskip
%
%Dear Christian

Stefano Laporta  has recently published analytic values of all 4-loop
electron magnetic moment gauge sets, and they look very intriguing.
Sergey  A. Volkov has started evaluating individual 5-loop vertex
diagrams and is in position to estimate numerically the size of 5-loop
gauge sets. So it might be a good time to make a new attempt to prove/disprove
the QED finiteness conjecture. This set of notes is my current best attempt
to motivate this effort.

I do not take my 1977 ``gauge-set approximation'' very literally - its
content is only that if individual gauge sets can be bounded to anything
growing slower than combinatorially, quenched QED (and hopefully full
QED) is a finite theory, not an asymptotic series. But the form of vertex
gauge sets is very suggestive; it is defined in terms of $N$-photon
exchanges. So to me the wordline formalism seem the most promising way
forward.

While the quenched QED magnetic moment of electron is the cleanest
possible physical calculation one can do, for sociological reasons most
effort on high order estimates has gone in other directions. Do you have
a formula for spinor QED magnetic moment that one could attempt to
analyze more closely?

%As I have returned to this problem only last month, after a 40 year
%hiatus :) I would be grateful for any further pointers to the literature.

\item[2017-06-12 Christian]
I have been
fascinated by the finiteness conjecture ever after Gerald Dunne and I
were led there, too, from Euler-Heisenberg, Borel analysis,
Ritus mass shift and worldline instantons fifteen years ago.
The things have picked up a lot
just during the last year or so, namely:

\begin{enumerate}
  \item
Idrish Huet, Michel Rausch and I are working on the Affleck,
Alvarez and Manton (AAM) exponentiation conjecture\rf{AffAlMa82} in 1+1 QED.
We had a nice parameter integral representation for the three-loop
Euler-Heisenberg for quite a while, but could not get a sufficient number
of weak-field expansion coeffs out of it (so far what we got points to
AAM not holding at three loops - rather the coeffs go asymptotically
BELOW the AAM prediction - but this is very preliminary). Michel was just
visiting, and we now got a really nice algorithm (based on the polynomial
invariants of the dihedral group) for the analytical calculation of the
coeffs from the more difficult (nonplanar) 3-loop EH diagram.
  \item
If AAM really fails, then presumably the AAM worldline instanton needs
refinement at higher loops. My former student Naser Ahmadiniaz (now
postdoc in Korea) has made some progress with this.
  \item
For many years I am planning to apply the worldline formalism to (g-2),
in particular to the important graphs with the
light-by-light subdiagram. For this subdiagram I have, since a long time,
an off-shell representation that is permutation symmetric, manifestly
gauge invariant, without spurious UV poles, and moreover allows one to
trivially integrate out the one low-energy photon leg. What held me back
was that the various worldline representations that existed hitherto for
the electron line were all somewhat cumbersome, and seemed not suitable
for high-order calculations. Precisely this problem we have been working
on here for the last half year, and things have fallen into place really
nicely, we have a Bern-Kosower type master formula for the electron line
dressed with any number of photons, in vacuum and in the presence of a
constant field, and one of our students is already programming it. I am
now definitely trying to assemble a collaboration to attack the QED (g-2),
and I anticipate that your notes will be quite useful for motivating my
collaborators.

\end{enumerate}
Gies \etal\rf{GiSaVa05} have some intriguing numerical results from
worldline Monte Carlo (in section 5). Another thing I find quite
interesting are Ritus\rf{Ritus80} most recent
papers\rf{Ritus06,Ritus13,Ritus15} on the value of the bare fine
structure constant - he confirmed to me in an email from last December
(at age 90) that he definitely does not think that the bare charge is
infinite.

I am appending a summary of our efforts in this line of work\rf{HuTrSc17a}
which I wrote as a contribution to the 5th Winter Workshop on
Non-Perturbative Quantum Field Theory. 22-24 March 2017,
Sophia-Antipolis.

\item[2017-06-18 ] <ahmadiniaz.naser@gmail.com>
(currently a postdoctoral fellow at the Center for Relativistic Laser
Science (CoReLS), Institute for Basic Science (IBS), Gwangju, South
Korea).
\\
My main research
interests are in amplitude calculations in QED, QCD and quantum gravity
from the worldline formalism. Recently, we have been interested in
(off-shell) tree-level amplitudes for QED processes in vacuum as well as
in the presence of classical background fields which will also be applied
to higher order corrections.

\item[2017-06-20 James P. Edwards] <jedwards@ifm.umich.mx>
\\
%Instituto de Física y Matemáticas
%Universidad Michoacana de San Nicolás de Hidalgo
\HREF{www.ifm.umich.mx/ifm/index.php/fisca/academicos/edwards/}
{homepage}
\\
I am currently working with Christian on the worldline approach. In some
of our recent work we have been working on more efficient ways to
determine g-2 based upon a worldline expression for the dressed spinor
propagator (both in vacuum and in a constant background).

\item[2017-06-11 Predrag] to
%\\
% Professor
Dirk Kreimer <kreimer@physik.hu-berlin.de>
%\\
%Humboldt-Universit\"at zu Berlin\\
%Institut f\"r Mathematik\\
%\bigskip
%
%Dear Dirk
%
%\medskip

Stefano Laporta  has recently published analytic values of all 4-loop
electron magnetic moment gauge sets, and they look very intriguing.
It might be a good time to make a new attempt to prove/disprove
the QED finiteness conjecture. This set of notes is my current best attempt
to motivate this effort.

If individual gauge sets can be bounded to anything growing slower than
combinatorially, quenched QED (and hopefully full QED) is a finite
theory, not an asymptotic series. The form of vertex gauge sets is
suggestive of a way forward; it is defined in terms of $N$-photon
exchanges. Hopf algebras might hold the key, but I do not know how to use
these ideas.

I would be grateful for any further pointers to the literature.
Anybody else I should send these notes to? And I'm
finally ready to stand up an be counted - I can fly to Berlin on a short
notice if that helps.

%\bigskip
%
%\noindent
%best regards
%\\
%Predrag
%\newpage

\item[2017-06-14 Dirk Kreimer]
consider coming to
\HREF{http://www2.mathematik.hu-berlin.de/~kreimer/wp-content/uploads/houches2018_1116.html}
{Les Houches — June 4-15, 2018} on structures in local quantum field theory.



\end{description}
