% reducesymm/QFT/Gribov.tex
% Predrag  created              Sep 2 2013
% continues siminos/blog/dailyBlog.tex as of that date

\chapter{Gribov ambiguity}
\label{c-Gribov}


% Clipped from \refref{atlas12}:
% \\
Symmetry reduction in dynamics (including classical field theories such
as the \NSe) closely parallels the reduction of gauge symmetry in
quantum field theories. There, the freedom of choosing moving frames
 is called `gauge freedom' and a
particular prescription for choosing a representative from each gauge
orbit is called `gauge fixing'. Just like the slice hyperplanes
may intersect a group orbit many times, a gauge
fixing submanifold may not intersect a gauge orbit, or it may intersect
it more than once (`Gribov ambiguity')\rf{Gribov77,VaZw12} In this
context a chart is called a `Gribov' or `fundamental modular' region and
its border is called a `Gribov horizon' (a convex manifold in the
space of gauge fields). The Gribov region is compact and bounded by the
Gribov horizon. Within a Gribov region the `Faddeev-Popov operator'
(analogue of the group orbit tangent vector) is strictly positive, while on
the Gribov horizon it has at least one vanishing eigenvalue.

\input{dailyBlogKS}

\section{Daily Gribov blog}
\label{s-DailyGBlog}



\begin{description}
\item[2013-12-07  Predrag]
here are a few things I would like to understand in dynamics language:
\begin{enumerate}
  \item All Gribov copies papers say that electromagnetism $U(1)$ is trivial;
  Landau gauge defines a slice normal to all gauge orbits. Why is it that we find
  \SOn{2}\ not trivial?
  \item All Gribov copies papers focus on the copy that includes the
  origin. Are they not automatically excluding the `non-perturbative' solutions,
  \ie\ neighborhoods of turbulent \rpo s that we care about?
  \item What does `gauge potential' connection $A^\mu$ mean in the dynamical
  systems language? How does it get promoted to a dynamical field?
  \item Landau gauge seems to correspond to template's group tangent vector.
  Does any commonly used gauge correspond to `co-moving frame'?
\end{enumerate}


\item[2013-11-23  Predrag] Moved all matters Gribov from
\texttt{siminos/blog/lit.tex} to here.

\item[2011-07-08 Predrag] from \refref{FrCv11}:                     \toCB
\\
% At this point
It is worth noting that imposing the global and fixed slice
%\refeq{PCsectQ}
is not the only way to separate equivariant dynamics into `group
dynamics' and `shape' dynamics\rf{BeTh04}. In modern mechanics and even
field theory (where elimination of group-directions is called
`gauge-fixing') it is natural to separate the flow {\em locally} into
group dynamics and a transverse, `horizontal'
flow\rf{Smale70I,AbrMars78}, by the `method of
connections'\rf{rowley_reduction_2003}. From our point of view, such
approaches are not useful, as they do not reduce the dynamics to a
lower-dimensional \reducedsp\ $\pS/\Group$.


\item[2012-05-20 Jeff Greensite] has written a book\rf{Greensite11} of
possible interest, \emph{An introduction to the confinement problem}.
I have put a GaTech eBook copy
\HREF{http://ChaosBook.org/library/Greensite11.pdf}{Greensite11.pdf} into
ChaosBook.org/library.

\HREF{http://en.wikipedia.org/wiki/Gribov_ambiguity}{Gribov ambiguity wiki}
(edits by Predrag):

Gauge fixing means choosing a representative from each gauge orbit. The
space of representatives is a submanifold and represents the gauge fixing
condition. Ideally, every gauge orbit will intersect this submanifold
once and only once. This is generally impossible globally, especially for
non-abelian gauge theories, because of topological obstructions and the
best that can be done is make this condition true locally. A gauge fixing
submanifold may not intersect a gauge orbit at all or it may intersect it
more than once. This is called a Gribov\rf{Gribov77} ambiguity.

Gribov ambiguities lead to a nonperturbative failure of the BRST
symmetry, among other things.

A way to resolve the problem of Gribov ambiguity is to restrict the
relevant functional integrals to a single \emph{Gribov region} or {\em
fundamental modular region} whose boundary is called a \emph{Gribov
horizon}.

{\em Gribov copies} play a crucial role in the infrared (IR) regime while
it can be neglected in the perturbative ultraviolet (UV)
regime\rf{Gribov77,Zwanz89,Zwanz93}. The restriction to the Gribov region
(defined in such a way that the Faddeev-Popov operator is strictly
positive) can be achieved by adding a nonlocal term, commonly known as
`horizon term', to the YM action\rf{Zwanz89,Zwanz93,Zwanz92}. This is a
nonlocal term in the 4-dimensional Euclidean space, written as an
integral over the `horizon function.'

Greensite: ``In non-Abelian theories, there are many gauge copies -
Gribov copies - that satisfy the Coulomb gauge condition. The Gribov
region is the space of all Gribov copies with positive Faddeev-Popov
eigenvalues. Configurations of the Gribov horizon have at least one FP
eigenvalue $\lambda =0$. What counts for confinement is the density of
eigenvalues $\rho(\lambda)$ near $\lambda =0$, and the `smoothness' of
these near-zero eigenvalues.

The Gribov horizon is a convex manifold in the space of gauge fields,
both in the continuum and on the lattice. The Gribov region, bounded by
that manifold, is compact.
''

Amusingly, they can find the first 200 eigenstates of the lattice
Faddeev-Popov operator on each time-slice of each lattice configuration
by the Arnoldi algorithm.

See also Heinzl\rf{Heinzl96,HeRuSch08}, as well as
\refrefs{RuSchVo02,MaScha94,vanBaal91,vanBaal97,DellAnZwan91,Cutkosky84,Singer78}.
Review of \refref{VaZw12} promises ``to give a pedagogic review of the
ideas of Gribov and the subsequent construction of the GZ action,
including many other topics related to the Gribov region.''


\item[2012-06-15 Daniel]
Wow! Kinda lost me here... How does this apply exactly? Is the
point to make an analogy between our slices and these Gribov regions?
All this QFT stuff kind of comes out of left field. Discussing
it with Predrag convinced me that it may be interesting to have this here
because there appear to be some strong parallels with what we're doing. I
may need a little massaging to make it more palatable, though.

\item[2012-06-15 Predrag]
Laufer and Orland\rf{LauOrl12} say in
{\em The geometry of {Yang-Mills} orbit space on the lattice}: ``
We find coordinates, the metric tensor, the inverse metric tensor and the
Laplace-Beltrami operator for the orbit space of Hamiltonian SU(2) gauge
theory on a finite, rectangular lattice. This is done using a complete
axial gauge fixing. The Gribov problem can be completely solved, with no
remaining gauge ambiguities.
''

\item[2012-06-15 Evangelos]
This seems very interesting and I have to read it. It will most probably
confuse people, if we are really addressing fluid-dynamicists.

{\bf [2012-06-15 Predrag]} I am trying to reach out to quantum filed
theorists, that's why these facts are made explicit here - otherwise they
think it is just plumbing, nothing to do with Fundamental Physics..

Maybe shorten it? For instance Faddeev-Popov operator is not introduced
here, so we might avoid reference to it altogether. In the last sentence,
is the Gribov region or the Gribov horizon a convex manifold?

\item[2010-09-28 ES: Faddeev-Popov ghosts]                    \toCB
(moved to here from froehlich/blog)
\\
From my random readings, supposedly making up for my inability to attend
colloquia in French: Faddeev in a
\HREF{http://www.scholarpedia.org/article/Faddeev-Popov_ghosts}{scholarpedia
article} discusses the difficulties in Yang-Mills quantization that led
him and Popov to introduce fictitious fields, now known as
\emph{Faddeev-Popov ghosts}. The problem was that of gauge fixing,
essentially of working on a slice. Faddeev says:

`` %\begin{ttfamily}
It was clear that the equivalence principle had to be taken into account.
In the functional integral framework, the equivalence principle implies
that one has to integrate over classes of gauge equivalent fields instead
of integrating over all fields $A_\mu^a$.

The choice of the representatives in the classes of equivalent fields is
realized by means of a gauge condition (gauge fixing), for instance,
\[
    \partial_{\mu} A_{\mu}^{a} = 0 .
\]
This condition defines a plane in the set of all fields, which is
intersected by the gauge orbits defined by
\[
    A_{\mu} = A_{\mu}^{a}t_{a} \to A_{\mu}^{\Omega}
            = \Omega A_{\mu} \Omega^{-1} + \partial_{\mu} \Omega \Omega^{-1} .
\]
In this context, the difference among abelian and non-abelian cases
becomes clear. In the abelian case, we take $\Omega(x) =
\exp{i\Lambda(x)}$ and a gauge orbit is defined by
\[
    A_{\mu} \to A_{\mu} + \partial_{\mu} \Lambda ,
\]
which is just a linear shift. Thus all the abelian orbits intersect the
gauge surface at the same angle.

In the non-abelian case, the gauge orbit equations are non-linear and the
intersection angle depends on the field parameterizing the orbit. It is
clear that this must be taken into account in the functional integral.
'' %\end{ttfamily}

See the wikipedia link to put this in the proper context. As an abelian
case Faddeev lists quantum electrodynamics where the group is $U(1)$ (the
same as in \cLe\ if we think in terms of complex variables). As a
non-abelian example he lists the Standard Model: $U(1)\times SU(2) \times
SU(3)$.

Faddeev relies on the group being connected to write group action in
exponential form, as Stefan does. $\On{2}$ in Kuramoto-Sivashinsky is
non-abelian so I am worried that there might be more work required. Are
all non-connected compact Lie groups non-abelian and vice versa?

The weight factor Faddeev and Popov introduced might be helpful in
trace-formulas for non-abelian groups.

\item[2012-06-15 Evangelos]
Something that still confuses me: Is `Faddeev-Popov operator' really the
gauge orbit tangent vector, or is it the group generator? {\bf
[2012-06-15 Predrag]} I think `group generator' would not have enough
information (why should it be position dependent?), so it should be
something like our group orbit tangent vector, except positivity of
eigenvalues sounds like a projection on a given direction across the
slice.

\item[2013-11-27  Predrag]
Williams\rf{Williams02} {\em {QCD}, gauge-fixing, and the {Gribov}
problem}seems like a nice reference.

 	Privatdozent Dr. habil.
 \HREF{http://www.tpi.uni-jena.de/~axm/} {Axel Maas} is cute - has a blog.
He also finds gauge invariance a \HREF{http://axelmaas.blogspot.com/2013/09/blessing-and-bane-redundancy.html}
{blessing and bane}.
In it, \arXiv{1309.1957}, a paper with one of the worst abstracts
ever written, is explained in gentle terms. Some snippets from the article:

``
The minimal Landau gauge\rf{Maas11se} is obtained by minimizing the Hilbert square norm
\beq
|| A ||^2 = \int | A_\mu^b(x) |^2 d^4x,
\eeq
to some local minimum (in general not an absolute minimum) with respect to local gauge transformations $g(x)$.  These act according to ${^g}A_\mu = g^{-1} A_\mu g + g^{-1} \p_\mu g$.  At a local minimum, the functional $F_A(g) \equiv ||{^g}A||^2$ is stationary and its second variation is positive.  It is well known that these two properties imply respectively that the Landau gauge (transversality) condition is satisfied, $\p \cdot A = 0$, and that the Faddeev-Popov operator is positive {\it i. e.} $(\omega, M(A) \omega) \geq 0$ for all $\omega$.  Here the Faddeev-Popov operator acts according to
\beq
\label{Macts}
M^{ac}(A) \omega^c = -  D_\mu^{ac}(A) \p_\mu \omega^c,
\eeq
where the gauge covariant derivative is defined by $D_\mu^{ac}(A) \omega^c = \p_\mu \omega^a + f^{abc} A_\mu^b \omega^c$, and the coupling constant has been absorbed into $A$.  Configurations $A$ that satisfy these two conditions are said to be in the (first) Gribov region\rf{Gribov77} which we designate by $\Omega$.  It is known\rf{Maas11se} that in general there are more than one local minimum of $F_A(g)$, and we do not specify which local minimum is achieved.  This gauge is realized numerically by minimizing a lattice analog of $F_A(g)$ by some algorithm, and the local minimum achieved is in general algorithm dependent. However, for all commonly employed algorithms this does not yield different expectation values, as they all are equivalent to an averaging over the first Gribov region\rf{Maas11se} {Maas:2013vd}.
''

{Maas:2013vd}
  A.~Maas,
  %``Local and global gauge-fixing,''
  PoS ConfinementX {\bf }, 034 (2012)
  [arXiv:1301.2965 [hep-th]].


Some snippets from
\HREF{http://www.tpi.uni-jena.de/~axm/southampton12.pdf} {Maas (2012)}:

QCD is a gauge theory
with a Lie-algebra-element valued gauge-field
$A_\mu = \Lg_aA_\mu^a$.
Gauge transformations change the gauge fields, but leave physics
invariant: gauge orbits. Set of all gauge-equivalent copies is a gauge
orbit. Gauge fields not unique. Quantities that change along a gauge
orbit are gauge-dependent. Gauge transformations do not change physics.

Gauge freedom: Choice of
internal coordinate system.

Path integral averages with flat weight over all gauge copies.
Gauge-fixing is the introduction of a non-flat weight, such that all
copy- independent quantities remain unchanged.
Two possibilities
\\
(1)
Averaging over all or some
copies - must be normalized
\\
(2)
Single out one copy as representative.

Any gauge fixing yields a residual set of gauge copies: Residual gauge
orbit.

Residual copies are related by a residual gauge symmetry.

Gribov copies if isolated.

An example of a Gribov copy: Instanton
[Maas, EPJC 06].

Gribov-Singer ambiguity.

Landau-gauge hypersurface can be decomposed into (Gribov) regions.

Innermost Gribov region is convex and bounded and pierced at least once
by all gauge orbits: Contains all information.

In perturbation theory: Local gauge condition in Landau gauge
$\partial^\mu A_\mu=0$
Sufficient for perturbation theory. Insufficient beyond perturbation theory.
There are gauge-equivalent configurations which obey the same
local gauge-condition: Gribov copies. There are no known local gauge conditions, which lead to a
unique gauge configuration.

\item[2012-06-14 Predrag]
\HREF{http://marcofrasca.wordpress.com/about/}{Frasca} is either insane
or just yet another ignorant field theorist: ``I have worked on almost
all fields of physics'' (???). I checked the publication list, and it is
no  L.D. Landau. But the blog is informative:

{\bf [2012-06-05 Frasca]} (edited by Predrag):

``The answer to the question of the mass gap in Yang-Mills theory has
made enormous progress mostly by the use of lattice computations and,
quite recently, with the support of theoretical analysis. Contrarily to
common wisdom, the most fruitful attack to this problem is using Green
functions. The reason why this was not a greatly appreciated approach
relies on the fact that Green functions are gauge dependent.
Nevertheless, they contain physical information that is gauge independent
and this is exactly what we are looking for: The mass gap.''

(Predrag: this I interpret in the spirit of Gutzwiller - the Green function
is coordinate dependent, but it's trace - which yield the spectrum - is
coordinate invariant.)

``\HREF{http://marcofrasca.wordpress.com/2011/01/28/the-saga-of-landau-gauge-progators-a-short-history/}
{The Saga of Landau-Gauge Propagators: A Short History}'' is a good read:

``We cannot forever ignore the low energy behavior of QCD as its complete
understanding could have impact at unexpected large scales.''


\item[2013-12-07  Predrag]
About the
\HREF{http://marcofrasca.wordpress.com/2011/02/03/the-gribov-obsession/}
{Gribov obsession/}: Frasca says ``Gribov copies should be renamed Gribov
obsession as I did. If you want a fine description of the problem you can
read appendix H in the paper by Alfred Actor\rf{Actor79}, or the
beautiful lectures by Silvio Sorella and Rodrigo Sobreiro\rf{SobSor05a}
{\em Introduction to the {Gribov} ambiguities} are pedagogical.
See also
\refref{SobSor05}.

``Gribov himself proposed a solution limiting solutions to the so called
first Gribov horizon as Gribov pointed out that the set of gauge orbits
can be subdivided in regions with the first one having the Fadeed-Popov
determinant with all positive eigenvalues and the next ones with
eigenvalues becoming zero and then going to negative. ''

Residual freedom: impose Landau gauge condition. It reduces configuration
space to a hypersurface. Leaves the non-perturbative gauge freedom to
choose between gauge-equivalent Gribov copies: Residual gauge orbit.

There is no possibility using a local condition to restrict to one
representative of the residual gauge orbit. So, leave the global color
symmetry unfixed.

Possibilities to complete the Landau gauge:
\\
1) Take the representative inside the fundamental
modular region
\\
2) Average over (part of) the residual gauge orbit
\\
3) Average over the complete residual orbit: Hirschfeld-like
\\
4)
Average over the elements in the first Gribov region:
Minimal Landau gauge.

A local minimum of $\int d^dx A_\mu^a A^\mu_a$
defines the first Gribov region.
The Gribov region $\Omega$ is  defined as the set of
field configurations fulfilling the Landau gauge condition and for
which the Faddeev-Popov operator,
\begin{eqnarray}
\mathcal{M}^{ab} &=&  -\partial_{\mu}\left( \p_{\mu} \delta^{ab}
    + g f^{acb} A^c_{\mu} \right) \;,
\end{eqnarray}
is strictly positive, namely
\begin{eqnarray}
\Omega &\equiv &\{ A^a_{\mu}, \; \partial_{\mu} A^a_{\mu}=0,
    \; \mathcal{M}^{ab}  >0\} \;.
\end{eqnarray}
The boundary, $\partial \Omega$, of the region $\Omega$ is called the
(first) Gribov horizon. However, this region $\Omega$ still
contains a number of Gribov copies and is therefore still ``larger'' than
the fundamental modular region (FMR), which is completely free of Gribov
copies. Unfortunately, it is unknown how to treat the FMR analytically
~~{Semenov,Dell'Antonio:1991xt,Dell'Antonio2,vanBaal:1991zw}.

Nele Vandersickel PhD thesis\rf{Vandersickel11} {\em A study of the
Gribov-Zwanziger action: from propagators to glueballs} might be
pedagogical. Her
\HREF{http://physik.uni-graz.at/~dk-user/talks/Vandersickel20100225.pdf}
{talk} gives a compact overview, might be useful for writing this up.
Chapter 3 of her thesis\rf{Vandersickel11}, \arXiv{1104.1315}, gives a
pedagogic overview of the Gribov-Zwanziger framework, not available yet
in the literature.

This 2010 \HREF{https://sites.google.com/site/facingqcd/} {workshop}
gives a pretty good idea of who is who in the field. I am curious what
Nele is doing today - \HREF{https://biblio.ugent.be/person/802000264158}
{her papers} stop in 2012.

Dudal~\etal\rf{DGSVV08} is a very impressive, huge paper.


\item[2016-02-17  Predrag]
van den Heuvel and van Baal\rf{vdHvB95} in {\em Dynamics on the {SU(2)}
fundamental domain} mean \slice\ when the say ``fundamental domain''.
They use our `nearest point on the orbit' slice condition.

``
Let ${\cal A}$ be the set of all gauge fields ${\cal A}  : S^3 \to
su(2)$. The \emph{physical configuration space} is the space of gauge
orbits ${\cal A}/\Group$. Let $\Group$ be the set of all gauge
transformations $g : S^3 \to SU(2)$. We would like to have a
\emph{fundamental domain}, that is, a set of gauge fields which is in
one-to-one correspondence with the physical configuration space.
''

``
As a representative of an orbit, we choose the gauge field that has
lowest norm.'' \ie, the same \slice\ condition as what we do.

This work is continued by Cucchieri\rf{Cucchieri98} {\em Numerical study
of the fundamental modular region in the minimal {Landau} gauge}, who
studies the so-called \emph{fundamental modular region}, a
region free of Gribov copies, in the minimal Landau gauge for pure $SU(2)$
lattice gauge theory.



\end{description}
\renewcommand{\ssp}{a}

%\newpage %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\printbibliography[heading=subbibintoc,title={References}]
