\ifsvnmulti
 \svnkwsave{$RepoFile: lyapunov/students.tex $}
 \svnidlong {$HeadURL$}
 {$LastChangedDate$}
 {$LastChangedRevision$} {$LastChangedBy$}
 \svnid{$Id$}
\fi

\chapter{Log of student's projects}
\label{c-students}

\section{Xiong Ding's project}
\label{sect:introXD}

The goal of the summer project is that Xiong Ding adopts the Ginelli
\etal\rf{GiChLiPo12} method for computation of covariant vectors to
computation of \po\ Floquet vectors,  and implements it for the \po s and
\rpo s in our \KS\ database.


\begin{description}

\item[2013-06-27 Predrag] Our first goal is to learn
how to compute \emph{all} eigenvectors (CLVs) and Floquet multipliers
(in literature mistakenly often called `Lyapunov exponents') of
\JacobianMs\ for \po s computed in \refref{SCD07}.
Please study this reference in depth. For \KS\
calculations of \refchap{sect:LyapKS}, system size $L=22$, this is 62
eigenvectors. We have at least 40,000 \rpo s.

You will first check whether you agree with Kazumasa's results for his
favorite \po\ \PO{10.3} with $\period{p} = 10.253$, then the new work
starts. \PO{10.3} is the prime orbit period of a pre-periodic orbit,
the full period is twice that, see Fig. 6.1\,(i) \emph{Selected
relative periodic and pre-periodic orbits} in \refref{SCD07}.

Save your code in a \texttt{siminos/xiong/} subfolder. You want to
confirm the result in \refref{SCD07}, see
\HREF{http://www.cns.gatech.edu/~predrag/papers/SCD07.pdf}{sect. 4} in
that paper:
    \PC{the source file is in \texttt{siminos/rpo\_ks/}}

\begin{quote}
``[...] Indeed, numerically the covariant Lyapunov
vectors\rf{ginelli-2007-99} of the $L=22$ chaotic attractor separate
into 8 ``physical'' vectors with small Lyapunov exponents
$(\Lyap_j) = (0.048,$ 0, 0, $-0.003$, $-0.189$, $-0.256$,
$-0.290$, $-0.310$),
and the remaining 54 ``hyperbolically isolated'' vectors with rapidly
decreasing exponents
$(\Lyap_j)
= (-1.963$,   $-1.967$,   $-5.605$,   $-5.605$,  $-11.923$,  $-11.923$,
 $\cdots) \approx -(j/\tildeL)^4$,
in full agreement with the Yang \etal\rf{YaTaGiChRa08} investigations
of KS for large systems sizes.
 [...]''
\end{quote}

\item[2013-06-27 Predrag]
This section is an evolving draft of Xiong Ding's project report;
daily progress is reported in \refsect{sect:dailyBlXD}.


Please {\color{red} write down here}, in your project's
log, the equations you are integrating (you can clip and paste
equations from elsewhere in this repository to save time).
This will be a continuously evolving draft of the project report.
When you
write up the project report it will all be needed, so you might just
as well do it now.

If we know where are you in the project, we can help you.
Write up full fledged
derivations and explanations of what you are learning - clear enough
that some of that text can be later used in your project report, or a
publication, if we get that far. Clear exposition is 1/3 of research.



My vision of the project is sketched in \refchap{c-draft}. Studying
\refchap{s:LyapunovVec} is also a good idea. \refChap{sect:LyapKS}
deals specifically with the \KS\ calculations.
Search for\\
``2012-02-06 Evangelos Talked with Hugues at MPIPKS, created a to-do list:''
\\
to see Evangelos and Hugues work plan.

\item[2013-06-27 Predrag]
Setting up Xiong Ding for the svn (subversion) repository access

\texttt{svn://zero.physics.gatech.edu/siminos}

you will
need this userID, password, so save it in your secret stash of magic
words:
\\
user: xiong  password: Lyapunov

to access repositories read about subversion/svn and VPN access on
\HREF{http://cns.gatech.edu/CNS-only} {cns.gatech.edu/CNS-only} - to
access any of the internal CNS pages you will need

cnsuser           cnsweb

\HREF{http://ChaosBook.org/library}{Click here} for internal reprints and
books collection. You will need

student           Lautrup

I do not have the patience to enter every paper in the listing on this
internal home page, so many papers are there but not listed. Their PDF
file name is the same as their bibTeX name for the publication. For
example, a book by Wassim M. Haddad\rf{HaCh08} has BibTeX name
\texttt{HaCh08}, so I put a copy of their book into the
ChaosBook.org/library as

\HREF{http://ChaosBook.org/library/HaCh08.pdf}
{ChaosBook.org/library/HaCh08.pdf}.

If you download some publications that would be nice to save for other
people, let me know, and I'll show you how to do it.

To compile the blog,
\begin{verbatim}
 cd siminos/lyapunov/
 pdflatex blog; bibtex blog; pdflatex blog
\end{verbatim}

To help a bit with starting writing, I will
edit your text (you can see edits by using diff in Tortoise).

Here is how you can
\Xiongedit{highlight your edits}, and here is how you can add a dated
footnote\Xiong{2013-03-03}{This is a footnote. Always enter the date, so
we know how old it is.}.

In the first draft, do not worry about
English, just focus on the formulas. Write
down what a starting graduate student would need to know to be able
to set up the calculation.

If you want source files for ChaosBook.org to clip and past from, you can
\begin{verbatim}
  svn checkout svn:\\zero.physics.gatech.edu\dasbuch
\end{verbatim}

Always trim the figures, avoid lots of white
space around the edges.

If you look into directory
siminos/chaos/ and pdflatex blog.tex, you can see an example of what
a student work log looks like (after a semester of so of work).
For examples of
what final reports look like, see \wwwcb{/projects}.


Come to me or Skype me if there is something you
cannot figure out.

\end{description}


\subsection{Covariant vectors}

\begin{description}

\item[2013-06-27 Predrag]
I rewrote ChaosBook.org chapters
\HREF{http://chaosbook.org/paper.shtml\#stability} {``Linear stability''},
\HREF{http://chaosbook.org/paper.shtml\#invariants} {``Cycle
stability''}, and
\HREF{http://chaosbook.org/paper.shtml\#Lyapunov} {``Lyapunov
exponents''}, to help graduate students. Any suggestion for improving
these chapters is greatly appreciated.

                              \inCB
{Finite-time} {Lyapunov} or \emph{characteristic} exponents and
associated \emph{principal axes} in the theory of dynamical systems
are discussed in chapter
\HREF{http://chaosbook.org/paper.shtml\#Lyapunov} {``Lyapunov
exponents''}. Oseledec \emph{Lyapunov exponents}\rf{lyaos} are the
$\zeit\to\infty$ limit of these. \emph{Floquet multipliers} and
\emph{eigen-vectors} are property of finite-time, compact invariant
solutions, such as \po s and \rpo s; they are explained in chapter
\HREF{http://chaosbook.org/paper.shtml\#invariants} {``Cycle
stability''}. \emph{Stability exponents}\rf{GoSuOr87} are the
corresponding long-time limit, and --in one of those frustrating
historical accidents-- the corresponding
\emph{stability} eigenvectors are misnamed in recent papers
\emph{Lyapunov vectors}, even though \emph{they are not} the
eigenvectors that correspond to the Lyapunov exponents. Sorry, not my
fault. The prefix `covariant' is meant to distinguish the two kinds of
eigenvectors. That's just confusing, for no good reason - Lyapunov
has nothing to do with linear stability described by the \jacobianM\
$\jMps$, as far as I know his paper\rf{Lyap1892}
is about $(\transp{\jMps}\!\jMps)$ and the associated principal axes.

Oseledec proofs are important in mathematics, but not
sensible for computational work in dynamical systems. For me the
Goldhirsch, Sulem and Orszag\rf{GoSuOr87} exposition is the clearest
(read it \HREF{http://chaosbook.org/library/GoSuOr87.pdf}{here}).
They correctly distinguish \emph{Lyapunov} eigenvalues and eigenvectors
from the \emph{stability} eigenvalues and eigenvectors.

At that time Goldhirsch \etal\ had no proof that the long-time Lyapunov
exponents converge to the stability exponents (Kazumasa, do you have
some more recent paper that you prefer to Goldhirsch \etal?).

Trevisan and Pancotti\rf{TrePan98} \emph{Periodic orbits, Lyapunov
vectors, and singular vectors in the Lorenz system} (see above, read
it \HREF{http://chaosbook.org/library/TrePan98.pdf}{here}))
apparently need to be cited for the observation that covariant
vectors reduce to Floquet eigenvectors in the particular case of a
{\po}. Seems so obvious it is in ChaosBook without attribution, and
Ruelle and Eckmann\rf{eckerg} surely say the same (though I have not
checked). Most importantly for our project, they say:

``The leading Lyapunov vectors, as defined here, as well as the
asymptotic final singular vectors, are tangent to the attractor,
while the leading initial singular vectors, in general, point away
from it. Perturbations that are on the attractor can be found in the
subspace of the leading Lyapunov vectors.''

They have very nice pictures for Lorenz unstable orbits illustrating that.

\item[2013-06-27 Predrag]
Graduate students are brave people who immediately jump into the deep
end of the pool, without any testing. Here are two 2\dmn\ maps, the
Lozi map (for $a=1.85, b=0.3$)
% \PC{need figure of Lozi strange attractor}
\bea
   x_{n+1} &=& 1-a |x_{n}|+ b y_n  \continue
   y_{n+1} &=& x_{n}
\,,
\label{e_lozi_def}
\eea
and the H\'enon map (for $a=1.4, b=0.3$; described at length in
\refchap{c-Henon})
\bea
    x_{n+1}&=&1-ax^2_n+b y_n
        \continue
    y_{n+1}&=& x_n
\,,
\label{eq2.1a}
\eea
for which fixed points are available analytically\rf{DasBuch} (for
the Lozi map all periodic points are available analytically). So see
what Floquet exponents
your program returns for \po s of these humble maps, and then for
ergodic trajectories, before running the programs on ergodic
trajectories in zillion dimensions.

If your programs only apply to continuous time flows (ODEs rather
than maps), ChaosBook\rf{DasBuch} has \eqva\ and \po s for 2\dmn\
Lorenz and R\"ossler flows, and Evangelos has lots of \rpo s for the
\cLe; make sure your programs work on small systems first.

\end{description}


\subsection{\KSe}

\PC{2013-03-19 Predrag copied this from \refref{SCD07}, to set up
the conventions for the \KSe\ calculations.}
%
In the formulation
adopted here, the time evolution of the `flame front velocity'
$u=u(x,t)$ on a periodic domain $u(x,t) = u(x+L,t)$ is given by
\beq
  u_t = F(u) = -{\textstyle\frac{1}{2}}(u^2)_x-u_{xx}-u_{xxxx}
    \,,\qquad   x \in [-L/2,L/2]
    \,.
\ee{ks}
Here $t \geq 0$ is the time, and $x$ is the spatial coordinate.
The subscripts $x$ and $t$ denote partial derivatives with respect to
$x$ and $t$. In what follows
we shall state results of all calculations either in units of the
`dimensionless system size' $\tildeL$, or the system size $L = 2 \pi
\tildeL$. All numerical results presented in this report
are for $\tildeL=22/2\pi \simeq 3.5014$.
Spatial periodicity $u(x,t)=u(x+L,t)$
makes it convenient to work in the Fourier space,
\beq
  u(x,t)=\sum_{k=-\infty}^{+\infty} a_k (t) e^{ i k x /\tildeL }
\,,
\ee{eq:ksexp}
with the $1$-dimensional PDE \refeq{ks}
replaced by an infinite set of
ODEs for the complex Fourier coefficients $a_k(t)$:
\beq
\dot{a}_k= \pVeloc_k(a)
     = ( q_k^2 - q_k^4 )\, a_k
    - i \frac{q_k}{2} \sum_{m=-\infty}^{+\infty} a_m a_{k-m}
\,,
\ee{expan}
where $q_k = k/\tildeL$.
Since $u(x,t)$ is real, $a_k=a_{-k}^\ast$, and we can replace the
sum by a $k > 0$ sum.


\begin{description}

\item[2013-06-27 Predrag to Xiong Ding]
Read this blog, study the relevant \KS\
papers\rf{SCD07,Christiansen97}

\end{description}


\section{Daily log}
\label{sect:dailyBlXD}

This section is the daily (or at least, bi-weekly)  log of the Xiong Ding's
work.

\begin{description}

\item[2013-06-27 Predrag] Xiong Ding  <dingxiong203@gmail.com>,
1. year Georgia Tech physics graduate student, joined the
collaboration as a summer research project.
Xiong Ding will enter notes on his day-to-day progress into this log.

\item[2013-06-27 Evangelos] Maybe Xiong Ding could start with Francesco
    \etal\ review paper\rf{GiChLiPo12} instead of the blog? Francesco
    gave a talk here and made it seem very easy task to implement the
    method for a system of ODEs and also his paper seems readable.
    \\
{\bf [2013-06-27 Predrag]}
I agree, that's why the paper is referred to in the 1. line of
\refsect{sect:introXD} :) So does Kazumasa, see the next entry.

\item[2013-06-27 Kazumasa  to Xiong Ding]
Great! Please read (apart from ChaosBook.org)
%Francesco
Ginelli \etal\rf{ginelli-2007-99}
\HREF{http://prl.aps.org/abstract/PRL/v99/i13/e130601}{PRL}, and the
\HREF{http://arxiv.org/abs/1212.3961}{long follow-up paper}\rf{GiChLiPo12},
(much more detailed than the PRL).
Otherwise you will not understand anything on the codes (and the project).

My C++  code and README is in siminos/kazz/code/. You compile and run
it as C++. Note that, though I extracted a part of my long code
relevant to our project, it contains many functions which you probably
will not need.

Hugues and I
strongly recommend that you write your own code from zero, without
reference to mine. This is the best way to learn what's going on in
the code. Don't try to write a code with full functionality from the
beginning, but start with a minimal code, which only computes, say,
Lyapunov exponents.
Confirming that this works as you expect, you can add a code to compute
the covariant vectors from a simple forward-backward process (and
check if the exponents values computed from the covariant vectors
agree with those from the Gram-Schmidt method). For practical use,
you have to implement the "block-by-block" computation of the
vectors, to overcome memory issues (see discussions in Sec. 4.2 in
Ginelli \etal\rf{GiChLiPo12}, \arXiv{1212.3961}). This is
the first step you should reach in this project.

For numerical integration of the KS equation, I used the
operator-splitting algorithm (Adams-Moulton method + Heun's method),
typically with time step 0.005. For more detail, read Sect.~II~A in
Takeuchi \etal\rf{TaGiCh11}. Stiffness matters, implicit methods are
common ways to overcome this problem,
 and that's why I used the Adams-Moulton method. To further improve,
it's better to split the linear and non-linear terms
 by the operator splitting method and use the implicit method only to
 the linear terms where the stiffness is.
I'm not saying that my algorithm is the best way to simulate the KS
equation, but at least it's suited to practical use.

\item[2013-06-27 Predrag  to Xiong Ding]
A discussion of stiffness in integrating PDEs can be found
\HREF{http://www.pvv.ntnu.no/~berland/talks/berland05expintro.pdf}
{here}. \HREF{http://www.math.ntnu.no/num/expint/} {Here} is Matlab
package for testing 47 (!) schemes.

I agree with Kazumasa: you want to write your own code,
but not necessarily reinvent the wheel. There is KS code various
places in this svn repository, for example in
\texttt{siminos/matlab/}, that you might want to compare performance
of your code with. Evangelos can help you with that.


Study the methods described in
\HREF{http://www.cns.gatech.edu/~predrag/papers/SCD07.pdf}
{Appendix A} of \refref{SCD07}.
The authors have lots of experience and they went to the
    trouble of explaining the best methods they know, so why not read
    them and implement them?

The plot of what you get for $L=22$
Lyapunov spectrum should confirm the published results (see
{\bf [2009-09-13 Ruslan]} on \refpage{sect:LyapKS},
\reffig{fig:lyapSpecCLG}, \reffig{fig:lyapSpec1},
\reffig{fig:lyapSpec}, and \reftab{tab:ks22shad}) and the
figure\PC{which one? Always state the number} in Kazumasa \etal\
paper\rf{TaGiCh11} (read it
\HREF{http://chaosbook.org/library/KoSa11.pdf} {here}).

\item[2013-06-27 Ruslan]
You can use the method developed by Cox and
Matthews\rf{cox02jcomp} and improved by Kassam and
Trefethen\rf{ks05com}, where the linear part is integrated exactly.
My Matlab code {\tt ksfmetd2.m} is based on this method and it is
stable and reasonably accurate with time step as large as 0.25.  It
also solves variational problem, so can be used to calculate Lyapunov
exponents.  See {\tt ksfmlyap.m} in {\tt siminos/matlab/ruslan/},
read {\tt 00ReadMe.txt}.  If you want to compute Lyapunov exponents
using the standard GS procedure, the relevant code is in {\tt
ksdupo.m} in the same folder (look for the cell "\%\% Compute
Lyapunov exponents of KS (using ksfmlyap and GS)").

\item[2013-06-27 Predrag]
Ruslan uses the same \KSe\ convention as Kassam and
Trefethen\refrefs{ks05com}. I have saved the Trefethen Matlab code
\HREF{http://chaosbook.org/extras/Trefethen/kursiv.m} {here}. Perhaps
you want to c++ it, see how it runs for you, or run it in Matlab and
compare with your code. There might be something else useful on
\HREF{http://chaosbook.org/extras/}{ChaosBook.org/extras} homepage,
for example the simulations by the spring 2007 GaTech chaos class.
You can search the blog for 'Trefethen' for other discussions
(Kazumasa has reservations about the Trefethen\rf{ks05com} algorithm,
but Ruslan is OK with it).
Siminos has other codes, if needed, on a different repository.

\item[2013-01-21 Evangelos] The \KS\ data you need is in \\
\texttt{siminos/matlab/ruslan/kse22orbits.mat},
\\
in a structure called \texttt{eq}.
Eigenvalues are in the field eq.eig and right
eigenvectors are in the field eq.evec. [e.g. eq(k).evec(:,1) is the
eigenvector which corresponds to the first eigenvalue eq(k).eig(1) of
the k'th equilibrium]. However, I have not used the data for a long
time so, it would be better if Ruslan verifies how the Fourier modes
are stored (I think that the numbers in a column vector correspond to
real and imaginary part of the Fourier modes
\[
 (a_1,\, b_1,\, a_2,\, b_2,\, \ldots a_N,\, b_N)
\]
and thus here there are $31+1$ complex Fourier modes (the zero'th mode
is not included)).

In order to get the left eigenvectors you will need to compute the
stability matrix.

If you run into problems or have questions please email me so that we
can arrange to talk through Skype.


\item[2013-06-27 Ruslan] To compute stability matrix in Matlab, use
\\
\texttt{siminos/matlab/ruslan/ksfm.m}:
\\ {\tt [f, df] = ksfm(0,eq(1).a,22.0)},
where {\tt df} will be the stability matrix of $EQ_1$.

The idea of the method is described in \refref{SCD07} Appendix A and
B.  The Jacobian matrix is calculated from (B.1) which uses solutions
of (A.4) and (A.6).  Matlab code {\tt ksfmetd2.m} solves (A.4) and
(A.6) simultaneously.  The Jacobian matrix is output in {\tt da}.
You don't need anything else.

\item[2013-06-14 Predrag]
Take a few days to read through this blog, and then we talk,
Skype or in person.
Enjoy :)

To start working on the covariant vectors project,
the
\HREF{http://www.cns.gatech.edu/CNS-only/subv.html} {subversion repository}
(type   cnsuser   cnsweb) is called
\begin{verbatim}
svn checkout svn://zero.physics.gatech.edu/siminos
username, password:     xiong and Lyapunov
\end{verbatim}
Chris Marcotte <christopher.marcotte@gmail.com, can help you get
started. Daniel Borrero dborrero@gatech.edu and Qi Ge
    qge30@gatech.edu also know how to do it. Failing that, I'll help
you via Skype predragcvitanovic.

\item[2013-07-03 Xiong Ding] Here is my first blog post:
A question about the \KS\ equation.
It seems that there are many versions of KS equation,
and some of them are totally different.
The one you used in \refref{SCD07}, ``On the state space geometry of
the KS flow in a periodic domain'' is
\beq
 u_t=-\frac{1}{2}(u^2)_x-u_{xx}-u_{xxxx} \ee{XD-KSeSCD07} However, the
equation in \refref{KNSks90}, ``Back in the saddle again: a computer
assisted study of the KS equation'' is:
\beq
 u_t+4\bigtriangleup^2 u
    +\alpha (\bigtriangleup u +\frac{1}{2}(\bigtriangledown u)^2)=0
\ee{XD-KSeSCD07}
The first order term is different.

\item[2013-07-04 Predrag] That is always annoying when one reads
KS literature - I asked that Evangelos include various transformations
in his thesis, but I cannot find that text, only this:
``The parameter $\alpha$ of \refrefs{KNSks90,ksgreene88} is
related to the system size by $\tildeL=\sqrt{\alpha/4}$.''

In Evangelos
\HREF{../blog/blog.pdf} {blog} we write

``
The {1\dmn} \KSe\ as given in \refrefs{ku,siv} is (up to overall scaling
factors):
\beq
    y_t=-y_x^2/2-y_{xxxx}-y_{xx}
\,,\qquad       x \in [0,L]
\,,
    \label{eq:KSeOR}
\eeq
with periodic boundary conditions in the $[0,L]$ interval. The form used
here
\beq
    u_t=uu_x-u_{xxxx}-u_{xx}
\,.
    \label{eq:KSeAP}
\eeq
is obtained from \refeq{eq:KSeOR} by differentiating with respect to $x$
and setting \PCedit{$u=-y_x$}.
In study of {1\dmn} \KS\ \eqva\ $x$ in \refrefs{LanThesis,ksgreene88}
is interpreted as ``time", so $y$ is a height of a front,
and $y_x$ its ``velocity".
In the literature  both forms of the equation are
referred to as the \KSe.
''

In Evangelos blog I rescued the text that he did not use in his thesis,
copied below.
Let us know whether this resolves your quandry,
and please do remember to include this in your project
report, so it is available to future students (I should really
include it into an appendix to ChaosBook...)            \toCB

\item[2007-01-20 Evangelos]
Notes on Greene and Kim\rf{ksgreene88}. A copy is in ChaosBook.org/library,
\HREF{http://chaosbook.org/library/index.html}{click here}.

\noindent\textbf{\Eqva\ according to Greene and Kim:}
%
The form of \KSe\ studied by Greene and Kim\rf{ksgreene88} is
\beq
    y_t=-4y_{xxxx}-\alpha\left(y_{xx}+\frac{1}{2}y_x^2
            -\frac{1}{4\pi}\int_0^{2\pi}y_x^2\ dx\right)
\,,\qquad       x \in [0,2\pi]
\,,
    \label{eq:KSeGreeneKim}
\eeq
with  periodic boundary condition on the interval $[0,2\pi]$.
Mean elastic energy density of a spatial profile $y(x,t)$ is defined by
\beq
    E=\frac{1}{2\pi}\int_0^{2\pi}y_x^2\, dx\,.
    \label{KSenergy}
\eeq
% \ES{This definition does not only apply to equilibria. Greene and
% Kim also study the transfer of energy between modes
% in non-stationary trajectories.}.
Taking the derivative of \refeq{eq:KSeGreeneKim}
with respect to $x$ and substituting $y_x=-u$ leads to
\[
    u_t=4u_{xxxx}+\alpha\left(u_{xx}-uu_x\right)
\,.
\]
Rescaling
\beq
    \tilde{x}=\frac{\sqrt{\alpha}}{2} x
\,,\qquad
    \tilde{t}=\frac{\alpha^2}{4} t
\,,\qquad
    \tilde{u}=\frac{2}{\sqrt{\alpha}} u
    \label{eq:GKscale}
\eeq
brings the Greene-Kim equation to the form \refeq{eq:KSeAP} used here.
The dimensionless system size $\tildeL=L/2\pi$ is related to
the Greene-Kim parameter
through $\tildeL=\sqrt{\alpha}/2$.
The system size $L=22$ studied here corresponds to $\alpha=49.0395$.

The ``kinetic energy'' reads in our units:
\PC{shouldn't there be a prefactor $\frac{1}{2}$?}
\beq
    \tilde{E}=\frac{1}{L}\int_0^{L}\tilde{u}^2\, d\tilde{x}\,.
\eeq
Integrating \refeq{ks} in $[0,L]$ we get $c=\tilde{E}$,
since the terms involving $u_x$ and $u_{xxx}$ vanish due to periodicity.
From the scalings \refeq{eq:GKscale} we have $\tilde{E}=\frac{4}{\alpha}E$.


\item[2013-07-03 Xiong Ding] In \refref{SCD07} you point out KS
    equation is Galilean invariant: if $u(x,t)$ is a solution, then
    $u(x-ct)-c$ is also a solution. but I am confused, because when I
    substitute it into the equation I get a minus sign. If I change the
    form to be $u(x-ct)+c$, then it is OK.

\item[2013-07-04 Predrag] Thanks for noticing this - you are correct, see
2nd paragraph of Section ChaosBook {\em 26.1.1 Scaling and symmetries.}
(all ChaosBook section and eq.~numbers refer to the stable
\HREF{http://chaosbook.org/version14/pdf.shtml} {version 14} of Dec.
31 2012). The error has propagated from Evangelos
\HREF{http://chaosbook.org/projects/theses.html} {thesis},
{\em section 5.1.2 Symmetries of Kuramoto-Sivashinsky system} to the article.

Embarrassing, but too small an error to change the arXiv version of \refref{SCD07}
(though Evangelos could fix it in his thesis)...

Now I note that the ChaosBook {\em Exercise
26.1.
Galilean invariance of the Kuramoto-Sivashinsky equation} is also in error.
As you are checking the KS papers anyhow, you would help me a lot if you also
checked {\em Chapter 27 Turbulence?} as you went along - that chapter is still
only a draft, has been very hard to write up concisely...
Here is the exercise - please correct it
(I'll transfer it back to ChaosBook):                   \toCB

\Exercise{Galilean invariance of the \KSe.}{\label{exer:GalInv}
%Predrag         4apr2013
\index{Galilean invariance}
\index{invariance!Galilean}
\begin{enumerate}
\item
    Verify that the  \KSe\ is
Galilean invariant: if $u(x,t)$ is a solution, then
$v+u(x+2vt,t)$, with $v$ an arbitrary constant velocity, i
s also a solution.
\item
Verify that mean
\[
\expct{u} = \frac{1}{L} \int_L \! dx \, u
\]
is conserved by the flow.
\item
Argue that the choice
%RESTORE in CB: \refeq{GalInv}
of the vanishing  mean
velocity, $\expct{u}=0$ leads
to no loss of generality in calculations that follow.
\item
\tangent
[thinking is extra cost] Inspection of various
``turbulent" solutions of \KSe\ reveals subregions of
``traveling waves" with locally nonzero  $\expct{u}$.
Is there a way to use Galilean invariance locally, even though
we eliminated it by the $\expct{u}=0$ condition?
\end{enumerate}
    } % end \exercise{Galilean invariance of the \KSe

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\solution{exer:GalInv}{Galilean invariance of the \KSe.}{
% Predrag                           4jul2013
    %
{\bf [2013-06-14 Predrag to Xiong Ding]}
Please write up your solution here; but first let's get subversion
working, so you can check out the ChaosBook source code.
        }% end \solution{Galilean invariance of the \KSe



\item[2013-07-04 Predrag] I assume we are going to work on the
    covariant vectors project, so please try to learn how to check out
    the repository (see {\bf [2013-06-14 Predrag]} above).



\end{description}

\renewcommand{\ssp}{a}
