\section{\textsc{Publications}}
% \subsection{In Press %{\color{see}\small available at \href{www.cns.gatech.edu/~siminos/publications.php}{\url{www.cns.gatech.edu/~siminos/publications.php}} }
% }
\subsection{In preparation}
	\cvitem{}{\siminos, M.~Grech, S.~Skupin, T.~Schlegel, and V.\,T.~Tikhonchuk,
	{\emtitle When does a relativisticaly intense laser pulse propagate in a plasma?}, 
	in preparation (2012) %{\color{see}\httplink[arxiv:1101.1318]{arxiv.org/abs/1101.1318}}
	}
\subsection{Submitted}
	\cvitem{}{P.~Cvitanovi{\'c}, D.~Borrero-Echeverry, K.~M.~Carroll, B.~Robbins, \siminos\ and
	L. Zhang,
	{\emtitle Cartography of high-dimensional flows: A visual guide to sections and slices}, 
	submitted to CHAOS (2012) %{\color{see}\httplink[arxiv:1101.1318]{arxiv.org/abs/1101.1318}}
	}
\subsection{Journal Articles}
	\cvitem{\doi{http://dx.doi.org/10.1063/1.3624498}
	\pdf{http://www.cns.gatech.edu/~siminos/papers/SSL11-2.pdf}\cntPaper}
	{G.~S\'anchez-Arriaga, \siminos\ and E.~Lefebvre,
	{\emtitle Relativistic solitary waves with phase modulation embedded in long laser pulses in plasmas,} 
	{\emjournal Phys. Plasmas} \textbf{18} 082304 (2011) %{\color{see}\httplink[arxiv:1101.1318]{arxiv.org/abs/1101.1318}}
	}
	\cvitem{\doi{http://dx.doi.org/10.1103/PhysRevE.83.056402}
	\pdf{http://www.cns.gatech.edu/~siminos/papers/sbg11.pdf}\cntPaper}
	{\siminos, D.~B\'enisti and L.~Gremillet, 
	{\emtitle Stability of nonlinear Vlasov-Poisson equilibria through spectral deformation and Fourier-Hermite expansion,} 
	{\emjournal Phys.~Rev.~E} \textbf{83} 056402 (2011) %{\color{see}\httplink[arxiv:1101.1318]{arxiv.org/abs/1101.1318}}
	}
	\cvitem{\doi{http://dx.doi.org/10.1088/0741-3335/53/4/045011}
	\pdf{http://www.cns.gatech.edu/~siminos/papers/SSL10.pdf}\cntPaper}
	{G.~S\'anchez-Arriaga, \siminos\ and E.~Lefebvre, 
	{\emtitle Relativistic solitary waves modulating long laser pulses in plasmas,} 
	{\emjournal Plasma~Phys.~Contr.~Fusion} \textbf{53}, 045011 (2011)%{,\color{see}\httplink[arxiv:1101.1004]{arxiv.org/abs/1101.1004}}
	}
 	\cvitem{\doi{http://dx.doi.org/10.1016/j.physd.2010.07.010}
	\pdf{http://www.cns.gatech.edu/~siminos/papers/SiCvi10.pdf}\cntPaper}
	{\siminos\ and  P.~Cvitanovi\'c, 
 	{\emtitle Continuous symmetry reduction and return maps for high-dimensional flows,}
 	{\emjournal Physica D} \textbf{240}, 187--198 (2011)
	}
 	\cvitem{\doi{http://dx.doi.org/10.1103/PhysRevLett.105.015001}
	\pdf{http://www.cns.gatech.edu/~siminos/papers/PhysRevLett.105.015001.pdf}\cntPaper}
	{D.~B\'enisti, O.~Morice, L.~Gremillet, \siminos\ and D.\,J.~Strozzi,
	{\emtitle Self-organization and threshold of stimulated Raman scattering,}
	{\emjournal Phys. Rev. Lett.} \textbf{105}, 015001 (2010)
	}
	\cvitem{\doi{http://dx.doi.org/10.1063/1.3464467}
 	\pdf{http://www.cns.gatech.edu/~siminos/papers/nlin_group.pdf}\cntPaper}
	{D.~B\'enisti, O.~Morice, L.~Gremillet, \siminos\ and D.\,J. Strozzi,
	{\emtitle Nonlinear group velocity of an electron plasma wave,}
	{\emjournal Phys. Plasmas} \textbf{17}, 082301 (2010)
	}
	\cvitem{\doi{http://dx.doi.org/10.1063/1.3494223}
	\pdf{http://www.cns.gatech.edu/~siminos/papers/braman.pdf}\cntPaper}
	{D.~B\'enisti, O.~Morice, L.~Gremillet, \siminos\ and D.\,J.~Strozzi,
	{\emtitle Nonlinear kinetic description of Raman growth using an envelope code, and comparisons with Vlasov simulations,}
	{\emjournal Phys. Plasmas} \textbf{17}, 102311 (2010)
	}
 	\cvitem{\doi{http://dx.doi.org/10.1137/070705623}
	\pdf{http://www.cns.gatech.edu/~siminos/papers/CDS10.pdf}\cntPaper}
	{P.~Cvitanovi\'c, R.\,L.~Davidchack and \siminos,
 	{\emtitle On the State Space Geometry of the Kuramoto-Sivashinsky Flow in a Periodic Domain,}
 	{\emjournal SIAM J. Appl. Dyn. Syst.} \textbf{9}, 1 (2010)
	}
\subsection{Thesis}
	\cvitem{\html{http://smartech.gatech.edu/handle/1853/28215}\pdf{http://www.chaosbook.org/projects/Siminos/thesis.pdf}}{E. Siminos, Recurrent spatio-temporal structures in presence of continuous symmetries,
	  PhD Thesis, School of Physics, Georgia Institute of Technology,  Atlanta, GA, USA, May 2009
	}
  