%%%% for public version, toggle \draftfalse in setup2modes.tex
%    (that removes all comments, the blog)

% reducesymm/cgang/2modes.tex    this is master file:    pdflatex 2modes
%     then:    pdflatex def2modes; bibtex def2modes; pdflatex def2modes; pdflatex def2modes

% until 2012-08-20 this was in svn repo siminos/cgang/2modes.tex

\documentclass[aip,cha,
reprint,
secnumarabic,
nofootinbib, tightenlines,
nobibnotes, showkeys, showpacs,
superscriptaddress,
%preprint,%
%author-year,%
%author-numerical,%
]{revtex4-1}

\newcommand{\version}{2modes ver. 2.4, Nov 15 2014}
% Daniel                    ver. 2.4, Nov 15 2014 - first version past arxiv v1.0 submission
% Predrag                   ver. 2.3, Nov 11 2014
% Predrag                   ver. 2.2, Jul 24 2014
% Predrag                   ver. 2.1, Jul 14 2014
% Burak                     ver. 2.0, Jul  8 2014
% Predrag                   ver. 1.3, May 11 2014
% Burak                     ver. 1.2, May  6 2014
% Predrag                   ver. 1.1, Nov 16 2013
% Burak                     ver. 1.0, Oct  6 2013
% Predrag                   ver. 0.3, Aug  1 2012
% Predrag                   ver. 0.2, Apr 30 2012}
% Predrag from atlas12      ver. 0.1, Apr 25 2012}

        \input setup2modes
        \input def2modes

\begin{document}

\title[Periodic orbit analysis of a system with continuous symmetry]
{Periodic orbit analysis of a system with continuous symmetry - a tutorial}

\author{Nazmi Burak Budanur}
\email{budanur3@gatech.edu}
\affiliation{
 School of Physics and Center for Nonlinear Science,
 Georgia Institute of Technology,
 Atlanta, GA 30332
}
\author{Daniel Borrero-Echeverry}
\affiliation{
 School of Physics and Center for Nonlinear Science,
 Georgia Institute of Technology,
 Atlanta, GA 30332
}
\affiliation{
 Department of Physics,
 Reed College,
 Portland OR 97202
}
\author{Predrag Cvitanovi\'{c}}
\affiliation{
 School of Physics and Center for Nonlinear Science,
 Georgia Institute of Technology,
 Atlanta, GA 30332
}
    \ifdraft
\date{\today}
    \else
\date{14 November 2014}
   \fi

\begin{abstract}
Dynamical systems with translational or rotational symmetry arise
frequently in studies of spatially extended physical systems, such as
Navier-Stokes flows on periodic domains. In these cases, it is natural to
express the state of the fluid in terms of a Fourier series truncated to a
finite number of modes. 
\revision{
Here we study a 4-dimensional toy model with chaotic dynamics and 
\SOn{2} symmetry similar to those appear in fluid dynamics problems. 
} 
A crucial step in the analysis of such a system
is symmetry reduction. We use the model to illustrate different
symmetry-reduction techniques. Its relative equilibria are conveniently
determined by rewriting the dynamics in terms of a symmetry-invariant
polynomial basis. However, for the analysis of its chaotic dynamics, the `method of
slices', which is applicable to very high-dimensional problems, is
preferable. We show that a Poincar\'e section taken on the `slice' can be
used to further reduce this flow to what is for all practical purposes a
unimodal map. This enables us to systematically determine all relative
periodic orbits and their symbolic dynamics up to any desired period. We
then present cycle averaging formulas adequate for systems with continuous symmetry 
and use them to compute dynamical averages using relative periodic orbits. The convergence
of such computations is discussed.
\end{abstract}

\pacs{02.20.-a, 05.45.-a, 05.45.Jn, 47.27.ed, 47.52.+j, 83.60.Wc}
\keywords{
symmetry reduction,
equivariant dynamics,
relative equilibria,
relative periodic orbits,
periodic orbit theory,
method of slices,
moving frames, chaos
}
\maketitle

\begin{quotation}
Periodic orbit theory provides a way to compute dynamical averages for
chaotic flows by means of {\cycForm s} that relate the time averages of
observables to the spectra of unstable periodic orbits. Standard
{\cycForm s} are valid under the assumption that the stability
multipliers of all periodic orbits have a single marginal direction
corresponding to time evolution and are hyperbolic in all other
directions. However, if a dynamical system has $N$ continuous symmetries,
periodic orbits are replaced by relative periodic orbits, invariant
$(N+1)$-dimensional tori with marginal stability in $(N+1)$ directions.
Such exact invariant solutions arise in studies of turbulent flows, such
as the pipe flow or the plane Couette flow, which have continuous symmetries.
In practice, the translational invariance of these flows is approximated by 
carrying out simulations in periodic domains so that the state of the fluid 
is conveniently expressed as a Fourier series, truncated to a large but finite 
number (from tens to thousands) of Fourier modes. This paper is a tutorial on 
how such problems are to be analyzed using periodic orbit theory. We illustrate 
all the necessary steps using a simple `\twomode' model as an example.
\end{quotation}

\section{Introduction}
\label{s:intro}

Recent experimental observations of traveling waves in pipe flows have
confirmed the intuition from dynamical systems theory that invariant solutions
of \NSe\ play an important role in shaping the \statesp\ of turbulent
flows\rf{science04}. When one casts fluid flow equations in a
particular basis, the outcome is an infinite dimensional dynamical system
that is often equivariant under transformations such as
translations, reflections and rotations. For example, when periodic
boundary conditions are imposed along the streamwise direction, the equations
for pipe flow retain their form under the action of streamwise translations, 
azimuthal rotations and reflections about the central axis, \ie, they are equivariant
under the actions of $\SOn{2} \times \On{2}$. In this case it is natural
to express the state of the fluid in a Fourier basis. However,
as the system evolves, the nonlinear terms in the equations mix the
various modes, so that the state of the system evolves not only along the
symmetry directions, but also along directions transverse to them.
This complicates the dynamics and gives rise to high dimensional coherent
solutions such as \reqva\ and \rpo s, which take on the roles played by
\eqva\ and \po s in flows without symmetry.

This paper provides a step-by-step tutorial on how periodic orbit theory can be 
applied to flows with continuous symmetries. Such an analysis should ultimately 
be applicable to turbulent flows, once sufficiently many exact invariant solutions
become numerically accessible. As an example, we study a \twomode\
\SOn{2} equivariant 
\revision{normal form} 
that has the minimal dimensionality required for chaotic
dynamics. The rest of the paper is organized as follows: In \refsect{s:symm} we
define basic concepts and briefly review the relevant symmetry reduction
literature. In \refsect{s:twoMode}, we introduce the \twomode\ model
system and discuss several of its symmetry-reduced representations, \DBedit{and utilize a
symmetry-reduced polynomial representation to find the only \reqv\ of the
system}\DB{201-11-15}{I would throw this out, at least as a highlight in the ``review'' of the
paper. Our take-home message is that you could use our approach on turbulent flows, so the
fact that we can cheat by using invariant polynomials shouldn't be celebrated}
. In \refsect{s:numerics}, we show how the \mslices\ can be used to
quotient the symmetry and reduce the dynamics onto a symmetry-reduced
\statesp\ or `\slice '. A Poincar\'e section taken on the \slice\ then
reduces the 4\dmn\ chaotic dynamics in the full \statesp\ to an approximately
one-dimensional, unimodal Poincar\'e return map. The return map is then
used to construct a finite grammar symbolic dynamics for the flow and
determine {\em all} \rpo s up to a given period. In \refsect{s:DynAvers},
we present {\cycForm s} adequate for systems with continuous symmetries
and use the relative periodic orbits calculated in \refsect{s:numerics} 
to calculate dynamically interesting observables. Finally, in \refsect{s:concl},
we discuss possible applications of the \mslices\ to various spatially
extended systems.

The main text is supplemented by two appendices. \refAppe{s:newton} describes 
the multi-shooting method used to calculate the \rpo s.
\refAppe{s:schur} discusses how periodic Schur decomposition can be used
to determine their Floquet multipliers, which can differ by 100s
of orders of magnitude even in a model as simple as the \twomode\ system.


\input{contsymm}
\input{flow}
\input{numerics}
\input{recycle}
\input{conclusion}

\begin{acknowledgments}
We are grateful to Evangelos Siminos for his contributions to this project
and Mohammad M.~Farazmand for a critical reading of the manuscript.
We acknowledge stimulating discussion with
Xiong Ding,
Ruslan L.~Davidchack,
Ashley P.~Willis,
Al Shapere
and
Francesco Fedele.
We are indebted to the 2012 ChaosBook.org class, in particular to
Keith M.~Carroll,
Sarah Flynn,
Bryce Robbins,
and
Lei Zhang,
for the initial fearless fishing expeditions into the enormous sea of the
parameter values of the \twomode\ model.
P.~C.\ thanks the family of late G.~Robinson,~Jr.
and
NSF~DMS-1211827 for support. D.~B.\ thanks M.~F.\ Schatz for support during
the early stages of this work under NSF~CBET-0853691.
\end{acknowledgments}

\appendix
\input{newton}
\input{schur}

\bibliography{../bibtex/siminos}


%%%%%%%%% end of the paper proper %%%%%%%%%%%%%%%%%%%%%%%%

\ifdraft
    \onecolumngrid

    \newpage
\input ../blog/flotsam2modes
% \input flotsam % Predrag 2014-07-14 absorbed into flotsam2modes
    \newpage
    \section{{\twoMode} simulations blog}
    \label{chap:Mathematica}
\input ../blog/Mathematica

    \newpage
    \section{{\twoMode} daily blog}
    \label{chap:2modes}
\input ../blog/2modes
    \newpage
    \section{Burak' s {\twomode}}
    \label{chap:2modesBB}
\input ../blog/2modesBB % Predrag 2013-0810 Burak, git version only

\addcontentsline{toc}{section}{last blog entry}

\fi % end \ifdraft

\end{document}
