\section{Conclusions and discussion}
\label{s:concl}

In this tutorial, we have studied probably the simplest dynamical system which 
exhibits chaos and is equivariant under a continuous symmetry transformation.
We have shown that reducing this symmetry simplifies the qualitative dynamics 
to a great extent and enables one to find all \rpo s of the systems via 
conventional techniques such as Poincar\'e sections and return maps. In 
addition, we have also shown that one can also extract quantitative information 
from the \rpo s by computing cycle averages. 

We motivated our study of the \twomode\ system by the resemblence of its 
symmetry structure to that of the spatially extended systems; and the steps we 
outlined here are, in principle, applicable to physical systems that are 
described by $N$-Fourier mode truncations of PDEs such as $1D$ 
Kuramoto-Sivashinsky\rf{SCD09b}, $3D$ pipe flows\rf{WiShCv14}, etc. However, 
for these higher dimensional problems, there remains other challenges, both 
theoretical and numerical. 

We found in \refsect{s:numerics} that \twomode\ dynamics can be completely 
described by a unimodal return map on the Poincar\'e section we constructed 
after continuous symmetry reduction. In a high dimensional system, finding such 
an easy symbolic dynamics, or any symbolic dynamics at all is a challenging 
problem on its own. For desymmetrized (confined in the odd subspace) $1$D 
Kuramoto-Sivashinsky system \refref{lanCvit07} finds a bimodal return map after 
reducing the discrete symmetry of the problem, however, for turbulent fluids, 
we do not know any study that simplifies the flow to such an extent.

In \refsect{s:DynAvers}, we have shown that the symbolic dynamics and the 
associated grammar rules greatly affects the convergence of the \cycForm s and 
we have just mentioned that finding a finite symbolic description of a flow is 
not so easy. Good news is there exists other methods of ordering cycle 
expansion terms, for example, ordering in their stability and discarding terms
that are above a threshold; so that one expects the remaining terms to form
shadowing combinations and converge faster than exponential. Bad news is 
whichever method of term ordering is deployed, the cycle expansions are as good
as the shortest cycle that one fails to find. Symbolic dynamics solves both 
of problems at once since it puts the cycles into a topological order so that 
one cannot miss any accessible cycle and shadowing combinations naturally occur
when the expansion is ordered in topological length. The question one might ask 
is: When there is no symbolic dynamics, how can we make sure that we find all 
periodic orbits of a flow up to some cycle period? 

In the searches of the cycles of high dimensional flows, one usually looks at
the near recurrences of the ergodic flow, and then runs Newton searches 
starting nearby these recurrences to find if they are influenced by an exact 
recurrence. Such an approach does not answer the question we just asked to full 
confidence, however, one may argue that the dynamically important cycles 
influence recurrences of the ergodic flow, and hence cycles found this way are 
those that are relevant for computing averages. Of course, such an argument can 
only be convincing if it is supported by numerical evidence.

To sum up, we have shown that the periodic orbit theory successfully extends to
the systems with continuous symmetries, yet, one still needs to think about 
remaining challenges, which we discussed above, associated with high 
dimensional systems. Once these are overcome, we would then be able to extract 
quantitative information from turbulence by using its exact unstable solutions, 
which would be big news.
