\section{Cycle Averages}
\label{s:DynAvers}

\renewcommand{\zeit}{\ensuremath{t}}  %time variable

So far, we have explained how to find the \rpo s of the \twomode\ system
and compute their stability. However, we have not yet
said anything about what to do with these numbers. We begin this section with
an overview of the main results of the periodic orbit theory. Our review starts
by recapitulating the presentation of \refref{DasBuch}, but then, in
\refsect{s-ContFac}, explains how
the theory is modified in the presence of continuous symmetries.\rf{Cvi07}
In \refsect{s-CycExp}, we present cycle expansions and
explain how to approximate the Poincar\'e section in
\reffig{fig:psectandretmap}\,(d), in order to obtain a better convergence of
the spectral determinants. The numerical results are discussed in
\refsect{s-NumResults}.

\subsection{Classical trace formula}

Consider the {\evOper} $\Lop^\zeit$, the action of which evolves a
weighted density $\rho(\ssp,\zeit)$ in the \statesp,
\bea
    \rho(\ssp',\zeit) &=& [\Lop^\zeit \rho ] (\ssp')
    = \int\!d \ssp \, \Lop^\zeit (\ssp', \ssp) \, \rho(\ssp,0)
    \continue
    \Lop^\zeit (\ssp', \ssp)
    &=&
    \delta (\ssp' - \flow{\zeit}{\ssp})\,
        e^{\beta \Obser^\zeit(\ssp)}
\,,
\label{e-EvOper}
\eea
where $\beta$ is an auxiliary variable and $\Obser^\zeit (\ssp)$ is the
integrated value of an observable $\obser (\ssp)$ along the trajectory
$\ssp(\zeit)=\flow{\zeit}{\ssp}$,
\beq
    \Obser^\zeit (\ssp ) = \int_0^{\zeit} d \zeit'
                              \obser(\flow{\zeit'}{\ssp})
\,.
\eeq
When $\beta = 0$, the \evOper\ \refeq{e-EvOper} evolves the initial density of
\statesp\ points to its new form after time $\zeit$; this form of the
evolution operator is known as the {\FPoper}. The
multiplicative weight $\exp(\beta \Obser^\zeit(\ssp))$
will enable us to compute the value of the observable $\obser$ averaged over
the natural measure.

As the integrated observable $\Obser^\zeit (\ssp)$, additive along the
trajectory, is exponentiated in \refeq{e-EvOper}, the \evOper\
is multiplicative along the trajectory:
\beq
   \Lop^{\zeit_1 + \zeit_2} (\ssp',\ssp) =
    \int d\ssp'' \Lop^{\zeit_2} (\ssp', \ssp'')
                   \Lop^{\zeit_1} (\ssp'', \ssp) \, .
   \label{eq-SemiGroupKernel}
\eeq
This semigroup property allows us to define the {\evOper} as the formal
exponential of its infinitesimal generator \Aop:
\beq
   \Lop^\zeit = e^{\Aop \zeit} \, .
   \label{eq-EvOpExp}
\eeq
Let $\rho_{\beta} (\ssp)$ be the eigenfunction of
\refeq{e-EvOper} corresponding to the leading eigenvalue of $\Aop$ (\ie, the one
with the largest real part) for a given $\beta$,
\beq
    \left[ \Lop^\zeit \rho_{\beta} \right] (\ssp)
    =
    e^{\zeit s(\beta )} \rho_{\beta}(\ssp)
    \, .
    \label{eq-EigenvalueRel}
\eeq
If the system under study is
ergodic, then an invariant `natural measure' $\rho_0(\ssp)$ with
eigenvalue $s(0) = 0$ exists, and the long time average of an observable is
then its \statesp\ average over the natural measure:
\beq
    \langle \obser \rangle = \int d \ssp \, \obser(\ssp) \rho_0 (\ssp) \, .
    \label{e-obserAvg}
\eeq
By evaluating the action of the {\evOper} \refeq{e-EvOper} for
infinitesimal times, one finds that the
long-time averages of
observables, as well as of their higher moments, are given by
derivatives of $s(\beta)$:
\bea
    \langle \obser \rangle &=&
    \lim_{\zeit \rightarrow \infty} \frac{1}{\zeit} \langle \Obser^\zeit \rangle =
        \left. \frac{\partial s(\beta) }{\partial \beta} \right|_{\beta = 0}
        \, , \continue
        \Delta \; &=&
        \lim_{\zeit \rightarrow \infty} \frac{1}{\zeit}
        \langle (\Obser^{\zeit})^2 - \langle \Obser^\zeit \rangle^2 \rangle =
        \left. \frac{\partial^2 s(\beta) }{\partial \beta^2} \right|_{\beta = 0}
    \, ,
    \label{eq-moments}
    \\
    \, &\vdots & \nonumber
\eea
For example, if the observable \obser\ is a velocity
$\dot{\ssp}=\vel(\ssp)$, the integrated observable $\Obser^\zeit$ is the
displacement $\ssp(\zeit)$ in $d$ dimensions, and $\Delta/2d$ is
Einstein's diffusion coefficient.

In order to obtain $s(\beta)$, we construct the resolvent of \Aop , by taking
the Laplace transform of \refeq{eq-EvOpExp}:
\beq
   \int_0^{\infty} \!d\zeit\, e^{-s\zeit} \Lop^\zeit = (s-\Aop)^{-1} \, ,
   \label{eq-ResolventA}
\eeq
the trace of which peaks at the eigenvalues of \Aop. By taking the
Laplace transform of $\Lop^\zeit$ and computing its trace
by $\tr \Lop^\zeit = \int d\ssp \, \Lop^\zeit (\ssp,\ssp)$, one obtains the
classical trace formula\rf{pexp}:
\beq
\sum_{\alpha=0}^{\infty} \frac{1}{s-s_{\alpha}} = \sum_p \period{p}
\sum_{r=1}^{\infty} \frac{e^{r(\beta \Obser_p - s \period{p})}}{\oneMinJ{r}}
\ee{e-ClassicalTraceFormula}
that relates the spectrum of the {\evOper} to the spectrum of periodic
orbits. Here  $s$ is the auxiliary
variable of the Laplace transform and $s_{\alpha}$ are the eigenvalues of \Aop . The
outer sum on the right hand side runs over the `prime cycles' $p$ of the system,
\ie, the shortest \po s of period $\period{p}$. $\Obser_p$ is the value of
the observable integrated along the prime cycle and $\monodromy_p$ is the
transverse monodromy matrix, the eigenvalues $\Obser^\zeit (\ssp)$ of
which are the Floquet multipliers of $p$ with the marginal ones excluded.
In the derivation of \refeq{e-ClassicalTraceFormula}, one assumes that
the flow has a single marginal direction, namely the $\vel(\ssp)$ tangent
to the periodic orbit, and evaluates the contribution of each \po\ to the
trace integral by transforming to a local coordinate system where one of
the coordinates is parallel to the flow, while the rest are transverse.
The integral along the parallel direction contributes the factors of
$\period{p}$ in \refeq{e-ClassicalTraceFormula}. The transverse integral
over the delta function \refeq{e-EvOper} contributes the factor of
$1/\oneMinJ{r}$.

\subsection{Decomposition of the trace formula over irreducible representations}
\label{s-ContFac}

The classical trace formula \refeq{e-ClassicalTraceFormula} accounts for contributions from \po s
to long time dynamical averages. However, \rpo s of equivariant systems are almost never
periodic in the full \statesp. In order to compute the contributions of \rpo s
to the trace of the \evOper, one has to factorize
the \evOper\
into the irreducible subspaces of the symmetry group. For discrete symmetries,
this procedure is studied in \refref{CvitaEckardt}. For the quantum systems
with continuous symmetries (abelian and 3D rotations), the factorization of
the semiclassical Green's operator is carried out in \refref{Creagh93}.
Reference \cite{Cvi07} addresses the continuous factorization of the \evOper\ and its
trace; we provide a sketch of this treatment here. 

We start by stating, without
proof, that a square-integrable field $\psi (\ssp)$ over a vector space can be
factorized into its projections over the irreducible subspaces of a group
$\Group$:
\beq
    \psi (\ssp) = \sum_m \mathbb{P}_m \psi (\ssp) \, ,
\eeq
where the sum runs over the irreducible representations of $\Group$ and
the projection operator onto the $m$th irreducible subspace, for a continuous
group, is
    \PC{2014-11-10: From now on, a problem - if I redefine $D(\cdots)$ as \matrixRep,
    cannot easily revert to the concise $g$ notation...}
\beq
    \mathbb{P}_m = d_m \int_\Group d \mu(\LieEl) \chi_m (\LieEl(\theta))
                            \mathbb{D}(\theta)
\,.
\ee{e-ProjectionOperator}
Here, $d_m$ is the dimension of the representation, $d \mu(g)$ is the
normalized Haar measure, $\chi_m (\LieEl)$ is the character of $m$th
irreducible representation, and $\mathbb{D}(\theta)$ is the operator that
transforms a scalar field defined on the \statesp\ as
$\mathbb{D}(\theta) \rho (\ssp) = \rho(\matrixRep(\theta)^{-1} \ssp)$.
    \DB{2014-11-10}{Added missing parenthesis back on 2014-11-3 to
    this expression. Did I put it in the right place? I think so, but please double check and then delete this comment.}
For
our specific case of a single $\SOn{2}$ symmetry,
\bea
d_m &\rightarrow& 1\, , \\
\int_G d \mu(g) &\rightarrow& \oint \frac{d \theta} {2 \pi} \, , \\
\chi_m (\LieEl(\theta)) &\rightarrow& e^{- \ii m \theta } \, .
\eea
Because the projection
operator \refeq{e-ProjectionOperator} decomposes scalar fields defined over the \statesp\
into their irreducible subspaces under action of $\Group$, it can be used to
factorize the \evOper. Thus, the kernel
of the \evOper\ transforms under the action of $\mathbb{D}(\theta)$ as
\bea
    \mathbb{D}(\theta) \Lop^\zeit (\ssp', \ssp) &=&
        \Lop^\zeit (\matrixRep(\theta)^{-1} \ssp', \ssp)\,,
    \continue
    &=& \Lop^\zeit (\ssp', \matrixRep(\theta) \ssp) \,, \continue
    &=& \delta (\ssp' - \matrixRep(\theta) f^\zeit (\ssp)) e^{\beta \Obser^\zeit(\ssp)}\, ,
    \label{e-gEvOper}
\eea
where the second step follows from the equivariance of the system under
consideration. \Rpo s contribute to $\mathbb{P}_m \Lop^\zeit = \Lop_m^\zeit$ since when its
kernel is modified as in \refeq{e-gEvOper}, the projection involves an integral
over the group parameters that is non-zero when $\theta= - \theta_{\rpprime}$, the phase shifts of the
\rpo s. By computing the trace of $\Lop_m^\zeit$, which in addition to the integral
over \statesp , now involves another integral over the group parameters, one
obtains the $m$th irreducible subspace contribution to the classical trace as
\beq
\sum_{\alpha=0}^{\infty} \frac{1}{s-s_{m, \alpha}} = \sum_p T_{\rpprime}
\sum_{r=1}^{\infty} \frac{\chi_m (\LieEl^r(\theta_{\rpprime}))
            e^{r(\beta \Obser_{\rpprime} - s T_{\rpprime})}}{\oneMinJred{r}} .
\ee{e-ReducedTraceFormula}
The reduced trace formula \refeq{e-ReducedTraceFormula} differs from the
classical trace formula \refeq{e-ClassicalTraceFormula} by the group character
term, which is evaluated at the \rpo\ phase shifts, and the reduced monodromy
matrix $\monodromyRed$, which is the $(d-N-1)\times(d-N-1)$ reduced Jacobian
for the \rpo\ evaluated on a Poincar\'e section in the \reducedsp . The eigenvalues
of $\monodromyRed$ are those of the \rpo\ Jacobian \refeq{e-rpoJacobian}
excluding the marginal ones, i.e., the ones corresponding to time evolution and evolution
along the continuous symmetry directions.

Since we are only interested in the leading eigenvalue of the \evOper , we
only consider contributions to the
trace \refeq{e-ClassicalTraceFormula} from the projections
\refeq{e-ReducedTraceFormula} of the $0$th irreducible subspace. For the $\SOn{2}$ case at hand, these can be written
explicitly as
\beq
\sum_{\alpha=0}^{\infty} \frac{1}{s-s_{0, \alpha}} = \sum_p \period{p}
\sum_{r=1}^{\infty} \frac{e^{r(\beta \Obser_p - s \period{p})}}{\oneMinJred{r}} \, .
\ee{e-tracem0}
This form differs from the classical trace formula
\refeq{e-ClassicalTraceFormula} only by the use of the reduced monodromy matrix
\DBedit{instead of the full monodromy matrix}\DB{2014-11-10}{What is the non-reduced monodromy matrix called?
I put down full but don't know if this is the correct nomenclature} since
the $0$th irreducible representation of $\SOn{2}$ has character $1$. For this
reason, cycle expansions,\rf{AACI} which we cover next, are applicable
to \refeq{e-tracem0} after the replacement
$\monodromy \rightarrow \monodromyRed$.

\subsection{Cycle expansions}
\label{s-CycExp}

While the classical trace formula \refeq{e-ClassicalTraceFormula} and its
factorization for systems with continuous symmetry \refeq{e-ReducedTraceFormula} manifest
the essential duality between the spectrum of an observable and that of
the \po s and \rpo s, in practice, they are hard to work with since the
eigenvalues are located at the poles of \refeq{e-ClassicalTraceFormula} and
\refeq{e-ReducedTraceFormula}. The dynamical zeta function
\refeq{e-DynamicalZeta}, which we derive below, provides a perturbative expansion form that
enables us to order terms in decreasing importance while computing
spectra for the \twomode\ system. As stated earlier, \refeq{e-tracem0}
is equivalent to \refeq{e-ClassicalTraceFormula} via substitution
$\monodromy \rightarrow \monodromyRed$, so this derivation works for either. 

We start by defining the
`spectral determinant':
\beq
  \det (s-\Aop) = \exp \left( - \sum_p \sum_{r=1}^{\infty}
      \frac{1}{r} \frac{e^{r(\beta \Obser_p - s \period{p})}}{\oneMinJ{r}} \right)\, ,
\ee{e-SpectralDeterminant}
whose logarithmic derivative ($(d/ds) \ln \det(s - \Aop)$) gives
the classical trace formula \refeq{e-ClassicalTraceFormula}.
The spectral determinant \refeq{e-SpectralDeterminant} \DB{2014-11-10}{What's up with G65?} is easier to work
with since the spectrum of $\mathcal{A}$ is now located at the zeros of
\refeq{e-SpectralDeterminant}. The convergence of \refeq{e-SpectralDeterminant}
is, however, still not obvious. More insight is gained by approximating
$\oneMinJ{r}$ by the product of expanding Floquet multipliers and then
carrying out the sum over $r$ in \refeq{e-SpectralDeterminant}. This
approximation yields
\bea
\oneMinJ{} &=& | (1 - \ExpaEig_{e,1})(1 - \ExpaEig_{e,2})... \continue
         &&(1 - \ExpaEig_{c,1}) (1 - \ExpaEig_{c,2}) ... | \nonumber \\
         &\approx& \prod_e |\ExpaEig_e| \equiv |\ExpaEig_p|,
    \label{e-LambdapApprox}
\eea
where $|\ExpaEig_{e,i}| > 1$ and $|\ExpaEig_{c,i}| < 1$ are expanding and
contracting Floquet multipliers respectively. By making this approximation, the sum over $r$ in
\refeq{e-SpectralDeterminant} becomes the Taylor expansion of natural logarithm. Carrying out this sum, brings the
spectral determinant \refeq{e-SpectralDeterminant} to a product (over prime
cycles) known as the dynamical zeta function:
\beq
1 / \zeta = \prod_p (1 - t_p) \, \mbox{where}, \, t_p = \frac{1}{|\ExpaEig_p|}
            e^{\beta \Obser_p - s \period{p}} z^{n_p} .
\ee{e-DynamicalZeta}
Each `cycle weight' $t_p$ is multiplied by the `order tracking term' $z^{n_p}$,
where $n_p$ is the topological length of the $p$th prime cycle. This polynomial
ordering arises naturally in the study of discrete time systems where the
Laplace transform is replaced by $z$-transform. Here, we insert the powers of
$z$ by hand, to keep track of the ordering, and then set its value to $1$ at the
end of calculation. Doing so allows us to write the dynamical zeta function
\refeq{e-DynamicalZeta} in the `cycle expansion' form by grouping its
terms in powers of $z$. For complete binary symbolic dynamics, where every binary symbol
sequence is accessible, the cycle expansion reads
\bea
1 / \zeta &=& 1 - t_0 - t_1 - (t_{01} - t_0 t_1 )  \label{e-CycleExpansion} \\
        && - [(t_{011} - t_{01}t_1) + (t_{001} - t_{01} t_0)] - ... \continue
        &=& 1 - \sum_f t_f - \sum_n \hat{c}_n \label{e-CurvatureExpansion},
\eea
\DBedit{where we labeled each prime cycle by its binary symbol sequence. In
\refeq{e-CurvatureExpansion} we grouped the contributions to the zeta function
into two groups: `fundamental' contributions $t_f$ and `curvature' corrections $c_n$.
The curvature correction terms are denoted explicitly by parentheses in \refeq{e-CycleExpansion} and
correspond to `shadowing' combinations where combinations of
shorter cycle weights, also known as `pseudocycle' weights, are subtracted from the weights of longer
prime cycles. Since the cycle weights in \refeq{e-DynamicalZeta} already
decrease exponentially with increasing cycle period, the cycle expansion
\refeq{e-CycleExpansion} converges even faster than exponentially when the
terms corresponding to longer prime cycles are shadowed.}\DB{2014-11-10}{Tried to make this clearer. Think I didn't change the content, but somebody more erudite than me please double check this.}

For complete binary symbolic dynamics, the only fundamental contributions to
the dynamical zeta function are from the cycles with topological length $1$, and all
longer cycles appear in the shadowing pseudocycle combinations.
More generally, if the symbolic dynamics is a subshift of finite type,\rf{DasBuch}
with the grammar of admissible sequences described by a finite set of pruning rules,
and the flow is uniformly hyperbolic, cycle expansions of {\Fd s}
are guaranteed to converge super-exponentially.\rf{hhrugh92}
A generic unimodal map symbolic dynamics is not a subshift of finite type.
However, we have shown in \refsect{s:numerics} that the Poincar\'e return map for
the \twomode\ system (\reffig{fig:psectandretmap}\,(d)) diverges at
$s \approx 0.98$ and approximated it as if its tip was located at the
furthest point visited by an ergodic trajectory. This brings the question of
whether we can approximate the map in \reffig{fig:psectandretmap}\,(d) in such a way
that corresponding symbolic dynamics has a finite grammar of pruning rules?
The answer is yes.

As shown in \reffig{fig:psectandretmap}\,(d), the cycles \cycle{001}
and \cycle{011} pass quite close to the tip of the cusp. Approximating the
map as if its tip is located exactly at the point where \cycle{001} cuts gives us
what we are looking for: a single grammar rule, which says that the symbol
sequence `00' is inadmissible. This can be made rigorous by the help of
kneading theory, however, the simple result is easy to see from the return map
in \reffig{fig:psectandretmap}\,(d): Cover the parts of the return map, which
are outside the borders set by the red dashed lines, the cycle \cycle{001} and
then start any point to the left of the tip and look at images. You will always
land on a point to the right of the tip, unless you start at the lower left
corner, exactly on the cycle \cycle{001}. As we will show, this `finite grammar
approximation' is reasonable since the orbits that visit outside
the borders set by \cycle{001} are very unstable, and hence, less
important for the description of invariant dynamics.

The binary grammar with only rule that forbids repeats of one of the symbols is
known as the `golden mean' shift,\rf{DasBuch} because it has a topological entropy of
$\ln \left(\left(1 + \sqrt{5}\right)/2\right)$. Binary itineraries of golden mean cycles can be easily
obtained from the complete binary symbolic dynamics by substitution
$0 \rightarrow 01$ in  the latter. Thus, we can write the dynamical zeta
function for the golden mean pruned symbolic dynamics by replacing $0$s in
\refeq{e-CycleExpansion} by $01$:
\bea
1 / \zeta &=& 1 - t_{01} - t_1 - (t_{011} - t_{01} t_1 )
              \label{e-GoldenMeanCycleExpansion}\\
        && - [(t_{0111} - t_{011}t_1) + (t_{01011} - t_{01} t_{011} ) ] - ...
          \nonumber
\eea
Note that all the contributions longer than topological length $2$ to the
golden mean dynamical zeta function are in form of shadowing combinations. In \refsect{s-NumResults},
we will compare the convergence of the cycle averages with and without the
finite grammar approximation, but before moving on to numerical results,
we explain the remaining details of computation.

While dynamical zeta functions are useful for investigating the convergence
properties, they are not exact, and their computational cost is same as that of
exact spectral determinants. For this reason, we expand the
spectral determinant \refeq{e-SpectralDeterminant} ordered in the topological
length of cycles and pseudocycles. We start with the following form of the
spectral determinant \refeq{e-SpectralDeterminant}
\beq
    \det (s-\Aop) =   \prod_p \exp \left( - \sum_{r=1}^{n_p r < N}
                             \frac{1}{r} \frac{e^{r(\beta \Obser_p - s \period{p})}
                                          }{\oneMinJ{r}} z^{n_p r} \right) \, ,
\ee{e-SpectralDeterminantExp}
where the sum over the prime cycles in the exponential becomes a
product. We also inserted the order tracking term $z$ and truncated the sum over cycle
repeats at the expansion order $N$. For each prime cycle, we compute the sum in
\refeq{e-SpectralDeterminantExp} and expand the exponential up to order
$N$. We then multiply this expansion with the contributions from previous cycles
and drop terms with order greater than $N$. This way, after setting $z=1$,
we obtain the spectral determinant truncated to cycles and pseudo-cycles of
topological length up to $\cl{p} \leq N$,
\beq
    F_N(\beta , s) = 1 - \sum_{n=1}^{N} Q_n(\beta,s) \,. %Switched arguments of Q_n for symmetry with F_N
    \label{e-NthOrderSpectDet}
\eeq
In what follows, we shall drop the subscript, $F_N \to F$, but actual
calculations are always done for a range of finite truncation lengths
$N$. Remember that we are searching for the eigenvalues $s(\beta)$ of the
operator \Aop\ in order to compute the moments \refeq{eq-moments}.
These eigenvalues are located at the zeros of the spectral
determinant, hence as  function of $\beta$ they satisfy the implicit
equation
\beq
    F(\beta, s(\beta )) = 0 \, .
    \label{e-FNimplicit}
\eeq
By taking derivative of \refeq{e-FNimplicit} with respect to $\beta$ and
applying chain rule we obtain
\beq
    \frac{d s}{d \beta} = - \left. \frac{\partial F}{\partial \beta} \right/
                                     \frac{\partial F}{\partial s}\, .
\eeq
Higher order derivatives can be evaluated similarly.
Defining
\bea
    \langle \Obser \rangle &=& -  \partial F / \partial \beta
                               \continue %\label{e-ObserAvg} \\
   \langle \period{}\,\rangle &=&  \partial F / \partial s
                    \,,\qquad
   \langle \period{}^2 \rangle \,=\,  \partial^2 F / \partial s^2
                        \label{eq-Tavg} \\
    \langle \Obser^2 \rangle &=&
      -  \partial^2 F / \partial \beta^2
          %\label{e-Obser2Avg} \\
                    \,,\qquad
    \langle \Obser \period{}\,\rangle \,=\,
         \partial^2 F / \partial \beta \partial s
          \, ,\nnu % \label{e-ObserTAvg}
\eea
we write the \cycForm s as
\bea
    \langle \obser \rangle
            &=& \langle \Obser \rangle / \langle \period{}\,\rangle \, ,
    \label{e-Avga} \\
   \Delta
            &=&
           \frac{ 1~}{\langle \period{}\,\rangle}\left(
          \langle \Obser^2 \rangle
           \,-\, 2 \frac{d s}{d\beta}
              \langle \Obser \period{}\,\rangle
           \,+\, \left(\frac{d s}{d \beta}\right)^2
                       \langle \period{}^2 \rangle \right)
           \continue
           &=&
   \frac{ 1~}{\langle \period{}\,\rangle}
   \langle (\Obser-\period{}\,\langle  \obser \rangle)^2 \rangle
    \label{e-Avgsigma} \, ,
\eea
with everything evaluated at $\beta=0$, $s=s (0)$.

By probability conservation, we expect that for an invariant measure
$\rho_0(\ssp)$, the eigenvalue $s(0)$ is $0$. However, we did not make
this substitution in \cycForm s since, in practice, our approximations to
the spectral determinant are always based on a finite number of \po s, so
that the solution of $F_N(0, s(0)) = 0$ is small, but not exactly $0$. This
eigenvalue has a special meaning: It indicates how well the \po s cover
the strange attractor. Following this interpretation, we define $\gamma =
- s(0)$ as the `escape rate': the rate at which the dynamics escape the
region that is covered by the \po s. Specifically, for our finite grammar
approximation; the escape rate tells us how frequently the ergodic flow
visits the part of the Poincar\'e map that we cut off by applying our
finite grammar approximation.

We defined $\langle T \rangle$ in \refeq{eq-Tavg} as a shorthand for a partial
derivative, however, we can also develop an interpretation for it by looking
at the definitions of the dynamical zeta function \refeq{e-DynamicalZeta} and the
spectral determinant \refeq{e-SpectralDeterminant}. In both series, the partial
derivative with respect to $s$ turns them into a sum weighted by the cycle
periods; with this intuition, we define $\langle T \rangle$ as the `mean cycle
period'.

These remarks conclude our review of periodic orbit theory and its
extension to the equivariant dynamical systems. We are now ready to present
our numerical results and discuss their quality.

\subsection{Numerical results}
\label{s-NumResults}

We constructed the spectral determinant \refeq{e-NthOrderSpectDet} to different
orders for two observables: phase velocity $\dot{\theta}$ and the leading
Lyapunov exponent. Remember that $\Obser_p$ appearing in
\refeq{e-SpectralDeterminantExp} is the integrated observable, so in order to
obtain the moments of phase velocity and the leading Lyapunov exponent from
\refeq{e-Avga} and \refeq{e-Avgsigma}, we respectively put in
$\Obser_p = \theta_p$, the phase shift of the prime cycle $p$, and
$\Obser_p = \ln |\Lambda_{p,e}|$, the logarithm of its expanding Floquet
multiplier of $\Lambda_{p,e}$.

In \refsect{s:visual}, we explained that \SOn{2} phase shifts correspond
to the drifts in configuration space. We define the
corresponding diffusion coefficient as
\beq
    D = \frac{1}{2 d}         \lim_{\zeit \rightarrow \infty} \frac{1}{\zeit}
        \langle \theta(\zeit)^2 - \langle \theta(\zeit) \rangle^2 \rangle
%      = \frac{\Delta_{\theta}}{2}
\,,
\eeq
where $d=1$ since the configuration space is one dimensional.
\PC{2014-11-11: this looks wrong. The full \statesp\ is 4\dmn,
this diffusion seems to be defined relative to \slice, which is 3\dmn.
Have to seriously rethink this - you might well be right.}

\reftabs{t-DynamicalAverages}{t-DynamicalAveragesNoGrammar} respectively show
the cycle averages of the escape rate $\gamma$, mean period
$\langle T \rangle$, leading Lyapunov exponent $\Lyap$, mean phase velocity
$\langle \dot{\theta} \rangle$ and the diffusion coefficient $D$
with and without the finite grammar approximation. In the latter, we input
all the \rpo s we have found into the expansion
\refeq{e-SpectralDeterminantExp}, whereas in the former, we discarded the
cycles with symbol sequence `00'.
\begin{table}
    \caption{Cycle expansion estimates for
             the escape rate $\gamma$, average cycle period $\langle T \rangle$,
             Lyapunov exponent $\lambda$, average phase velocity
             $\langle \dot{\theta} \rangle$, and the diffusion coefficient $D$,
             using cycles up to length $N$ in the golden mean approximation
             \refeq{e-GoldenMeanCycleExpansion} of the symbolic dynamics.}
    \begin{tabular}{c|c|c|c|c|c}
     $N$ & $\gamma$ & $\langle T \rangle$ & $\lambda$ & $\langle \dot{\theta} \rangle$ & $D$ \\
    \hline
    1 & 0.249829963 & 3.6415122 & 0.10834917 & 0.0222352 & 0.000000 \\
    2 & -0.011597609 & 5.8967605 & 0.10302891 & -0.1391709 & 0.143470 \\
    3 & 0.027446312 & 4.7271381 & 0.11849761 & -0.1414933 &  0.168658 \\
    4 & -0.004455525 & 6.2386572 & 0.10631066 & -0.2141194 &  0.152201 \\
    5 & 0.000681027 & 5.8967424 & 0.11842700 & -0.2120545 &  0.164757 \\
    6 & 0.000684898 & 5.8968762 & 0.11820050 & -0.1986756 &  0.157124 \\
    7 & 0.000630426 & 5.9031596 & 0.11835159 & -0.1997353 &  0.157345 \\
    8 & 0.000714870 & 5.8918832 & 0.11827581 & -0.1982025 &  0.156001 \\
    9 & 0.000728657 & 5.8897511 & 0.11826873 & -0.1982254 &  0.156091 \\
    10 & 0.000728070 & 5.8898549 & 0.11826788 & -0.1982568 &  0.156217 \\
    11 & 0.000727891 & 5.8898903 & 0.11826778 & -0.1982561 &  0.156218 \\
    12 & 0.000727889 & 5.8898908 & 0.11826780 & -0.1982563 &  0.156220 \\
    \end{tabular}
    \label{t-DynamicalAverages}
\end{table}

\begin{table}
    \caption{Cycle expansion estimates of the escape rate $\gamma$, average
    cycle period $\langle T \rangle$, Lyapunov exponent $\lambda$, average
    phase velocity $\langle \dot{\theta} \rangle$, and the diffusion coefficient
    $D$ using all cycles found up to length $N$ .}
    \label{t-DynamicalAveragesNoGrammar}
    \begin{tabular}{c|c|c|c|c|c}
     $N$ & $\gamma$ & $\langle T \rangle$ & $\lambda$ & $\langle \dot{\theta} \rangle$ & $D$ \\
    \hline
    1 & 0.249829963 & 3.6415122 & 0.10834917 & 0.0222352 &  0.000000 \\
    2 & -0.011597609 & 5.8967605 & 0.10302891 & -0.1391709 &  0.143470 \\
    3 & 0.022614694 & 4.8899587 & 0.13055574 & -0.1594782 &  0.190922 \\
    4 & -0.006065601 & 6.2482261 & 0.11086469 & -0.2191881 &  0.157668 \\
    5 & 0.000912644 & 5.7771642 & 0.11812034 & -0.2128347 &  0.168337 \\
    6 & 0.000262099 & 5.8364534 & 0.11948918 & -0.2007615 &  0.160662 \\
    7 & 0.000017707 & 5.8638210 & 0.12058951 & -0.2021046 &  0.160364 \\
    8 & 0.000113284 & 5.8511045 & 0.12028459 & -0.2006143 &  0.159233 \\
    9 & 0.000064082 & 5.8587350 & 0.12045664 & -0.2006756 &  0.158234 \\
    10 & 0.000093124 & 5.8536181 & 0.12035185 & -0.2007018 &  0.158811 \\
    11 & 0.000153085 & 5.8417694 & 0.12014700 & -0.2004520 &  0.158255 \\
    12 & 0.000135887 & 5.8455331 & 0.12019940 & -0.2005299 &  0.158465 \\
    \end{tabular}
\end{table}
In \refsect{s-CycExp}, we motivated the finite grammar approximation by
claiming that it would lead to faster convergence of dynamical averages 
due to the nearly exact shadowing combinations of the golden mean zeta function
\refeq{e-GoldenMeanCycleExpansion}. This claim is supported by the data
in \reftabs{t-DynamicalAverages}{t-DynamicalAveragesNoGrammar}. Take, for example, the
Lyapunov exponent. This converges to $7$ digits for the $12^{\mathrm{th}}$ order
expansion when using the finite grammar approximation in
\reftab{t-DynamicalAverages}, but only converges to $4$ digits at this order in
\reftab{t-DynamicalAveragesNoGrammar}. Other observables compare similarly in
terms of their convergence in both cases. Note, however, that the escape rate
in \reftab{t-DynamicalAverages} converges to $\gamma = 0.000727889$, whereas
in \reftab{t-DynamicalAveragesNoGrammar} it gets smaller and smaller with an
oscillatory behavior. This is due to the fact that in the finite grammar
approximation, we threw out the part of attractor that corresponds to the
cusp of the return map in \reffig{fig:psectandretmap}\,(d) above the point cut 
by \cycle{001}.

In order to compare with the cycle averages, we numerically estimated the
leading Lyapunov exponent of the \twomode\ system using the method of
Wolf \etal\rf{WolfSwift85} This procedure was repeated 100 times for
different initial conditions, yielding a numerical mean estimate of
$\timeAver{\Lyap} = 0.1198 \pm 0.0008$. While the finite grammar
estimate $\Lyap_{FG} = 0.1183$ is within $0.6\%$ range of this value,
the full cycle expansion agrees with the numerical estimate. This is not
surprising since in the finite grammar approximation, we discard the
most unstable cycles to obtain faster convergence, and so can expect 
a slight underestimate of the Lyapunov exponent.


\renewcommand{\zeit}{\ensuremath{\tau}}  %time variable
