% siminos/blog/atlasRev.tex
% $Author$ $Date$

\section{Atlas paper revisions}
\label{sect:atlasRev}

{\color{red} enter here the internal discussion of referee comments}

\begin{description}

\item[2012-07-01 Predrag] Mein Gott. Will proud ignorance of fellow
plumbers ever end?

\end{description}

\subsection{Response 1.1}
\label{sect:Response1.1}

\begin{description}

\item[2012-07-01 Predrag] Not sure what to do with this report -
there are different paths to wisdom, and we cannot
write a different paper. Our point is that symmetry reduction is
necessary, so downplaying it is not what we want to do...

\bigskip

Referee: ``This paper is mainly a review of some techniques for symmetry reduction
that are useful for high-dimensional systems. In particular, the authors
present a method of constraining to a subspace called a "slice," or a set
of such subspaces. The method itself is not new, having been used in the
authors' references 3-6, but this paper presents a nice overview and
several interesting examples.''

\item[2012-07-16 Evangelos]
	Regarding novelty, I think that the idea of bringing together
	charts into an atlas has been proposed many times (including our
	own papers) but has not been implemented elsewhere. Should we
	emphasize this in the text?

{\bf [2012-07-16 Predrag]} to Evangelos: I have answered this question
several times in the blog above, but again, we presumably know the
literature, so
(1) can you give me one or more external references in which the
algorithm of constructing an atlas for the symmetry-reduced \statesp\ is
described?
(2) can you give me one or more external references in which the
$(d\!-\!2)$\dmn\ \emph{\poincBord} ${\cal S} \subset \PoincS$ is defined,
discussed, computed, explained?
(3) can you give me one or more references authored by us (or anyone
else) in which the \emph{ridge} is defined and constructed? I.e., a pair
of \KCedit{$(d\!-\!N)$\dmn\ } local \slice\ hyperplanes intersects in a
{ridge}, a \KCedit{$(d\!-\!N-\!1)$\dmn\ } hyperplane {\PoincS} of points
$\sspRed^*$ shared by a pair of charts and thus satisfying the \slice\
condition \refeq{PCsectQ0}.
{\bf [2012-07-17 Evangelos]} No, I can't and
that's exactly my point. This is all new but we do not emphasize it
properly, so it gives the impression of it all being review material.
I've slightly rephrased a sentence in the introduction to reflect novelty,
but I think we need to do more than this - emphasize it in the abstract,
conclusions and maybe introductory paragraph.
{\bf [2012-07-18 Predrag]} Now inserted bold novelty claims into the paper
and the response to the referees.

{\bf [2012-07-20 Evangelos]} Changed algorithm to prescription. As with
Poincar\'e sections, finding good templates cannot be done algoritmically,
it needs thinking.

Also, I would remove, if possible, credit to Chaosbook.org for figures -
people think that if this is already in a book, then it has to be review
material.
 {\bf [2012-07-17 Predrag]} Done.


\item[(0)  x ] 2012-07-01 Predrag:

\item[(3) |?|] 2012-07-17 Predrag:

\item[(4) |?|] 2012-07-01 Predrag:

\item[(4) |x|] 2012-07-01 Predrag:

\item[minor (1) |x|] {\bf [2012-07-18 Daniel]:} Removed all references to
61,506 dimensions and replaced it with D dimensions, specifying that D is
large or just saying ``high-dimensional'' wherever necessary. Also took
d and D which were used in the text for dimensionality and replaced them
all with D's. Wrote comment outlining this change in \texttt{response1.1.txt}.
{\bf [2012-07-18 Predrag]}
ChaosBook.org convention is to use $d$ for \statesp\ dimension, $D$ for
configuration space dimension; Navier-Stokes pipe, etc. Reverted to $d$
throughout.

\item[minor (3) |x|] {\bf [2012-07-17 Daniel]:} Modified the text to try
and clarify ``turbulence breaks all symmetries'' and explain that
solutions calculated in invariant subspaces have different dynamics and
may have nothing to do with turbulence. Included review of changes in
\texttt{response1.1.txt}.

\item[minor 6. p9 |x|] 2012-07-01 Predrag:
 referee does not understand the difference between the `method of
 co-moving frames' and the `method of connections',
 \reffig{fig:BeThMconnect}. Shouldn't we weave Evangelos clarification
 into the text proper? {\bf [2012-07-17 Evangelos]} Done.

\item[minor (9) |?|] 2012-07-01 Predrag: can you weave this into the text?: ``
\edit{
Blah, blah}
''

\item[minor (11) |?|] 2012-07-01 Predrag:


\item[2012-07-19 Evangelos] Referee says: I described the manuscript as
self-focussed because about one third of the references are to the
authors' work.

We respond:
We feel referee's pain, having recently refereed a paper with 12
self-references out of 44 total, many of them deliriously misplacing the
credit.
*   Now there are 9 self-references (from the list of 44) that authors
    deem essential to this paper. By comparison, our internal
    symmetry-reduction blog cites 306 references.

ES: ``having recently refereed a paper'' means that we (\ie\ Predrag)
refereed one such paper or that the referee refereed our paper?

\item[2012-07-19 Evangelos] The ref. says:

I do not understand the remarks about sections and strobing, at the end
of section II: isn't it obvious that a Poincare section is not a
projection? and is it not obvious that strobing is a Poincare section
with the section being defined by time taking certain values? I.e. in an
extended state space the two are equivalent, aren't they?

RESPONSE: Not sure what an `extended' state space is.
(1) Our students often internally visualize a (d-1)-dimensional Poincare
section as one views a 3 dimensional world projected onto a 2-dimensional
photo (from one photo one cannot reconstruct the 3 dimensions), so that
remark was meant for them.
(2) suppose you strobe a periodic orbit of period $\pi$ every second.
({\bf ES:} What does every second mean here?)
What you get is a set of points that erotically fills out a
1-dimensional loop, whereas *stationary* Poincare section is
0-dimensional.

ES: Regarding remark (2), I think the referee has in mind systems with
explicit time dependence where one usually considers time as an extra
depended variable and defines a stroboscopic Poincar\'e section.

Or maybe what he has in mind is that you could think of the stroboscopic
condition as defining implicitly a manifold in phase space, which you
then call a Poincar\'e section. But then one really needs to carefully
choose the strobing period in order to get something meaningful in the
case of dissipative chaotic or turbulent system (as Predrag's example
shows).

\item[2012-07-19 Predrag]  I now realize that `strobing' remarks confuse
rather than aid the reader, and have removed them from the text.

\item[2012-07-19 Evangelos] Removed: ``the bit rushed submission that
reviewers have received (graduate students firmly believe that papers are
written one day before the deadline)'' from response. Do you really want
to trust an anonymous reviewer with behind the scenes details?

\end{description}


\subsection{Response 1.2}
\label{sect:Response1.2}

\begin{description}

\item[2012-07-14 Predrag] In response to referee's plaint:
``I described the manuscript as self-focussed because about one third of
the references are to the authors work.'' I removed a bunch of
Cvitanovi\'ciania, including Froehlich and Cvitanovi\'c (2011) paper, with
a heavy heart - Stefan did this work as undergraduate...

{\bf [2012-07-16 Evangelos]} I think you should not remove this. This is were
		slicing is formulated as an extremum condition. Also,
		I think the reviewer is pissed off about not citing (his)
		fluid dynamics papers - he does not care about this one.

{\bf [2012-07-16 Predrag]} thanks - Froelich is back.


\end{description}

\section{Proofs}

\begin{description}

\item[2012-10-05 Predrag to Daniel, Evangelos] Apparently we have 48
hours starting Fri, Oct 5, 2012 at 8:21am. I'll submit the annotated; can
you guys go through

\texttt{siminos/atlas/Chaos-v2/proofCopy.pdf}

\noindent
and enter here (not the pdf file - I have a special `active' copy) in
this format
\\
(line \#) correction





From cha@aip.org Fri, Oct 5, 2012

proofs fare now available for your review.

The publisher will take no further action until you review page proofs
and notify the production staff of any necessary changes.

\noindent
Article Title: CARTOGRAPHY OF HIGH-DIMENSIONAL FLOWS:
               A VISUAL GUIDE TO SECTIONS AND SLICES\\
Manuscript Code \#: 12259CATR\\
AIP ID: 009295CHA\\
Corresponding Author: PREDRAG CVITANOVIC\\
Journal: Chaos\\
Section: FOCUS ISSUE: FIFTY YEARS OF CHAOS: APPLIED AND THEORETICAL\\

http://aipprod.aip.org/ap2/proof?id=009295CHA-26458

All proof corrections must be indicated, by you, in English, as
annotations in the PDF file.

To annotate the PDF file with any changes, follow these steps:
      * Open the PDF file, then choose
        Tools > Comment & Markup > Callout Tool.
        If you are using Acrobat Reader X, use the Sticky Note tool.
      * You should now be able to place your cursor anywhere within the
        PDF file and note any changes needed.
      * Read your proofs thoroughly and check all text, illustrations,
        captions, equations, and tables.
      * Indicate all corrections as annotations within the PDF file. Make
        sure you include all of your changes on the annotated proof that
        you return to us. Do not send us a revised manuscript, in lieu of
        an annotated PDF.
      * Multiple rounds of proof corrections will not be accepted; the
        production staff cannot accommodate corrections communicated by you
        separately.
      * The extent to which you have altered your original article will
        be evaluated. If it is determined that your corrections are
        extensive (new material, not copy editing errors), your article
        will be returned to the Editorial Office and this will delay
        publication.
      * Do not use TeX formatting in your annotations on the PDF proofs.

      * Please check each illustration and provide any feedback as
        annotations within the PDF file.
      * The appearance & quality of illustrations in the proofs directly
        reflect how they will appear in the final published paper online.
        Note that graphics may have been down-sampled to produce a more
        manageable file size for downloading by online users of the
        journal.  Actual viewing and output results will vary depending
        upon the reader's screen or output device resolution.  (Note that
        high-resolution versions of the graphics will be used to produce
        the printed version of the journal.)
      * If illustrations appear in color in these proofs, then they will
        appear in color in the online journal.
      * If, and only if, you have agreed to the color-figure charges,
        color illustrations will appear in color in the printed journal.
      * If you expected one or more of your figures to appear in color,
        but they appear only in grayscale in these PDF proofs, please add a
        note to the PDF.

      * Save the annotated PDF file.

RETURNING CORRECTIONS

      AIP's eProof
      website:  approve without corrections,
      upload a marked PDF file, and upload new figure files as necessary.

      URL: http://pubpartner.aip.org/aipportal/proof_login.jsp
      PIN -> 009295CHA-26458

\item[2012-10-05 Daniel] 
AIP COMMENT: Please provide the definition for “PDE.”

RESPONSE: Replace ``the construction of invariant, PDE discretization independent state space coordinates,5'' with ``the construction of invariant state space coordinates (independent of the discretization scheme chosen to solve the relevant partial differential equation)'' or just ``the construction of invariant state space coordinates''. PDE doesn't appear anywhere else in the paper, so neglecting the independent of PDE discretization part probably doesn't hurt the paper too much and is in Ref. 5. 
\\
\\
AIP COMMENT: Please update Ref. 10 with volume number, page number, and year if published.

RESPONSE: Not sure where R. R. Kerswell, ``Misunderstanding the turbulence in a pipe,'' (unpublished) came from. Is this some document that somebody has actually seen? It's not a preprint that I can find online. Somebody update it if possible. I emailed Rich about it, just in case...
\\
\\
AIP COMMENT: Please update Ref. 11 with volume number, page number, and year if published.

RESPONSE: Has A. P. Willis, P. Cvitanovic, and M. Avila, ``Revealing the state space of turbulent pipe flow by symmetry reduction,'' J. Fluid Mech. (submitted); 734 e-print arXiv:1203.3701. been accepted yet? Don't see anything in this regard on arxiv.org nor the JFM website. Predrag, please update the reference if you have any inside information. Contacted Ashley and Marc in case they know something but haven't told Predrag.
\\
\\
AIP COMMENT: Please provide the publisher detail in Ref. 14.

RESPONSE: I'm not sure what the official JFM citation format for websites is but assuming that it is something like the IEEE format (Authors. ``Title.'' Internet: www.chaosbook.org/tutorials, date published [date accessed]), the reference for chaosbook/tutorials is 
J. F. Gibson and P. Cvitanovi{\'c}. ``Movies of Plane Couette Flow.'' Internet: www.chaosbook.org/tutorials, 2012 [Oct. 5, 2012].
\\
\\
AIP COMMENT: Please update Ref. 43 with volume number, page number, journal title, and year if published.

RESPONSE: N. Vandersickel and D. Zwanziger, ``The Gribov problem and QCD dynamics,'' \emph{Physics Reports} or \emph{Phys. Rep.},  in press; 
e-print http://www.sciencedirect.com/science/article/pii/S0370157312002074 or maybe leave original arxiv link for eprint.
\\
\\
FURTHER COMPLAINTS: \\
(1) Fig. 4a got miniturized (in a pretty bad way!)! \\
(2) Fig. 7a got miniturized (in a pretty bad way!)! \\
(3) List of steps in slicing method of section V. CHARTING THE SLICE got indented weird or at least different from submitted version. I think they look better the way we submitted them originally.\\
(4) Figures 6, 8, 9, 10, and 11 got blown up (not horribly) but they looked better the way we submitted them.\\
(5) Figure 12 is slightly smaller than submitted.\\
(6) Captions for Figures 9 and 10 have weird spacing between lines when a line has a symbol with a superscript.\\
\\
There's probably more but it's 3 AM. I'll go through it with a fine toothed comb tomorrow morning.

Evangelos COMPLAINTS: 

(line 41) space s -> spaces
(line 115) space s -> spaces
(line 245) by Siminos^{,17} -> in Ref. 17
(line 301) solutions -> solution
(line 396) z=-b z -> \dot{z}=-bz
(line 427) Eq. (1) -> (1)
(line 485) prime should be superscripted
(line 525) space s -> spaces
(caption of Fig. 6, second line) primes should be  superscripted


\end{description}