% siminos/blog/bronsky-2005.tex
% $Author$ $Date$

\section{Uncertainty estimates and $L_2$ bounds for the KS}

% Predrag moved from thesis/chapters/ back to blog  oct  2 2009
% Predrag moved from blog/ to thesis/chapters/      jun 26 2008
{\bf
Jared C. Bronski
%\footnote{Department of Mathematics,
% University of Illinois Urbana-Champaign, 1409 W. Green St, Urbana IL 61801}\\
and
Tom Gambill}


Nicolaenko, Scheurer and Temam\rf{NSTks85} gave the first long-time
boundedness result for the Kuramoto-Sivashinsky equation, showing that
$\limsup_{t \rightarrow \infty} |\!|u|\!|_2 \le C L^{\frac{5}{2}}$ for
odd initial data, as well as showing that bounds on the $L_2$ norm imply
bounds on the dimension of the attractor.

The $L_2$ estimate was improved by Collet, Eckmann, Epstein and
Stubbe\rf{CEEksgl93} who  extended it to any mean-zero initial data and
improved the exponent from $\frac{5}{2}$ to $\frac{8}{5}$, and by
Goodman~rf~{Good}, who extended it to any mean-zero initial data with the
same exponent.

these papers that the function $|\!|u-\phi|\!|_2^2$ is a Lyapunov
function for an appropriately chosen $\phi$ and $|\!|u|\!|_2$
sufficiently large.

bounds which do not fit into Lyapunov function framework:
Ilyashenko~rf~{Ilyashenko}, Otto  and Giacomelli/rf{GiacoOtto05}. The
latter, which treats the KS equation as a perturbation of the Burger's
equation, is currently the best estimate, establishing that
\[
\limsup_{t \rightarrow \infty} |\!|u|\!|_2 = o(L^{\frac{3}{2}}).
\]

In this paper we give an elementary argument of the Lyapunov function
type which establishes the slightly weaker result
\[
\limsup_{t \rightarrow \infty} |\!|u|\!|_2 = O(L^{\frac{3}{2}}).
\]
Our proof applies equally to the destabilized Kuramoto-Sivashinsky (dKS)
equation:
\[
u_t = -u_{xxxx} - u_{xx} + \gamma  u + u u_x ~~~~~~\gamma > 0 ~~~~~ \int u(x,0)dx = 0.
\]

It was shown by Wittenberg~rf~{WITTENBERG} that this equation
has stationary solutions which satisfy $|\!|u|\!| \propto L^{\frac{3}{2}}.$
Since a Lyapunov function argument for the KS equation also applies
to the dKS equation (for sufficiently small $\gamma$) Wittenberg
argued that $\frac{3}{2}$ is the best exponent that one can expect
from the Lyapunov function approach.

Similar ideas of considerably greater generality have been used by
Constantin and Doering to establish bounds on energy dissipation in
fluids, and generally go by the name `background flow
method.'~rf~{CD1,CD2}

%
{\bf PC}:
 % Past: Michael Loss has not taught us how to bound $E$ by
 %  Sobolev bounds. Neither has Spiegel. Next: But
Constantin says that the answer is in \refrefs{temam85ks}. Eckmann says:
the  best bound is by Otto\rf{GiacoOtto05}; the only bound close to k=0,
better in essential way. See also \refref{bronski2005}. Eckmann had
$L^{8/5}$, but conceptually Otto is the best. Recheck whether it is $|u|$
or $E \propto L^{3/2}$. When the solution is big, how long can it stay
big? They found it cannot stay big for long.

from bounds on energy, we might be able to bound the number of \eqva\ as
function of systems size $L$, and thus be sure we have them all.

Next for you guys: read Lieb and Ruelle to learn how to bound $E$  by
Sobolev bounds.

The basic strategy is to choose a periodic function $\phi_x$ of zero mean
such that the following quadratic form is coercive,
\[
<\!\!u, K u\!\!>= \int u_{xx}^2 - u_x^2 + \phi_x u^2 \ge \delta |\!|u|\!|^2,
\]
for $u$ satisfying Dirichlet boundary conditions and some positive
$\delta$ independent of $L$.

This calculation is, in a sense, complementary to Lieb-Thirring type
inequalities. In Lieb-Thirring inequalities one attempts to maximize some
measure of the negative part of the spectrum of an operator over all
potentials with a fixed norm.

M.S. Jolly, I.G. Kevrekidis and E.S. Titi,
Approximate inertial manifolds for the {Kuramoto-Sivashinsky} equation:
analysis and computations,
Physica D 44 (1990), 38-60.

D. Michelson, Elementary particles as solutions of the Sivashinsky
equations, Physica D 44 (1990), 502-556.

E. Mitidieri and S.I. Pohozaev,
Apriori Estimates and Blow-up of Solutions to Nonlinear Partial
Differential Equations and Inequalities,
Procedings of the Steklov Institute of Mathematics Issue 3 vol 234
(2001).

Y. Cao and E.S. Titi

Y. Pomeau and P. Manneville,
Stability and fluctuations of spatially periodic flow,
J. Physique Lett. 40 (1979), 609-612.

P. Souplet,
Gradient blow-up for multidimensional nonlinear parabolic equations with
general boundary conditions,
Di. Inegral Eqns 15 (2002), 237-56.

E. Tadmor, The well-posedness of the {Kuramoto-Sivashinsky} equation,
SIAM Journal on Mathematical analysis 17 (1986), 884-893.


article{CD1} Constantin, P. and Doering, C. (1992).
Energy dissipation in shear driven turbulence.
{\it Phys. Rev. Lett.} {\bf 69} (11), 1648-1651.

article{CD2} Constantin, P. and Doering, C. (1995).
Variational Bounds for Dissipative Systems.
{\it Phys. D}, {\bf 82} (3), 221-228.

P. Constantin, C. Foias, B.  and R. Temam\rf{constantin_integral_1989}.

article{Foias}
Foias, C., Manley, O. and Temam, R. (1987).
Attractors for the {B\'{e}nard} problem:
{Existence} and physical bounds on their fractal dimensions.
{\it Nonlinear Anal.}
{\bf 11}, ~939-967.

article{Foias2}
Foias, C., Sell, G. R., and  Temam, R. (1988).
Inertial manifolds for nonlinear evolutionary equations.
{\it J. Diff. Eq.}
{\bf 73} (2), ~309-353.

C. Foias and I. Kukavica, Determing nodes for the {Kuramoto-Sivashinsky} equation,
J. Dynam. Di. Eq. 7 (1995), 365-373.

article{Good}
Goodman, J. (1994).
Stability of the {Kuramoto-Sivashinsky} and related systems.
{\it Comm. Pure Appl. Math.}
{\bf 47} (3), ~293-306.

article{Ilyashenko}
Il\'{}yashenko, Yu. S. (1992).
Global analysis of the phase portrait for the {Kuramoto-Sivashinsky} equation.
{\it J. Dynam. Diff. Eq.}
{\bf 4}  (4), ~585-615.

article{LMRT} LaQuey, R., Mahajan, S., Rutherford, P. and Tang, W. (1975).
{\it Phys. Rev. Lett} {\bf 34} (7) 391-394.

article{ST} Sell, G. and Taboada, M. (1992).
 Local dissipativity and attractors for the {Kuramoto-Sivashinsky}
equation in thin $2{\rm D}$ domains.
{\it Nonlinear Analysis} {\bf 18} (7), 671-687.

article{WITTENBERG}
Wittenberg, R. W. (2002).
Dissipativity, analyticity and viscous shocks in the (de)stabilized
{Kuramoto-Sivashinsky} equation.
{\it Phys. Lett. A}
{\bf 300} (4-5), 407-416.
