% siminos/blog/flotsam.tex
% $Author$ $Date$
%
% Predrag moved from thesis/chapters/ back to blog  oct  2 2009
% Predrag moved from blog/ to thesis/chapters/      jun 26 2008
% Predrag created file                              jun 20 2006




\chapter{Flotsam}

This chapter contains material which has not been included in papers
 in this git repository,  and/or
Ph.D. theses. Please clean up whatever seems
extraneous.

\section{Example dynamical systems used throughout this work}
    \label{s:exampleIntro}
% Chapter Introduction, section Example dynamical systems used throughout the thesis.
% $Author$ $Date$
% Label: s:exampleIntro

In this section we briefly introduce some dynamical systems
that will be used as simple examples to demonstrate various
concepts in later chapters.

\paragraph{Lorenz, refchap~{chap:Lorenz}:}
%
\ES{Perhaps say Lorenz flow, complex Lorenz flow etc.}
Lorenz introduced his celebrated equations\rf{lorenz} as a
severe truncation of the Navier-Stokes equations describing
Rayleigh-Benard flow. They read
\index{Lorenz equations}
\begin{align}
\dot{x} &= \sigma (y-x) \notag \\
\dot{y} &= \rho x - y - x z \\
\dot{z} &= x y - b z \notag
\end{align}
where $\sigma$ the Prandtl number, $\rho$ the Rayleigh number
and $b$ an aspect ratio of the problem.

\paragraph{{\CLf}, refchap~{chap:lasers}:}
%
Gibbon and McGuinness\rf{GibMcCLE82} introduced {\CLf}
\beq
\begin{split}
	\dot{x}_1 &= -\sigma x_1 + \sigma y_1\cont
	\dot{x}_2 &= -\sigma x_2 + \sigma y_2\cont
	\dot{y}_1 &= (r_1-z) x_1 - r_2 x_2 -y_1-e y_2 \cont
	\dot{y}_2 &= r_2 x_1 + (r_1-z) x_2 + e y_1- y_2\cont
	\dot{z} &= -b z + x_1 y_1 + x_2 y_2
	\label{eq:introCLeR}
\end{split}
\eeq
as a low-dimensional model of baroclinic instability in the
atmosphere. {\CLf} is equivariant under the action of
$\SOn{2}$.

 \paragraph{\KS, refchap~{chap:KSe}:}
 The \KS\ system
 \beq
   u_t = F(u) = -{\textstyle\frac{1}{2}}(u^2)_x-u_{xx}-u_{xxxx}
     \,,\qquad   x \in [-L/2,L/2]
 \ee{intro:ks}
 was introduced by
 Kuramoto and Tsuzuki\rf{ku} as a phase equation for
 reaction-diffusion systems described by Complex Ginzburg-Landau
 equation and independently by Sivashinsky\rf{siv} to describe
 instabilities in laminar flame fronts.


\section{Squirreled away for the next paper}
% from energy.tex

Rescuing back what we squirreled away for the thesis from
material omitted form the KS papers.

\begin{description}
\item[2006-08-20 Lan]
    I checked the equilibria for $L=22,\nu=1$.
\end{description}
