% reducesymm/blog/strategy.tex
% $Author$ $Date$

\chapter{Strategy, to write up}
% Predrag: this file is distinct from siminos/blog/strategy.tex
\begin{bartlett}{
Someone who makes the same mistake twice is not a wise man.
}
\bauthor{
An ancient Greek saying
    }
\end{bartlett}




\section{How to read me}

For those whose Freud needs brushing up:
`Desymmetrization and its discontents' is a pun
on \HREF{http://en.wikipedia.org/wiki/Civilization_and_Its_Discontents}
{Civilization and its discontents}.

Throughout:  {\footnotesize inCB} on the margin                 \inCB
indicates that the text has been transferred to an
article in siminos/*/,  or to ChaosBook.org
chapters, such as
\HREF{http://ChaosBook.org/chapters/continuous.pdf}
{continuous.pdf}.
 {\footnotesize 2CB} on the margin indicates that the text
still needs to be transferred an article or ChaosBook.org.      \toCB

This \texttt{blog.pdf} file is \emph{hyperlinked}.
There is a bunch of handy links throughout,
now that we went to the trouble of downloading papers and stealing books. Brilliant.
For example, if you click on
this: \arXiv{1103.4536}, you might find a interesting paper to read.
\HREF{http://chaosbook.org/library/KoSa11.pdf}{Clicking here} will
lead you to our internal ChaosBook.org/library:
You'll need to log in as \texttt{student} and then enter \texttt{Lautrup}.
\HREF{http://www.zotero.org/groups/cns}{Zotero.org} is great,
but as only Evangelos and Predrag use it right now,
it is faster to stick stuff into ChaosBook.org library.

Here is a novice's guide to desymmetrization bloggery:
\begin{itemize}
  \item
How to read the running blog: go first to the latest blog post, end
of \refchap{c-DailyBlog}.
  \item
If you are reading an article of common interest (which does not fit into
one of the specialized topics), it might be already in \refchap{c:lit};
in nay case, enter your notes at the end of \refchap{c-DailyBlog}.
  \item
Comments to ChaosBook.org go into \refchap{chap:ChaosBook} blog.
  \item
Periodic orbit theory comments belong to \refchap{chap:UPO} {\em
Periodic orbit theory}.
  \item
If Hamiltonian dynamics is your obsession, that's in
\refchap{sect:LiePolice} {\em Lie police}.
  \item
Slicing all things laser should be confined to
\refchap{chap:lasers} {\em Laser physics: The lingo}
  \item
Symmetry reduction in fluid dynamics is in \refchap{chap:fluids} {Fluids}
  \item
Geophysicists reside in
\texttt{siminos/baroclinic/BrCv12.tex}.
  \item
Guys writing the ultimate guide to slicing for the woman on the street,
\texttt{siminos/atlas/}, blog in \refchap{chap:atlas} {\em Atlas}.
  \item
Plumbers who ponder how to slice experimental data blog in
\refchap{c-exp} {\em Symmetry reduction of experimental data}.
  \item
Enter your ponderings on all things norm into \refsect{c-norms}
\emph{Norms, distances}, though some of that is also in \refchap{c:lit}
(for experimental data) and \refchap{sect:LiePolice} (for symplectic
distances).
  \item
Cardiologists (mere electricians, really) have gotten a divorce, too. That
blog's gone to \texttt{DOGS/saldana/excite.tex}.
  \item
All things `{geometric phase}' are in \refchap{c-geometric} {\em
Geometric phase}.

\end{itemize}


\section{Classification, keywords}

						\noindent
elsevier (first number is Elsevier only? or? The 2nd is \textbf{PACS})
10.020: 02.20.-a Group theory	\\
10.050: 02.50.-r Probability theory, stochastic processes, and statistics 	\\
        02.70.Bf Finite-difference methods \\
10.150: 05.40.Ca Noise	\\
10.180: 05.45.-a Nonlinear dynamics and chaos	\\
10.190: 05.45.Ac Low-dimensional chaos	\\
10.210: 05.45.Gg Control of chaos, applications of chaos	\\
10.220: 05.45.Jn High-dimensional chaos	\\
10.230: 05.45.Mt Quantum chaos - semiclassical methods	\\
10.240: 05.45.Pq Numerical simulations of chaotic systems	\\
10.305: 05.10.Gg Stochastic analysis methods	\\
10.390: 05.70.Ln Nonequilibrium and irreversible thermodynamics	\\
20.030: 05.10.Gg Stochastic analysis methods	\\
20.080: 05.45.-a Nonlinear dynamics and nonlinear dynamical systems	\\
20.090: 05.45.Mt Semiclassical chaos (quantum chaos)	\\
        42.65.Sf Dynamics of nonlinear optical systems; optical instabilities,
                 optical chaos and complexity, and optical spatio-temporal dynamics \\
60.090: 46.70.-p Application of continuum mechanics to structures	\\
        47.10.Fg 	Dynamical systems methods (in Fluid Mechanics)	\\
        47.27.ed 	Dynamical systems approaches (turbulent flows)	\\
70.050: 47.27.-i Turbulent flows, convection, and heat transfer	\\
70.110: 47.52.+j Chaos (in fluid dynamics)	\\
70.130: 47.54.+r Pattern selection; pattern formation	\\
70.150: 47.60.+i Flows in ducts, channels, nozzles, and conduits	\\
70.160: 47.62.+q Flow control	\\
        83.60.Wc Flow instabilities \\
        95.10.Fh Chaotic dynamics



						\noindent
\textbf{keywords}	\\
symmetry reduction,	\\
equivariant dynamics,	\\
relative equilibria,	\\
relative periodic orbits,	\\
return maps,	\\
slices,	\\
moving frames,	\\
Hilbert polynomial bases,	\\
invariant polynomials,	\\
Lie groups	\\

\section{Zoteromania}

\begin{description}

\item[2008-07-18 Predrag] about webtools for generating BibTeX:
www.zotero.org
        will pick up most books from Amazon, etc; but
        better to find a book first on
\HREF{http://www.worldcat.org}{www.worldcat.org}
          or
\HREF{http://scholar.google.com}{scholar.google.com}, then zotero it
          in a collection, and export in BibTeX format

\item[2009-12-22 Evangelos]
setting up a cns group at zotero.org

\item[2011-08-16 Predrag] moved the instructions to siminos/bibtex/zotero.txt

\end{description}

\section{Git with it}

\begin{description}

\item[2013-07-07 Predrag] Divakar says I have to do it:
    svn is so 20th century now we must git it, so I'm taking a part of
    siminos repository, converting this
    \texttt{svn} repository to \texttt{git}, pruning most of it, and
starting a new theory graduate student Burak (Nazmi B. Budanur
<burakbudanur@gmail.com>) on it. One can continue working on the svn
repository using
    \HREF{http://mojodna.net/2009/02/24/my-work-git-workflow.html}
    {this web page}, but here my ambition is
    \HREF{http://thomasrast.ch/git/git-svn-conversion.html} {only one
    way},
    \HREF{http://git-scm.com/book/en/Git-and-Other-Systems-Migrating-to-Git}
    {to create} a git repository, and work from then onward only within
    \texttt{git}. Did this in linux rather than windows, seems easier.

The author map \texttt{users.txt} is a text file
\begin{verbatim}
 svn log svn://zero.physics.gatech.edu/siminos |
 sed -ne 's/^r[^|]*|
 \([^ ]*\) |.*$/\1 = \1 <\1@zero.physics.gatech.edu>/p' |
 sort -u > users.txt

\end{verbatim}

that maps SVN usernames
to real names and email addresses for the git
history, with lines of the form:
\begin{verbatim}
svnuser = R. E. Alname <real@email.example.com>
\end{verbatim}
The result is pretty useless so it easier to create it by copying from svn@zero
editing:
\begin{verbatim}
repos/siminos/hooks/emaildict
\end{verbatim}
The project \texttt{reducesymm} is branchless (consists only of a single line of history), so:
\begin{verbatim}
git svn clone -A users.txt --no-metadata \
svn://zero.physics.gatech.edu/siminos reducesymm
\end{verbatim}
add your new Git server as a remote and push to it.
Here is an example of adding github.com server as a remote:
\begin{verbatim}
git remote add origin \
https://user:'Password'@github.com/user/reducesymm.git
\end{verbatim}
To have all branches and tags go up, first update local rep (pull),
then push to the server
\begin{verbatim}
git pull origin --all
git push origin --all
git push origin --tags
\end{verbatim}
All branches and tags are now on the Git server in a nice,
clean import.

\item[2013-08-08 Burak]
I started the
\HREF{https://groups.google.com/forum/\#!forum/reducesymm} {Google group} for  commits.

On github project page
\HREF{https://github.com/cvitanov/reducesymm}
{github.com/cvitanov/reducesymm}, do the following:

Settings (last item on the side bar) -> Service Hooks -> E-mails

If you add,

\texttt{reducesymm@googlegroups.com}

don't check the box [send from author], checking the box [Active]

The google group should start to get notifications on commits and
distribute it to members.  You can check it by clicking on Test Hook on
this page.

This is based on
\HREF{http://www.dynamis-technologies.com/blog/send-github-commits-to-a-google-group/}
{this blog post}.
When someone joins
to project, the person should also be made a member of the
\HREF{https://groups.google.com/forum/\#!forum/reducesymm} {Google group}.
This should take 3 or 4 clicks, one copy/paste and a captcha.

\item[2013-08-10 Predrag]
Who's member ``noreply", joined Aug 7, 2013? Apparently Burak? That's not very informative...

\item[2013-08-10 Burak]
``noreply'' is github's account which sends the notification e-mails. I gave that address membership so that mails sent from it are sent
to all members of the group.

\item[2013-08-20 Predrag] On MS Windows, I do not find
\HREF{http://git-scm.com} {git console} particularly useful (I have
not needed to use the console yet). However,
\HREF{http://windows.github.com/} {GitHub for Windows} is very easy
to use, it might be only thing you need. What I like about it is that
it, as a default, displays the diff on all edited files - in tortoise
svn you have to open that manually.

\item[2013-08-23 Predrag] Now it happened; Daniel pushed his edits to gitHub
while I was editing the same file. Dealing with merges is funky. In linux:
\begin{verbatim}
   > git stash save
   > git pull origin master
   > git stash pop
edit the conflicted files by hand, by searching for `<<<<<<<', then
   > git commit -a
   > git push origin master
\end{verbatim}


\item[2013-08-10 Predrag] svn date stamps files by modifying the entries in
\begin{verbatim}
$Author$ $Date$
\end{verbatim}
Can git git something like that?

\item[2013-08-25 Burak] Here is a quick start guide for Debian-based Linux (Ubuntu, Ubuntu variations and some others such as ChrunchBang) users that I originally wrote for Kamal:

Open up a terminal on Ubuntu (or any debian based linux distro. If you have Fedora, replace all "apt-get"s below with "yum"s) and type this to get whole packages for compiling \texttt{.tex} files on the project:

\begin{verbatim}
sudo apt-get install texlive-full
\end{verbatim}

This should download around 1-2 gb of data. If you don't have a preferred latex editor, you might want to get Kile, it's simple:

\begin{verbatim}
sudo apt-get install kile
\end{verbatim}

Default Kile installation comes with some standard latex packages but they do not include REVTeX 4.1 which is needed to compile some of the .tex files in project. That is why I recommend installing texlive-full in the very beginning.

You also need git:

\begin{verbatim}
sudo apt-get install git-core
\end{verbatim}

This one should go fast. After you get git, you need to make a clone of the project (https://github.com/cvitanov/reducesymm) on your own side. Open a terminal in a folder that you would like to store your research and type:

\begin{verbatim}
git clone https://github.com/cvitanov/reducesymm.git
\end{verbatim}

This will create a folder named reducesymm and project files, blogs, etc. will be in it.

Basic stuff that you can do with git is explained well in this post. http://alistapart.com/article/get-started-with-git you might wanna read this.

To participate in this project, Predrag should add you as a collaborator on git. To get updates on your side, this is what you should do:

On the reducesymm/ folder, type

git remote add upstream https://github.com/cvitanov/reducesymm.git

This will tell git that the original of your project is here: https://github.com/cvitanov/reducesymm.git

So when someone pushes an update to the project (you will get updates once you join the google group), you can download that update by typing:

\begin{verbatim}
git fetch upstream
\end{verbatim}

When you do this, on your own computer, there will  be two branches of the project, master and upstream. You will probably want to get your own copy to be most up-to-date version, for that you should type in:

\begin{verbatim}
git merge upstream/master
\end{verbatim}

When Predrag makes you a collaborator, you will be able to push your commits to the project. This is how it goes:

Let's say that you modified one of the project files, it will be listed when you type

\begin{verbatim}
git status
\end{verbatim}

In the reducesymm/ folder. To commit that change, you type:

\begin{verbatim}
git commit -am 'your message'
\end{verbatim}

\item[2013-08-25 Predrag] I believe I had already installed all the
packages Burak mentions both on \texttt{sinux} and \texttt{zero}
machines; let me know if something is missing.

\item[2013-09-02 Evangelos]
My gitHub username: vasimos and I've used my evangelos.siminos@gmail.com email address.


\item[2013-09-02 Predrag to Evangelos] Done. You seem to have 'forked' the repo. Is that a good thing to do? Or it just means you have a copy of it
    on gitHub, rather than locally on your machine? I have no idea. I just 'cloned' it.

\item[2013-10-02 Predrag] Got into intractable mess with simultaneous commits with other
people - my repo forked 3 branches, erase 1 hour of my edits. Locked Burak in my office
with the laptop - $n$ hours later he managed to commit my edits, at least some of them.
How he did it, he will never tell.
I had removed most of siminos blog files, but they are back again, together with the old
\texttt{blog.tex} ... so who knows what
other edits got lost.

I checked for linux
\HREF{} {git GUIs} and installed these on Kimberley's sinux.physics.gatech.edu,
to play with:

\texttt{apt-get install gitk}
\\
     \texttt{giggle}
\\
     \texttt{gitg}
\\
    \texttt{git-cola}
 \\
    \texttt{git-gui}


\end{description}
