% siminos/CLE/05fixMe.tex
% $Author$ $Date$

\section*{To fix}

{\bf remove these as you fix them:}

Class revtex4 Warning: If you want your PACS to appear in your output, use docu
ment class option showpacs.


LaTeX Warning: Label `MFdtheta' multiply defined.


LaTeX Warning: Label `EqMotionMovFramePC' multiply defined.


LaTeX Warning: Label `PCsectSin' multiply defined.


LaTeX Warning: Label `fig:CLEpcSect' multiply defined.


LaTeX Warning: Label `EqMotionMovFrame' multiply defined.

LaTeX Warning: Reference `f:MeanVelocityFrame' on page 6 undefined on input lin
e 199.


LaTeX Warning: Reference `chap:lasers' on page 13 undefined on input line 181.


LaTeX Warning: Reference `eq:FiniteRot' on page 24 undefined on input line 128.


LaTeX Warning: Reference `fig:PCunrot' on page 26 undefined on input line 303.


LaTeX Warning: Reference `fig:PCunrot' on page 26 undefined on input line 306.


LaTeX Warning: Reference `fig:PCunrot' on page 26 undefined on input line 323.


LaTeX Warning: Reference `sec:SymmDyn' on page 26 undefined on input line 324.


LaTeX Warning: Reference `fig:CLEx1x2z' on page 27 undefined on input line 427.


LaTeX Warning: Reference `fig:PCunrot1' on page 28 undefined on input line 500.


LaTeX Warning: Reference `LorenzR' on page 29 undefined on input line 561.


LaTeX Warning: Reference `Lorenz' on page 29 undefined on input line 561.


LaTeX Warning: Reference `inftmInv' on page 29 undefined on input line 561.


LaTeX Warning: Reference `sec:StabEq' on page 29 undefined on input line 561.

---------------------------------- BibTeX -------------


change the style file to Physica D (currently apsrev.bst)

../bibtex/siminos.bib
\\
Warning--I didn't find a database entry for "endnote50"

