% siminos/CLE/CLE.tex
% $Author$ $Date$

                            \newif\ifdraft \newif\ifpaper
  \drafttrue\paperfalse      % draft version, commented
%   \draftfalse\paperfalse     % final version, hyperlinks
% \draftfalse\papertrue      % final version, no hyperlinks, for printing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Predrag reorganized siminos/CLE/          2009-10-09
% Vaggelis created siminos/CLE/CLE.tex      2009-09-21
% Vaggelis created siminos/CLE              2009-02-05

\ifdraft
\documentclass[preprint,number,sort&compress]{elsarticle}
\else
\documentclass[final,number]{elsarticle}
\fi

\usepackage{amsmath,amsfonts,amssymb,amsbsy,amscd}
\usepackage{ifthen}
\usepackage{graphicx}
\usepackage[dvips]{color}
\usepackage[dvips,colorlinks]{hyperref}
\graphicspath{{../figs/}{../Fig/}}  %% directories with graphics files
\input ../inputs/def        % TEMPORARY
\input inputs/defsCLE.tex   % eventualy merge used commanDs into inputs/defsCLE.tex

% Use the next command to get Bibliography header appear. The problem is a conflict
% between elsarticle and amsref over \bibsection command.
 \renewcommand\bibname{References}
 \newcommand\bibsection{%
   \section*{\bibname\markright{\MakeUppercase{\bibname}}}}

\begin{document}
\journal{Physica D}
\begin{frontmatter}

			\title{
Continuous symmetry reduction and return maps for higher-dimensional flows
			}
% Continuous symmetry reduction and return maps for high dimensional flows
% Continuous symmetry reduced return maps for high dimensional flows
% Returm maps of high dimensional flows with continuous symmetry:
%                                           I. Symmetry reduction
% Towards continuous symmetry reduction in high dimensional flows

\author{Evangelos Siminos}
\ead{siminos@gatech.edu}
\author{Predrag Cvitanovi\'c}
\address{Center for Nonlinear Science,
School of Physics, Georgia Institute of Technology,
Atlanta, GA 30332-0430}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}

\date{\today}

        \begin{abstract}
We present two continuous symmetry reduction methods for
reducing high-dimensional dissipative flows to local return
maps. In the Hilbert polynomial basis approach the equivariant
dynamics is rewritten in terms of invariant coordinates. In the
`method of moving frames' or `method of slices' the state space
is `sliced' locally in such a way that each group orbit of
symmetry-equivalent points is represented by a single point. In
either approach, numerical computations can be performed in the
original state-space representation, and the solutions are then
projected onto the symmetry-reduced state space. The two methods
are illustrated by reduction of the complex Lorenz system, a
5-dimensional dissipative flow with rotational symmetry. While
the Hilbert polynomial basis approach appears not feasible for
high-dimensional flows, the symmetry reduction by the `method of
moving frames' offers hope.
        \end{abstract}

\begin{keyword}
continuous symmetry reduction,
relative equilibria,
relative periodic orbits,
return maps,
slices,
moving frames,
Hilbert polynomial bases
\PACS 05.45.-a \sep 47.27.ed \sep 42.65.Sf
% 05.45.-a 	Nonlinear dynamics and chaos
% 05.45.Jn 	High-dimensional chaos
% 47.10.Fg 	Dynamical systems methods (in Fluid Mechanics)
% 47.27.ed 	Dynamical systems approaches (turbulent flows)
% 47.52.+j 	Chaos in fluid dynamics
% 42.65.Sf 	Dynamics of nonlinear optical systems; optical instabilities,
% 		optical chaos and complexity, and optical spatio-temporal dynamics
\end{keyword}
\end{frontmatter}

\section{\label{s:intro} Introduction}
    \PC{please keep improving abstract, intro and conclusions at each proofreading.
    \\
    remove PC\{\} and PCedit\{\} after you have accepted/edited them}
    \input intro

% \section{\label{s:symDyn} Symmetries of dynamical systems}
    \input symDyn

% \subsection{\label{s:introCLE} An example: \CLe}
    \input{introCLE}

% \section{\label{s:symSol} Solutions of systems with continuous symmetry}
   \input{symSol}
   \input{CLEsols}

\section{\label{s:Hilbert} Symmetry reduction: Hilbert polynomial bases}
    \input symRedGeneral
    \input Hilbert

\section{\label{sec:mf} Symmetry reduction: Moving frames}
    \input movingFrames
    \input mfReqb
    \input mfLocal
    \input slice


\section{Conclusions}
    \input conclusions

\section*{Acknowledgements}
We are grateful to
D.~Barkley,
W.-J.~Beyn,
R.~Gilmore,
J.~Halcrow,
K.A.~Mitchell,
C.W.~Rowley,
R.~Wilczak,
and in particular R.L.~Davidchack for many spirited exchanges,
and J.F.~Gibson for a critical reading of the manuscript.
P.C. thanks the
James Franck Institute, U. of Chicago,
for hospitality, and Argonne National Laboratory and
G.~Robinson Jr. for partial support.
E.S. was supported by NSF grant DMS-0807574 and G.~Robinson,~Jr.

% Specify following sections are appendices. Use \appendix* if there
% only one appendix.
\appendix

\PublicPrivate{}{
\section{\label{s:StabReq} Stability of \reqva}
    \input stabReqva
    } %end \PublicPrivate{}{

% \bibliographystyle{elsarticle-harv}
\bibliographystyle{elsarticle-num}
\bibliography{../bibtex/siminos}

\ifdraft
% remove this when all problems have been fixed
    \newpage
    \input 05fixMe
    \input flotsam
\fi
\end{document}
