% siminos/CLE/stabReqva.tex
% $Author$ $Date$

% From siminos/thesis/chapters/lasersSym.tex

%\subsection{Stability of \reqva}
%\label{s:StabReq}

In method of connections of Rowley \etal\rf{rowley_reduction_2003}, 
which we do not use for symmetry reduction because of
its additional geometric phase, the vector field $\vel(\ssp)$ 
is decomposed into a part along the group action tangent and a part $\velRed$ 
transverse to it. On a relative equilibrium the transverse
vector field vanishes and the method of connections
accrues no spurious geometric phase. Here we will use the same
decomposition to compute stability of \reqva\ in reduced space.
We only treat the one-dimensional group case and simplify
notation by dropping all group parameter related indices, 
writing $\Lg_{ij}$ instead of $(\Lg_1)_{ij}$, \etc. 
Our treatment is similar in spirit to that
of Chossat and Lauterbach\rf{ChossLaut00} but our final
prescription allows to calculate stability in reduced space
from the equi\-vari\-ant vector field in a manner that
we find more straightforward.


We introduce the projection operator
\beq
 	\PperpOp_{ij}(\ssp)=\delta_{ij}-
    \frac{\groupTan(\ssp)_i \groupTan(\ssp)_j}{\groupTan(\ssp)^2}
\ee{transvProj}
that projects a $d$-dimensional flow $v(\ssp)$ onto
flow
\beq
	\dot{\sspRed} = \velRed(\ssp) = \vel(\ssp)
    - \groupTan(\ssp) \frac{\groupTan(\ssp)^T \vel(\ssp)}{\groupTan(x)^2}
\ee{transvFlowSlice}
in a $(d\!-\!1)$-dimensional {\slice} transverse to the
direction fixed by the point $\ssp$.

The {\stabmat} $\bar{\Mvar}_{ij}$ for the flow in the reduced space 
is then given by
    \PC{Here the projection operator \refeq{transvProj} is OK,
    as the action of the group on $\ssp_{\REQV{}{1}}$ is trivial?
    Not sure...
    }
 \beq
	\bar{\Mvar}_{ij} = \frac{\partial}{\partial x_j}(\PperpOp  \vf)_i
		= \Pperp_{ik}\frac{\partial  \vf_k}{\partial x_j}
          +\frac{\partial \Pperp_{ik}}{\partial x_j} \vf_k
		\label{eq:stabreqvdef}
 \eeq
Now
\bea
	\frac{\partial \Pperp_{in}}{\partial x_j}  &=&  -\frac{\partial}{\partial x_j}\left(\frac{\groupTan(x)_i \groupTan(x)_n}{\groupTan(x)^2}\right)\continue
% 			&=& -\left(\frac{\Lg_{iq}\delta_{jq} \groupTan(x)_n}{\groupTan(x)^2}+\frac{\groupTan(x)_i \Lg_{n\ell}\delta_{j\ell}}{\groupTan(x)^2}-\frac{\groupTan(x)_i \groupTan(x)_n}{\groupTan(x)^4}\frac{\partial}{\partial x_j}\groupTan(x)^2 \right)\continue
			&=& -\left(\frac{\Lg_{ij} \groupTan(x)_n}{\groupTan(x)^2}+\frac{\groupTan(x)_i \Lg_{nj}}{\groupTan(x)^2}-2\frac{\groupTan(x)_i \groupTan(x)_n}{\groupTan(x)^4}\Lg_{mj}\groupTan(x)_m \right)\continue
% 			&=& -\frac{1}{\groupTan(x)^2}\left(\Lg_{ij}\groupTan(x)_n+\groupTan(x)_i \Lg_{nj}-2\frac{\groupTan(x)_i \groupTan(x)_n}{\groupTan(x)^2}\Lg_{mj}\groupTan(x)_m \right)\continue
			&=& -\frac{1}{\groupTan(x)^2}\left(\groupTan(x)_n\left(\Lg_{ij}-\frac{\groupTan(x)_i }{\groupTan(x)^2}\Lg_{mj}\groupTan(x)_m\right)+\groupTan(x)_i\left( \Lg_{nj}-\frac{ \groupTan(x)_n}{\groupTan(x)^2}\Lg_{mj}\groupTan(x)_m\right) \right)\continue
% 			&=& -\frac{1}{\groupTan(x)^2}\left(\groupTan(x)_n\left(\delta_{im}-\frac{\groupTan(x)_i }{\groupTan(x)^2}\groupTan(x)_m\right)\Lg_{mj}+\groupTan(x)_i\left( \delta_{nm}-\frac{ \groupTan(x)_n}{\groupTan(x)^2}\groupTan(x)_m\right)\Lg_{mj} \right)\continue
			&=& -\frac{1}{\groupTan(x)^2}\left(\groupTan(x)_n \Pperp_{im} \Lg_{mj}+\groupTan(x)_i \Pperp_{nm} \Lg_{mj} \right)\,.
\eea
Therefore \refeq{eq:stabreqvdef} takes the form
\beq
	\mathbf{\bar{\Mvar}}=\PperpOp \mathbf{A}
         -\frac{1}{\groupTan(x)^2}\left(
          \left[\vf^T \groupTan(x)\right]
          \left(\PperpOp \Lg\right) +\groupTan(x) \otimes
            \left[\vf^T \left( \PperpOp \Lg\right)\right]
          \right)\,,
	\label{eq:reqvStab}
\eeq
where $A_{ij}=\frac{\partial \vf_i}{\partial x_j}$. This
expression allows to calculate \reducedsp\ stability of
\reqva\ working in the equi\-vari\-ant variables, without
explicit knowledge of the vector field  $\velRed$ in reduced
space. Applying \refeq{eq:reqvStab} for \reqv\ \REQV{}{1} of
\cLe\ we obtain eigenvalues \refeq{eq:CLeREQBstab} we computed in
polar coordinates, along with an extra zero eigenvalue. This
extra eigenvalue appears since we work here in
equi\-vari\-ant variables for which the $\SOn{2}$ symmetry
marginal direction exists. Being able to compute stability
eigenvalues of \reqva\ directly in equi\-vari\-ant variables
is convenient in higher dimensions where an explicit
change of variables to invariant ones is often intractable.

