% acknowledgements.tex
% $Author$ $Date$

I would like to thank my advisor, Predrag Cvitanovi\'c,
for 
\PublicPrivate{patient guidance}
	          {relentless indoctrination, frequently reminding me that
I threw away Lan's papers}, constant encouragement \PublicPrivate{}{, insistence that research has to be beautiful} and 
dogged perfectionism.
\PublicPrivate{}{In retrospect, I greatly appreciate Yueheng Lan's trailblazing
work
, but the first thing I did when I took over his
graduate student desk was to throw away his complete collection of
article on \KS\  and related PDEs. For which I steel feel like a Nazi,
especially because I am reminded of it every once in a while,
but it is what Lan himself instruct me to do, giving me no indication
that those were useful papers. Which, useful papers, I suspect are now
on his desk in China.
              }
%
% People I have interacted with, learned from
Numerics did rarely converge until Ruslan L. Davidchack stepped in
and taught me how to do things correctly, always providing
useful advice and totaly changing the route of this thesis.
Yueheng Lan has put a lot of effort in bequeathing his knowledge 
and experience in the problem, and I am grateful to him for pointing out to the
existence of the \EQV{1}~\eqv\ at the $L=22$ system size among many
other tips. I have
profited from John F.~Gibson's insights into \statesp\
visualization of high-dimensional fluid flows,
discussions with
Jonathan Halcrow on desymmetrization of Lorenz and other
flows, and
I thank
G.~Robinson,~Jr.\,, National Science Foundation grant
DMS-0807574 and a Gerondelis Foundation Scholarship
for partial financial support.
% PC: have you ever used PACE? Nope.
%Also a special thanks to the Georgia Tech Student Union
%who believed in us where NSF feared to thread,
%and generously funded our accessa
%to the Georgia Tech Public Access Cluster Environment (GT-PACE),
%essential to the
%computationally demanding \KS\ numerics.

I would like to thank my parents for their firm support that 
laid down the foundations for everything. Kalli has been cheerful company while I was
writing this thesis and her keystrokes helped randomize the content.
%
% and now an ode to Penelope:
Last and most important, I would like to thank Sophia for her encouragement, support, patience and
understanding that were so valuable in every stage of this work and without which I wouldn't
have make it.