% Halcrow thesis summary
% $Author$ $Date$


Recent experimental and theoretical advances\rf{science04}
support a dynamical vision of turbulence:
For any finite  spatial resolution, a turbulent flow follows approximately for a finite time
a pattern belonging to a { finite alphabet} of admissible patterns.
The long term dynamics is a {walk through the space of these unstable patterns}.
The question is how to characterize and classify such patterns?
Here we follow the seminal Hopf paper\rf{hopf48}, and  visualize
turbulence not as  a sequence of spatial snapshots in turbulent evolution,
but as a trajectory in an  $\infty$-$d$ \statesp\ in which an
instant in turbulent evolution is a {unique} point. In the dynamical systems approach,
theory of turbulence for a given system, with given boundary conditions,
is given by (a) the geometry of the \statesp\ and (b) the associated natural measure,
\ie, the likelihood that asymptotic dynamics visits a given \statesp\ region.

In this thesis this vision is pursued in the context of \KS\ equation,
one of the simplest physically interesting spatially extended
nonlinear systems. Relaxing the restriction of previous studies\rf{LanThesis,Christiansen97}
to discrete symmetry invariant subspace a new obstacle confronts our attempt to elucidate
the \statesp\ geometry. Continuous symmetry, in the form of a subgroup of \On{n} acting on \statesp, 
endorses \statesp\ with additional structure. Symmetry dictates the bifurcation
structure and the type of observed solutions. At the same time recurrence
becomes relative: Not only do we have to identify points along the orbit of the 
nonlinear group of time evolution but also points along the orbit of the linear symmetry group.
The latter identification, termed symmetry reduction, although conceptualy simple as the group action is linear, 
is hard to implement in practice. The high dimensionality of phase space that results from truncation
of a PDE is one of the reasons. The second reason is that, for most interesting group actions,
the process introduces singularities in the structure of the reduced phase space. 

Here we propose a procedure to efficiently project dynamics of solutions computed in the original
space to a reduced phase space. This is done in conjunction with eliminating time-translational
invariance with suitably chosen Poincar\'e sections to avoid singularities while setting the stage
to compute return maps that describe how unstable manifolds of solutions organize the flow. 
The procedure gives as insight, through simplifying visualization, on which solutions of \KS\ equation
are important in describing the geometry of state space. Moreover, for a low-dimensional, strongly-contracting flow
its application leads to a one-dimensional return map that encapsulates all the information about the
dynamics.


