%Abstract Template


\documentclass[11pt]{article}
\usepackage{calc}
\usepackage{color}
\usepackage{amsfonts}
\usepackage{latexsym}
\usepackage{placeins}
\ifx\pdftexversion\undefined
  \usepackage[dvips]{graphicx}
\else
  \usepackage[pdftex]{graphicx}
\fi
\usepackage{amssymb}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage[cp1250]{inputenc}
\usepackage[OT4]{fontenc}

\addtolength{\voffset}{-3.5cm} \addtolength{\textheight}{4cm}

\renewcommand\Authfont{\scshape\small}
\renewcommand\Affilfont{\itshape\small}
\setlength{\affilsep}{1em}

\newcommand{\smalllineskip}{\baselineskip=15pt}
\newcommand{\keywords}[1]{{\footnotesize\hspace{0.68cm}{\textit{Keywords}: }#1\par
  \vskip.7\baselineskip}}
\renewenvironment{abstract}[0]{\small\rm
        \begin{center}ABSTRACT
        \\ \vspace{8pt}
        \begin{minipage}{5.2in}\smalllineskip
        \hspace{1pc}}{\end{minipage}\end{center}\vspace{-1pt}}
\newcommand{\emailaddress}[1]{\newline{\sf#1}}

\let\LaTeXtitle\title
\renewcommand{\title}[1]{\LaTeXtitle{\large\textsf{\textbf{#1}}}}

%%%TITLE
\title{Reduction of continuous SO(2) symmetry of a 2-mode system using method of slices}
\date{}

%%AFFILIATIONS
\author[1]{Nazmi Burak Budanur}
\author[1]{P. Cvitanovi\'{c}}
\affil[1]{School of Physics and Center for Nonlinear Dynamics,
		  Georgia Inst. of Technology,
		  Atlanta, GA  30332, USA \emailaddress{budanur3@gatech.edu}}


%%DOCUMENT
\begin{document}
\maketitle

%%PLEASE PUT YOUR ABSTRACT HERE
\begin{abstract}

Danglmayr~\cite{Dang86} and Porter~\&\ Knobloch~\cite{PoKno05} have 
introduced a family of 2-Fourier mode SO(2)-equivariant ODEs  in order to study 
bifurcations of solutions of dynamical systems in presence of symmetries. A 
4-dimensional system of this kind is perhaps the simplest example of a system 
with a continuous symmetry that can exhibit chaos, so we use it to illustrate 
the role symmetries play in chaotic dynamics. We show that a continuous 
symmetry induces drifts in the 4-dimensional state space dynamics, drifts which 
obscure the chaotic dynamics. Change of equations of motions to a locally 
symmetry-invariant `comoving' frame does not eliminate these drifts: that is 
only attained by a \emph{symmetry reduction} - reformulation of dynamics in a 
3-dimensional symmetry-reduced state space, where every group orbit (set of all 
points reached by actions of the group of all symmetries of the equations of 
motion) is replaced by a point. Porter~\&\ Knobloch system is a particularly 
nice illustration of how this works, as in 3 dimensions we are able to 
visualize everything.

We compare three symmetry reduction methods: polar coordinates, invariant 
polynomial bases, and the `method of slices'. An invariant polynomial basis is 
convenient for determination of all relative equilibria of such system. Our 
conclusion, however, is that the most insight is offered by the method of 
slices. While in general a number of local slices are needed to cover a 
strange attractor~\cite{atlas12}, for the Porter~\&\ Knobloch system there we 
define a unique slice hyperplane that captures \emph{all} symmetry-reduced 
dynamics. A Poincar\'e return map within the slice hyperplane enables us to 
reduce the dynamics further, essentially to a unimodal map, and determine, in 
principle, all relative periodic orbits of the system. We can visualize each 
step of this process without having to project solutions onto a submanifold 
since the slice hyperplane for this system is three dimensional. We argue that 
our method can reduce the SO(2)-symmetry also for $N$-Fourier modes truncations 
of PDEs such as the Kuramoto-Sivashinsky, pipe flows, etc., as long as the 
amplitude of the first Fourier mode is non-zero.

\end{abstract}
%%THE END OF ABSTRACT

\begin{thebibliography}{99}
\small

\bibitem{Dang86} 
G. Danglmayr,  {\it Dyn. Sys.} {\bf 1}, pp. 159--185 (1986).

\bibitem{PoKno05} 
J. Porter and E. Knobloch, {\it Physica D} {\bf 201}, pp. 318--334 (2005).

\bibitem{atlas12} 
P. Cvitanovi\'{c}, D. Borrero-Echeverry, K. Carroll, B. Robbins and
E. Siminos, {\it Chaos} {\bf 22}, 047506 (2012).

\end{thebibliography}
\end{document}
