\svnkwsave{$RepoFile: siminos/baroclinic/OrtegaBlog.tex $}
\svnidlong {$HeadURL$}
{$LastChangedDate$}
{$LastChangedRevision$} {$LastChangedBy$}
\svnid{$Id$}

\chapter{Convectively coupled waves}
\label{chap:OrtegaBlog}

\section{Introduction}
\label{sect:CCWs}

\begin{description}

\item[2012-02-08 Predrag] Please write an introduction to
the ``Convectively coupled waves'' suitable to inclusion into
your thesis: what they are, and why should one care. Add relevant references
to this repository's siminos/bibtex/siminos.bib (no separate bibliographies,
it makes updating them a pain).

\end{description}

\section{Sebastian's blog}
\label{sect:OrtegaDaily}

This is Sebastian Ortega Arango's blog for PHYS 7224,  spring 2012
\emph{Nonlinear dynamics: Chaos, and what to do about it?} course project

\begin{description}

\item[2012-02-07 Predrag] Added Sebastian Ortega Arango
<sortega@gatech.edu> project to this blog; enabled svn access, updates
(sortega  baroclinic). First task: please study \refchap{chap:baroclinic},
and improve it in any way you see fit, edit, add better references etc.

\item[2012-02-07 Sebastian]
I think I would be able to relate the project with my line of research. I
am interested in \textbf{Convectively Coupled Waves}. I read that people at NYU
has worked the mathematics in this types of phenomena, and it appears
that interesting nonlinearities arise from it. Baroclinic instability
seems to bee an important driver.

\item[2012-02-08 Predrag] I would like the project to focus on a
geophysically important model that exhibits $\SOn{2}$ invariance,
hence propose to investigate the baroclinic instability first, then apply
that to your thesis research on Convectively Coupled Waves. Advantage of
starting with the baroclinic instability is that you can hit the ground running,
as Annalisa has simulation code ready to use.

\item[2011-10-15 Annalisa] Define and explain the Rossby radius for the
atmosphere in \refsect{sect:CCWs}. Mark here [~~] when done.

\item[2012-03-27 Sebastian] I am going to follow your advice and star
working with baroclinic stability first. I have written a draft of the
introduction, the idea is that it outlines the work to be done, and is
subject to changes. I think it is a good outline to introduce baroclinic
instability, and to introduce the model used in the simulations. However,
it still not clear to me how to include the nonlinear analysis. I will
continue reading about the \KSe\ and the \pCf\ for this. I will try to
write \refsect{s:intro} today.

\item[2012-03-27 Predrag] I'm glad you are getting started - go for it.

\item[2012-03-27 Sebastian]
Also, I find a thesis that might be interesting to read Veen's thesis,
\HREF{http://igitur-archive.library.uu.nl/dissertations/2002-0801-151812/c2.pdf}
{Chapter 2}.

\item[2012-03-27 Predrag] I have not read Lennart's thesis, but he does good work. Blog
here what you find of interest as you read it.

\item[2012-04-04 Sebastian]
Finished first and second section. I am uploading to show what I have done so far, but it is just a quick draft as far as redaction goes. I still have to read it again and correct it. Some times I write in this way, first I get all the ideas in paper and then iterate until I have a coherent document. Probably not the best way out there.

In short, the point that I am trying to get trough, is that there is a base flow given by geostrophy and the hydrostatic relation which might be stable depending on the slope of the isopycnals. Then I introduce the model Professor Annalisa used for her simulations (I believe is the one described by Philips in 1951).

I think it is important for me to get focus on the nonlinear aspects of the problem. So I will try to finish all the way to section \ref{s:stability} by this week (but not sure if I will have the time).

I was also wondering what should I do for the nonlinear section. But I guess I should first order my ideas and learn how to think of baroclinic instability in terms of dynamical systems. So far I think of it from the point of view of bifurcations (please correct me if wrong), where I have a equilibrium point in the dynamical system which becomes unstable (or disappears??) after some parameter increases (this is what I believe the linear theory does). But I am not sure what happens once the flow is unstable, my guess is that there would be some kind of strange attractor out there in the n-dimensional space. But I guess visualization of this would be quite hard, unless a low order spectral representation is used. It would be very interesting to hear your vision of the instability; so please let me know if there is a paper I can read for this, or what you think is happening.

Also, let me know of any changes that you think might be convenient.

\item[2012-04-05 Sebastian]
A quote from Rick Salmon book (might be good for Chaos book). After proving Ertel's Theorem: "Of course, we can prove all these results directly from (1.1) \footnote{Momentum equation, continuity equation, thermodynamic equation and equation of state} by pedestrian mathematical manipulations, but that only makes it harder to appreciate their physical significance"
\end{description}
